% Auto-generated from raw.txt
% Usage:
%   \vocabentry{word}{/ipa/ (pos)}{definition}{example}{note}

\vocabentry{Curriculum}
{/kəˈrɪkjələm/ (n)}
{The subjects comprising a course of study in a school or college.}
{The ministry is updating the national curriculum to include more digital literacy skills.}
{Đây là một danh từ dùng để chỉ chương trình giảng dạy tổng thể của một cơ sở giáo dục.}

\vocabentry{Pedagogy}
{/ˈpedəɡɑːdʒi/ (n)}
{The method and practice of teaching, especially as an academic subject or theoretical concept.}
{Innovative pedagogy can significantly improve student engagement in the classroom.}
{Đây là một danh từ chuyên ngành dùng để chỉ phương pháp và lý luận giảng dạy sư phạm.}

\vocabentry{Academic}
{/ˌækəˈdemɪk/ (adj)}
{Relating to education and scholarship.}
{The university provides a strong academic environment for research students.}
{Đây là một tính từ dùng để chỉ các vấn đề thuộc về học thuật, lý thuyết hoặc trường lớp.}

\vocabentry{Literacy}
{/ˈlɪtərəsi/ (n)}
{The ability to read and write.}
{The government is launching a new campaign to improve adult literacy rates in rural areas.}
{Đây là một danh từ dùng để chỉ trình độ học vấn cơ bản hoặc khả năng biết đọc biết viết.}

\vocabentry{Syllabus}
{/ˈsɪləbəs/ (n)}
{An outline of the subjects in a course of study or examination.}
{Students receive a copy of the syllabus on the first day of the semester.}
{Đây là một danh từ dùng để chỉ đề cương chi tiết của một môn học cụ thể.}

\vocabentry{Assessment}
{/əˈsesmənt/ (n)}
{The evaluation or estimation of the nature, quality, or ability of someone.}
{Continuous assessment is often more effective than a single final exam.}
{Đây là một danh từ dùng để chỉ sự đánh giá năng lực hoặc quá trình kiểm tra.}

\vocabentry{Compulsory}
{/kəmˈpʌlsəri/ (adj)}
{Required by law or a rule; obligatory.}
{Mathematics and English are compulsory subjects in most high schools.}
{Đây là một tính từ dùng để chỉ các môn học hoặc chương trình giáo dục bắt buộc.}

\vocabentry{Tertiary}
{/ˈtɜːrʃieri/ (adj)}
{Relating to education at university or college level.}
{Access to tertiary education has expanded significantly over the last decade.}
{Đây là một tính từ dùng để chỉ giáo dục bậc cao hoặc giáo dục đại học.}

\vocabentry{Vocational}
{/voʊˈkeɪʃənl/ (adj)}
{Relating to an occupation or employment.}
{Vocational training provides students with practical skills for the workplace.}
{Đây là một tính từ dùng để chỉ giáo dục nghề nghiệp, chuẩn bị kỹ năng cho một công việc cụ thể.}

\vocabentry{Plagiarism}
{/ˈpleɪdʒərɪzəm/ (n)}
{The practice of taking someone else's work or ideas and passing them off as one's own.}
{Plagiarism is a serious academic offense that can lead to expulsion from the university.}
{Đây là một danh từ dùng để chỉ sự đạo văn trong học thuật.}

\vocabentry{Dissertation}
{/ˌdɪsərˈteɪʃn/ (n)}
{A long essay on a particular subject, especially one written as a requirement for a university degree.}
{He is currently writing his doctoral dissertation on environmental law.}
{Đây là một danh từ dùng để chỉ luận văn hoặc luận án tốt nghiệp.}

\vocabentry{Enrolment}
{/ɪnˈroʊlmənt/ (n)}
{The action of enrolling or being enrolled.}
{The school has seen a 20\% increase in student enrolment this year.}
{Đây là một danh từ dùng để chỉ sự đăng ký nhập học hoặc tổng số lượng học sinh nhập học.}

\vocabentry{Faculty}
{/ˈfæklti/ (n)}
{A group of university departments concerned with a major division of knowledge.}
{She is a member of the Faculty of Arts and Sciences.}
{Đây là một danh từ dùng để chỉ khoa hoặc toàn bộ giảng viên của một trường đại học.}

\vocabentry{Tuition}
{/tuˈɪʃn/ (n)}
{Sum of money charged for teaching or instruction by a school.}
{High tuition fees have become a major barrier for many students.}
{Đây là một danh từ dùng để chỉ học phí hoặc sự giảng dạy.}

\vocabentry{Scholarship}
{/ˈskɑːlərʃɪp/ (n)}
{A grant or payment made to support a student's education, awarded on the basis of academic or other achievement.}
{She won a full scholarship to study medicine at Harvard University.}
{Đây là một danh từ dùng để chỉ học bổng.}

\vocabentry{Assignment}
{/əˈsaɪnmənt/ (n)}
{A task or piece of work assigned to someone as part of a course.}
{The deadline for the history assignment is next Friday.}
{Đây là một danh từ dùng để chỉ bài tập về nhà hoặc nhiệm vụ học tập được giao.}

\vocabentry{Tutorial}
{/tuːˈtɔːriəl/ (n)}
{A period of tuition given by a university tutor to an individual or a very small group.}
{I have a weekly tutorial where I can discuss my research problems with my supervisor.}
{Đây là một danh từ dùng để chỉ giờ hướng dẫn kèm cặp của giáo viên với nhóm nhỏ.}

\vocabentry{Qualification}
{/ˌkwɑːlɪfɪˈkeɪʃn/ (n)}
{A pass of an examination or an official record of achievement.}
{Having a teaching qualification is essential for working in public schools.}
{Đây là một danh từ dùng để chỉ bằng cấp hoặc chứng chỉ chuyên môn.}

\vocabentry{Discipline}
{/ˈdɪsəplɪn/ (n)}
{The practice of training people to obey rules; also a branch of knowledge.}
{Maintaining classroom discipline is essential for effective learning.}
{Đây là một danh từ dùng để chỉ kỷ luật hoặc một ngành học.}

\vocabentry{Interactive}
{/ˌɪntərˈæktɪv/ (adj)}
{Involving the exchange of information or ideas between people.}
{Interactive whiteboards have made lessons more engaging for children.}
{Đây là một tính từ dùng để chỉ tính tương tác trong phương pháp giảng dạy.}

\vocabentry{Holistic}
{/hoʊˈlɪstɪk/ (adj)}
{Characterized by the belief that the parts of something are intimately interconnected.}
{A holistic approach to education focuses on both academic and emotional development.}
{Đây là một tính từ chỉ phương pháp giáo dục toàn diện cả về trí tuệ lẫn tâm hồn.}

\vocabentry{Pedagogical}
{/ˌpedəˈɡɑːdʒɪkl/ (adj)}
{Relating to teaching.}
{Teachers are attending workshops to enhance their pedagogical skills.}
{Đây là một tính từ dùng để chỉ các yếu tố liên quan đến nghiệp vụ sư phạm.}

\vocabentry{Credential}
{/krəˈdenʃl/ (n)}
{A qualification, achievement, or aspect of a person's background.}
{You need strong academic credentials to apply for this research position.}
{Đây là một danh từ dùng để chỉ chứng chỉ hoặc năng lực chuyên môn được công nhận.}

\vocabentry{Alumni}
{/əˈlʌmnaɪ/ (n)}
{Former students or graduates of a particular school, college, or university.}
{The university's alumni often donate money to support new scholarships.}
{Đây là một danh từ dùng để chỉ các cựu sinh viên.}

\vocabentry{Enrichment}
{/ɪnˈrɪtʃmənt/ (n)}
{The action of improving or enhancing the quality or value of something.}
{Extracurricular activities provide mental enrichment for students.}
{Đây là một danh từ dùng để chỉ hoạt động bổ trợ hoặc làm giàu kiến thức.}

\vocabentry{Grant}
{/ɡrænt/ (n)}
{A sum of money given by a government or organization for a particular purpose.}
{The research team received a government grant to study climate change.}
{Đây là một danh từ dùng để chỉ khoản trợ cấp nghiên cứu hoặc học tập.}

\vocabentry{Illiteracy}
{/ɪˈlɪtərəsi/ (n)}
{The inability to read or write.}
{Eradicating illiteracy remains a top priority in developing nations.}
{Đây là một danh từ dùng để chỉ tình trạng mù chữ.}

\vocabentry{Instruction}
{/ɪnˈstrʌkʃn/ (n)}
{The act of teaching or providing someone with information.}
{Clear instruction from the teacher is vital for student success.}
{Đây là một danh từ dùng để chỉ sự hướng dẫn hoặc giảng dạy.}

\vocabentry{Knowledge-based}
{/ˈnɑːlɪdʒ beɪst/ (adj)}
{Relying on or using specialized knowledge.}
{Modern societies are becoming increasingly knowledge-based.}
{Đây là một tính từ dùng để chỉ các lĩnh vực hoặc nền kinh tế dựa trên kiến thức.}

\vocabentry{Mentor}
{/ˈmentɔːr/ (n)}
{An experienced and trusted adviser.}
{Having a mentor in college can help students make better career choices.}
{Đây là một danh từ dùng để chỉ người cố vấn hoặc người hướng dẫn giàu kinh nghiệm.}

\vocabentry{Objective}
{/əbˈdʒektɪv/ (n)}
{A thing aimed at or sought; a goal.}
{The main objective of this course is to improve students' writing skills.}
{Đây là một danh từ dùng để chỉ mục tiêu học tập hoặc đào tạo.}

\vocabentry{Prerequisite}
{/ˌpriːˈrekwəzɪt/ (n)}
{A thing that is required as a prior condition for something else to happen.}
{Passing Level 1 is a prerequisite for taking the Level 2 course.}
{Đây là một danh từ dùng để chỉ điều kiện tiên quyết để tham gia một khóa học.}

\vocabentry{Rote learning}
{/roʊt ˈlɜːrnɪŋ/ (n)}
{A memorization technique based on repetition.}
{Rote learning is often criticized for not encouraging critical thinking.}
{Đây là một danh từ dùng để chỉ việc học vẹt, học thuộc lòng máy móc.}

\vocabentry{Specialized}
{/ˈspeʃəlaɪzd/ (adj)}
{Requiring or involving detailed and specific knowledge or training.}
{Many graduates pursue specialized degrees to advance their careers.}
{Đây là một tính từ dùng để chỉ tính chuyên môn hóa cao.}

\vocabentry{Standardized test}
{/ˈstændərdaɪzd test/ (n)}
{A test that is administered and scored in a consistent manner.}
{Many universities use standardized tests to evaluate applicants.}
{Đây là một danh từ dùng để chỉ bài kiểm tra tiêu chuẩn hóa như IELTS, SAT.}

\vocabentry{Undergraduate}
{/ˌʌndərˈɡrædʒuət/ (n)}
{A student at a college or university who has not yet earned a bachelor's degree.}
{Many undergraduates participate in sports clubs and student organizations.}
{Đây là một danh từ dùng để chỉ sinh viên đại học chưa tốt nghiệp.}

\vocabentry{Graduate}
{/ˈɡrædʒueɪt/ (v/n)}
{To successfully complete a course of training; a person who has done so.}
{After four years of hard work, he finally graduated with honors.}
{Đây là một động từ chỉ sự tốt nghiệp hoặc danh từ chỉ người đã tốt nghiệp.}

\vocabentry{Postgraduate}
{/ˌpoʊstˈɡrædʒuət/ (n)}
{A student who has already received a degree and is working towards a higher degree.}
{Postgraduates often work as teaching assistants while doing their research.}
{Đây là một danh từ dùng để chỉ học viên sau đại học - Thạc sĩ/Tiến sĩ.}

\vocabentry{Workshop}
{/ˈwɜːrkʃɑːp/ (n)}
{A meeting at which a group of people engage in intensive discussion and activity.}
{I attended a workshop on digital marketing last weekend.}
{Đây là một danh từ dùng để chỉ buổi hội thảo thực hành hoặc tập huấn.}

\vocabentry{Lecture}
{/ˈlektʃər/ (n/v)}
{An educational talk to an audience; to deliver such a talk.}
{The professor gave an inspiring lecture on the future of renewable energy.}
{Đây là một danh từ/động từ chỉ bài giảng hoặc việc thuyết giảng ở đại học.}

\vocabentry{Seminar}
{/ˈsemɪnɑːr/ (n)}
{A class in which a small group of students discuss a subject with a teacher.}
{The history seminar focused on the causes of the industrial revolution.}
{Đây là một danh từ dùng để chỉ buổi hội thảo hoặc giờ thảo luận nhóm.}

\vocabentry{Boarding school}
{/ˈbɔːrdɪŋ skuːl/ (n)}
{A school where students reside during the term.}
{Boarding schools can help students develop independence from a young age.}
{Đây là một danh từ dùng để chỉ trường nội trú.}

\vocabentry{Co-educational}
{/ˌkoʊ edʒuˈkeɪʃənl/ (adj)}
{The education of pupils of both sexes together.}
{Most modern schools are co-educational institutions.}
{Đây là một tính từ dùng để chỉ môi trường giáo dục nam nữ học chung.}

\vocabentry{Distance learning}
{/ˈdɪstəns ˈlɜːrnɪŋ/ (n)}
{A method of studying in which classes are conducted over the internet.}
{Distance learning has become increasingly popular due to its flexibility.}
{Đây là một danh từ dùng để chỉ việc học từ xa.}

\vocabentry{Extracurricular}
{/ˌekstrəkəˈrɪkjələr/ (adj)}
{Pursued in addition to the normal course of study.}
{Participating in extracurricular activities can help students develop leadership skills.}
{Đây là một tính từ dùng để chỉ các hoạt động ngoại khóa.}

\vocabentry{Lifelong learning}
{/ˌlaɪflɔːŋ ˈlɜːrnɪŋ/ (n)}
{The ongoing search for knowledge for either personal or professional reasons.}
{Lifelong learning is essential in a rapidly changing world.}
{Đây là một danh từ dùng để chỉ việc học tập suốt đời.}

\vocabentry{Primary education}
{/ˈpraɪməri ˌedʒuˈkeɪʃn/ (n)}
{The first stage of formal education, typically for children aged 5 to 11.}
{Primary education is essential for building a strong foundation of knowledge.}
{Đây là một danh từ dùng để chỉ giáo dục tiểu học.}

\vocabentry{Secondary education}
{/ˈsekənderi ˌedʒuˈkeɪʃn/ (n)}
{Education beyond the elementary level, typically for ages 11 to 18.}
{In the UK, secondary education ends after students take their GCSEs.}
{Đây là một danh từ dùng để chỉ giáo dục trung học.}

\vocabentry{Technical school}
{/ˈteknɪkl skuːl/ (n)}
{A school providing instruction in specialized practical skills.}
{Technical schools prepare students for careers in fields like engineering and IT.}
{Đây là một danh từ dùng để chỉ trường kỹ thuật hoặc trường nghề.}

\vocabentry{Academic performance}
{/ˌækəˈdemɪk pərˈfɔːrməns/ (n)}
{The extent to which a student has achieved their educational goals.}
{Stress can negatively affect a student's academic performance.}
{Đây là một cụm danh từ dùng để chỉ kết quả hoặc thành tích học tập.}

\vocabentry{Cognitive}
{/ˈkɑːɡnətɪv/ (adj)}
{Relating to the mental action or process of acquiring knowledge.}
{Cognitive development is a key focus in early childhood education.}
{Đây là một tính từ dùng để chỉ các hoạt động thuộc về nhận thức và tư duy.}

\vocabentry{Continuous assessment}
{/kənˈtɪnjuəs əˈsesmənt/ (n)}
{Evaluation of a student's progress throughout a course.}
{Continuous assessment helps teachers identify students who need extra help.}
{Đây là một cụm danh từ dùng để chỉ việc đánh giá thường xuyên.}

\vocabentry{Critical thinking}
{/ˈkrɪtɪkl ˈθɪŋkɪŋ/ (n)}
{The objective analysis and evaluation of an issue in order to form a judgment.}
{Universities aim to develop students' critical thinking skills.}
{Đây là một cụm danh từ dùng để chỉ tư duy phản biện.}

\vocabentry{Diploma}
{/dɪˈploʊmə/ (n)}
{A certificate awarded by an educational institution for a course of study.}
{He earned a diploma in graphic design from a technical college.}
{Đây là một danh từ dùng để chỉ văn bằng hoặc chứng chỉ tốt nghiệp.}

\vocabentry{Dropout}
{/ˈdrɑːpaʊt/ (n)}
{A person who leaves school or college before finishing their studies.}
{Economic pressure forced many students to become dropouts.}
{Đây là một danh từ dùng để chỉ người bỏ học giữa chừng.}

\vocabentry{Elective}
{/ɪˈlektɪv/ (n)}
{An optional course of study.}
{In the second year, students can choose two electives based on their interests.}
{Đây là một danh từ dùng để chỉ các môn học tự chọn.}

\vocabentry{Formal education}
{/ˈfɔːrml ˌedʒuˈkeɪʃn/ (n)}
{Education normally delivered by trained teachers in a systematic way.}
{Formal education is just one part of a person's overall development.}
{Đây là một cụm danh từ dùng để chỉ giáo dục chính quy.}

\vocabentry{Higher education}
{/ˌhaɪər ˌedʒuˈkeɪʃn/ (n)}
{Education at universities or similar institutions.}
{More students are pursuing higher education than ever before.}
{Đây là một cụm danh từ dùng để chỉ giáo dục bậc cao/đại học.}

\vocabentry{Major}
{/ˈmeɪdʒər/ (n)}
{The main subject that a student studies at college or university.}
{Her major is international relations, and she hopes to work for the UN.}
{Đây là một danh từ dùng để chỉ chuyên ngành chính.}

\vocabentry{Master’s degree}
{/ˈmæstərz dɪˈɡriː/ (n)}
{A second-cycle academic degree awarded by universities.}
{He is studying for a master's degree in business administration.}
{Đây là một cụm danh từ dùng để chỉ bằng thạc sĩ.}

\vocabentry{Minor}
{/ˈmaɪnər/ (n)}
{A secondary subject that a student studies at university.}
{She chose history as her major and French as her minor.}
{Đây là một danh từ dùng để chỉ chuyên ngành phụ.}

\vocabentry{Orientation}
{/ˌɔːriənˈteɪʃn/ (n)}
{A program designed to introduce new students to a school.}
{The orientation week helps new students get used to the campus layout.}
{Đây là một danh từ dùng để chỉ buổi hướng dẫn hoặc định hướng nhập học.}

\vocabentry{Professor}
{/prəˈfesər/ (n)}
{A university teacher of the highest rank.}
{The professor published several groundbreaking papers on genetics.}
{Đây là một danh từ dùng để chỉ giáo sư.}

\vocabentry{Research}
{/rɪˈsɜːrtʃ/ (n/v)}
{The systematic investigation into a subject; to carry out such investigation.}
{The university is a leading institution in the field of medical research.}
{Đây là một danh từ/động từ dùng để chỉ việc nghiên cứu.}

\vocabentry{Student-centered}
{/ˈstuːdnt ˈsentərd/ (adj)}
{Focusing on the needs and interests of the students.}
{Student-centered learning encourages pupils to take an active role.}
{Đây là một tính từ dùng để chỉ phương pháp lấy học sinh làm trung tâm.}

\vocabentry{Thesis}
{/ˈθiːsɪs/ (n)}
{A long essay involving personal research, written for a university degree.}
{She is defending her master's thesis next month.}
{Đây là một danh từ dùng để chỉ luận văn tốt nghiệp.}

\vocabentry{Transcript}
{/ˈtrænskrɪpt/ (n)}
{An official record of a student's work, showing courses and grades.}
{You need to submit your academic transcript when applying for a visa.}
{Đây là một danh từ dùng để chỉ bảng điểm học tập.}

\vocabentry{Abilities}
{/əˈbɪlɪtiz/ (n)}
{Possession of the means or skill to do something.}
{The test is designed to measure students' mathematical abilities.}
{Đây là một danh từ số nhiều dùng để chỉ năng lực hoặc khả năng.}

\vocabentry{Absenteeism}
{/ˌæbsənˈtiːɪzəm/ (n)}
{The practice of regularly staying away from work or school without good reason.}
{High levels of absenteeism can lead to poor academic results.}
{Đây là một danh từ dùng để chỉ tình trạng trốn học hoặc vắng mặt không lý do.}

\vocabentry{Accreditation}
{/əˌkredɪˈteɪʃn/ (n)}
{Official recognition that an institution meets certain standards.}
{The college is seeking international accreditation for its business program.}
{Đây là một danh từ dùng để chỉ sự kiểm định hoặc công nhận chất lượng giáo dục.}

\vocabentry{Achievement}
{/əˈtʃiːvmənt/ (n)}
{A thing done successfully, typically by effort or skill.}
{Graduation is a significant achievement for any student.}
{Đây là một danh từ dùng để chỉ thành tựu hoặc kết quả đạt được.}

\vocabentry{Aptitude}
{/ˈæptɪtuːd/ (n)}
{A natural ability to do something.}
{He showed an early aptitude for learning foreign languages.}
{Đây là một danh từ dùng để chỉ năng khiếu hoặc khả năng tự nhiên.}

\vocabentry{Autonomous}
{/ɔːˈtɑːnəməs/ (adj)}
{Having the freedom to govern itself or control its own affairs.}
{The university encourages students to become autonomous learners.}
{Đây là một tính từ dùng để chỉ tính tự chủ trong học tập.}

\vocabentry{Bachelor's degree}
{/ˈbætʃələrz dɪˈɡriː/ (n)}
{An undergraduate degree awarded by colleges and universities.}
{She received her bachelor's degree in economics last June.}
{Đây là một cụm danh từ dùng để chỉ bằng cử nhân.}

\vocabentry{Campus}
{/ˈkæmpəs/ (n)}
{The grounds and buildings of a university or college.}
{The university campus has a large library and several sports facilities.}
{Đây là một danh từ dùng để chỉ khuôn viên trường học.}

\vocabentry{Chancellor}
{/ˈtʃænsələr/ (n)}
{The head of some universities.}
{The Chancellor delivered a speech at the graduation ceremony.}
{Đây là một danh từ dùng để chỉ hiệu trưởng hoặc giám đốc đại học.}

\vocabentry{Classroom}
{/ˈklæsruːm/ (n)}
{A room in which a class of pupils or students is taught.}
{Modern classrooms are equipped with digital projectors and Wi-Fi.}
{Đây là một danh từ dùng để chỉ lớp học.}

\vocabentry{Competence}
{/ˈkɑːmpɪtəns/ (n)}
{The ability to do something successfully or efficiently.}
{Students must demonstrate competence in a foreign language to graduate.}
{Đây là một danh từ dùng để chỉ năng lực hoặc sự thành thạo.}

\vocabentry{Dean}
{/diːn/ (n)}
{The head of a college or university faculty.}
{The Dean of the Medical Faculty is a world-renowned surgeon.}
{Đây là một danh từ dùng để chỉ trưởng khoa.}

\vocabentry{Degree}
{/dɪˈɡriː/ (n)}
{An academic rank conferred by a college or university.}
{He is working hard to earn a degree in computer science.}
{Đây là một danh từ dùng để chỉ bằng cấp đại học.}

\vocabentry{Education}
{/ˌedʒuˈkeɪʃn/ (n)}
{The process of receiving or giving systematic instruction.}
{Education is the most powerful tool for changing the world.}
{Đây là một danh từ dùng để chỉ nền giáo dục hoặc sự giáo dục.}

\vocabentry{Educator}
{/ˈedʒukeɪtər/ (n)}
{A person who provides instruction or education; a teacher.}
{Maria Montessori was a famous Italian educator and physician.}
{Đây là một danh từ dùng để chỉ nhà giáo dục.}

\vocabentry{Elementary}
{/ˌelɪˈmentri/ (adj)}
{Relating to the most basic stages of a subject.}
{The elementary level of the language course focuses on simple grammar.}
{Đây là một tính từ dùng để chỉ mức độ cơ bản hoặc tiểu học.}

\vocabentry{Equality}
{/iˈkwɑːləti/ (n)}
{The state of being equal, especially in status, rights, and opportunities.}
{The government is committed to promoting equality in schools.}
{Đây là một danh từ dùng để chỉ sự bình đẳng trong giáo dục.}

\vocabentry{Examination}
{/ɪɡˌzæmɪˈneɪʃn/ (n)}
{A formal test of a person's knowledge or proficiency.}
{Students are preparing for their final examinations in June.}
{Đây là một danh từ dùng để chỉ kỳ thi hoặc bài kiểm tra.}

\vocabentry{Expulsion}
{/ɪkˈspʌlʃn/ (n)}
{The action of depriving someone of membership in an organization.}
{Bullying can lead to the expulsion of the student involved.}
{Đây là một danh từ dùng để chỉ sự đuổi học.}

\vocabentry{Fellowship}
{/ˈfeloʊʃɪp/ (n)}
{A grant of money for research or further study.}
{He was awarded a prestigious fellowship to study at Oxford.}
{Đây là một danh từ dùng để chỉ học bổng nghiên cứu dành cho nghiên cứu sinh.}

\vocabentry{Foundation}
{/faʊnˈdeɪʃn/ (n)}
{The lowest load-bearing part of a building; an underlying basis.}
{Primary school provides the foundation for all future learning.}
{Trong giáo dục dùng để chỉ kiến thức nền tảng.}

\vocabentry{Genius}
{/ˈdʒiːniəs/ (n)}
{Exceptional intellectual or creative power or other natural ability.}
{Albert Einstein was widely recognized as a scientific genius.}
{Đây là một danh từ dùng để chỉ thiên tài.}

\vocabentry{Grade}
{/ɡreɪd/ (n)}
{A mark indicating the quality of a student's work.}
{She was disappointed with her grade in the math test.}
{Đây là một danh từ dùng để chỉ điểm số hoặc lớp học.}

\vocabentry{Handout}
{/ˈhændaʊt/ (n)}
{A piece of printed information provided free of charge.}
{The teacher distributed a handout containing the main points of the lecture.}
{Đây là một danh từ dùng để chỉ tài liệu phát tay trong lớp học.}

\vocabentry{Headmaster}
{/ˌhedˈmæstər/ (n)}
{A man who is the head teacher of a school.}
{The headmaster spoke to the students about the importance of discipline.}
{Đây là một danh từ dùng để chỉ hiệu trưởng nam.}

\vocabentry{Headmistress}
{/ˌhedˈmɪstrəs/ (n)}
{A woman who is the head teacher of a school.}
{The headmistress is very supportive of the school's sports teams.}
{Đây là một danh từ dùng để chỉ hiệu trưởng nữ.}

\vocabentry{Home-schooling}
{/ˌhoʊm ˈskuːlɪŋ/ (n)}
{The education of children at home by their parents.}
{Home-schooling has become more common during the global pandemic.}
{Đây là một danh từ dùng để chỉ việc tự giáo dục tại nhà.}

\vocabentry{Incentive}
{/ɪnˈsentɪv/ (n)}
{A thing that motivates or encourages someone to do something.}
{The scholarship acts as an incentive for students to study harder.}
{Đây là một danh từ dùng để chỉ sự khuyến khích hoặc động lực.}

\vocabentry{Inclusion}
{/ɪnˈkluːʒn/ (n)}
{The action or state of including or of being included within a group.}
{The policy promotes the inclusion of students with disabilities in mainstream schools.}
{Đây là một danh từ dùng để chỉ sự hòa nhập giáo dục.}

\vocabentry{Institute}
{/ˈɪnstɪtuːt/ (n)}
{An organization having a particular purpose, especially one that is scientific or educational.}
{The Massachusetts Institute of Technology is a world leader in engineering.}
{Đây là một danh từ dùng để chỉ học viện hoặc viện nghiên cứu.}

\vocabentry{Intellectual}
{/ˌɪntəˈlektʃuəl/ (adj)}
{Relating to the intellect.}
{Reading books is an excellent way to stimulate intellectual growth.}
{Đây là một tính từ dùng để chỉ các yếu tố thuộc về trí tuệ hoặc trí óc.}

\vocabentry{Internship}
{/ˈɪntɜːrnʃɪp/ (n)}
{The position of a student or trainee who works in an organization to gain work experience.}
{Doing an internship at a law firm helped her gain practical experience.}
{Đây là một danh từ dùng để chỉ kỳ thực tập.}

\vocabentry{Invigilator}
{/ɪnˈvɪdʒɪleɪtər/ (n)}
{A person whose job is to watch people taking an exam.}
{The invigilator reminded students that they were not allowed to talk during the exam.}
{Đây là một danh từ dùng để chỉ giám thị coi thi.}

\vocabentry{Junior}
{/ˈdʒuːniər/ (adj)}
{Relating to young people or those at a lower level.}
{He is a junior student at the local high school.}
{Đây là một tính từ dùng để chỉ cấp bậc thấp hơn hoặc dành cho trẻ em.}

\vocabentry{Kindergarten}
{/ˈkɪndərɡɑːrtn/ (n)}
{An establishment where children below the age of five are educated.}
{Children learn through play and socialization in kindergarten.}
{Đây là một danh từ dùng để chỉ trường mẫu giáo.}

\vocabentry{Laboratory}
{/ˈlæbrətɔːri/ (n)}
{A room or building equipped for scientific experiments or research.}
{The science class spent the afternoon in the biology laboratory.}
{Đây là một danh từ dùng để chỉ phòng thí nghiệm.}

\vocabentry{Language}
{/ˈlæŋɡwɪdʒ/ (n)}
{The method of human communication.}
{Learning a second language can open up many career opportunities.}
{Đây là một danh từ dùng để chỉ ngôn ngữ.}

\vocabentry{Learner}
{/ˈlɜːrnər/ (n)}
{A person who is learning a subject or skill.}
{Every learner has a different style of acquiring new information.}
{Đây là một danh từ dùng để chỉ người học.}

\vocabentry{Learning}
{/ˈlɜːrnɪŋ/ (n)}
{The acquisition of knowledge or skills through experience or study.}
{Digital technology has transformed the process of learning.}
{Đây là một danh từ dùng để chỉ việc học hoặc kiến thức học được.}

\vocabentry{Lesson}
{/ˈlesn/ (n)}
{A period of learning or teaching.}
{The music lesson was canceled because the teacher was sick.}
{Đây là một danh từ dùng để chỉ bài học hoặc tiết học.}

\vocabentry{Library}
{/ˈlaɪbreri/ (n)}
{A building or room containing collections of books and other materials.}
{I spent the whole morning studying in the university library.}
{Đây là một danh từ dùng để chỉ thư viện.}

\vocabentry{Mastery}
{/ˈmæstəri/ (n)}
{Comprehensive knowledge or skill in a particular subject.}
{It takes years of practice to achieve mastery of a musical instrument.}
{Đây là một danh từ dùng để chỉ sự thành thạo hoặc làm chủ kiến thức.}

\vocabentry{Matriculation}
{/məˌtrɪkjuˈleɪʃn/ (n)}
{Formal admission to a university.}
{Matriculation ceremonies are a long-standing tradition in many older universities.}
{Đây là một danh từ dùng để chỉ việc trúng tuyển và nhập học đại học.}

\vocabentry{Measure}
{/ˈmeʒər/ (v)}
{Ascertain the size, amount, or degree of something.}
{Examinations are used to measure a student's understanding of the course material.}
{Trong giáo dục dùng để chỉ việc đo lường hoặc đánh giá.}

\vocabentry{Methodology}
{/ˌmeθəˈdɑːlədʒi/ (n)}
{A system of methods used in a particular area of study or activity.}
{The research paper explains the methodology used to collect the data.}
{Đây là một danh từ dùng để chỉ hệ phương pháp nghiên cứu hoặc giảng dạy.}

\vocabentry{Motivation}
{/ˌmoʊtɪˈveɪʃn/ (n)}
{The reason or reasons one has for acting or behaving in a particular way.}
{Teachers strive to keep their students' motivation high throughout the term.}
{Đây là một danh từ dùng để chỉ động lực học tập.}

\vocabentry{Multidisciplinary}
{/ˌmʌltidiˈsɪpləneri/ (adj)}
{Combining or involving several academic disciplines.}
{The project uses a multidisciplinary approach to solve urban problems.}
{Đây là một tính từ dùng để chỉ tính đa ngành.}

\vocabentry{Nursery school}
{/ˈnɜːrsəri skuːl/ (n)}
{A school for young children, mainly between the ages of three and five.}
{Children learn important social skills at nursery school.}
{Đây là một danh từ dùng để chỉ trường mầm non.}

\vocabentry{Nurture}
{/ˈnɜːrtʃər/ (v)}
{Care for and encourage the growth or development of.}
{It is important to nurture a child's natural curiosity and creativity.}
{Đây là một động từ dùng để chỉ sự nuôi dưỡng hoặc bồi dưỡng tài năng.}

\vocabentry{Option}
{/ˈɑːpʃn/ (n)}
{A thing that is or may be chosen.}
{Geography is one of the options available for students in their third year.}
{Trong giáo dục thường chỉ môn học tự chọn.}

\vocabentry{Oral}
{/ˈɔːrəl/ (adj)}
{Spoken rather than written.}
{The French exam consists of both a written and an oral section.}
{Đây là một tính từ dùng để chỉ các bài kiểm tra hoặc bài tập vấn đáp.}

\vocabentry{Outline}
{/ˈaʊtlaɪn/ (n)}
{A general description or plan giving the essential features of something.}
{I prepared a brief outline of my essay before I started writing.}
{Đây là một danh từ dùng để chỉ dàn ý hoặc đề cương bài viết.}

\vocabentry{Overcome}
{/ˌoʊvərˈkʌm/ (v)}
{Succeed in dealing with a problem or difficulty.}
{With hard work, he was able to overcome his learning difficulties.}
{Đây là một động từ dùng để chỉ việc vượt qua khó khăn trong học tập.}

\vocabentry{Participation}
{/pɑːrˌtɪsɪˈpeɪʃn/ (n)}
{The action of taking part in something.}
{Class participation is an important part of the final grade.}
{Đây là một danh từ dùng để chỉ sự tham gia đóng góp trong lớp học.}

\vocabentry{Pass}
{/pæs/ (v)}
{Be successful in an examination, test, or course.}
{She was relieved to hear that she had passed her driving test.}
{Đây là một động từ dùng để chỉ việc thi đỗ hoặc vượt qua một kỳ thi.}

\vocabentry{Philosophy}
{/fəˈlɑːsəfi/ (n)}
{The study of the fundamental nature of knowledge; also a set of beliefs.}
{The school's philosophy is based on encouraging independence and creativity.}
{Đây là một danh từ dùng để chỉ triết học hoặc quan điểm giáo dục.}

\vocabentry{Plenitude}
{/ˈplenɪtuːd/ (n)}
{An abundance.}
{There is a plenitude of online resources available for language learners.}
{Trong giáo dục thường chỉ sự phong phú của nguồn tài liệu.}

\vocabentry{Potential}
{/pəˈtenʃl/ (n)}
{Having or showing the capacity to become or develop into something in the future.}
{Teachers strive to help every student reach their full potential.}
{Đây là một danh từ dùng để chỉ tiềm năng.}

\vocabentry{Practical}
{/ˈpræktɪkl/ (adj)}
{Relating to experience or use rather than theory.}
{The course provides students with practical skills that are useful in the workplace.}
{Đây là một tính từ dùng để chỉ tính thực hành hoặc thực tế.}

\vocabentry{Preschool}
{/ˈpriːskuːl/ (n)}
{Relating to the time before a child is old enough to go to primary school.}
{Many children attend preschool to prepare them for formal education.}
{Đây là một danh từ dùng để chỉ giáo dục tiền tiểu học.}

\vocabentry{Principal}
{/ˈprɪnsəpl/ (n)}
{The person with the highest authority in a school.}
{The principal announced several changes to the school's safety policies.}
{Đây là một danh từ dùng để chỉ hiệu trưởng trường học.}

\vocabentry{Private}
{/ˈpraɪvət/ (adj)}
{Not provided or funded by the government.}
{Private schools often have smaller class sizes and better facilities.}
{Đây là một tính từ dùng để chỉ các trường tư thục.}

\vocabentry{Profound}
{/prəˈfaʊnd/ (adj)}
{Very great or intense.}
{The book had a profound impact on my understanding of history.}
{Trong giáo dục thường chỉ sự hiểu biết sâu sắc.}

\vocabentry{Progress}
{/ˈprɑːɡres/ (n/v)}
{Forward movement toward a destination; to develop toward a better condition.}
{The teacher is pleased with the student's progress in mathematics.}
{Đây là một danh từ/động từ dùng để chỉ sự tiến bộ.}

\vocabentry{Pupil}
{/ˈpjuːpl/ (n)}
{A person who is taught by another, especially a schoolchild.}
{The school has over five hundred pupils aged five to eleven.}
{Đây là một danh từ dùng để chỉ học sinh tiểu học.}

\vocabentry{Purpose}
{/ˈpɜːrpəs/ (n)}
{The reason for which something is done or created.}
{The main purpose of the research is to identify the causes of soil erosion.}
{Đây là một danh từ dùng để chỉ mục đích học tập.}

\vocabentry{Quest}
{/kwest/ (n)}
{A long or arduous search for something.}
{His lifelong quest for knowledge took him to many different universities.}
{Trong học thuật chỉ sự tìm kiếm tri thức.}

\vocabentry{Quiz}
{/kwɪz/ (n)}
{A test of knowledge, especially as a competition or game.}
{We have a vocabulary quiz every Friday morning.}
{Đây là một danh từ dùng để chỉ bài kiểm tra ngắn hoặc trò chơi đố vui kiến thức.}

\vocabentry{Range}
{/reɪndʒ/ (n)}
{A set of different things of the same general type.}
{The course covers a wide range of topics in modern history.}
{Dùng để chỉ phạm vi kiến thức hoặc tài liệu.}

\vocabentry{Reflective}
{/rɪˈflektɪv/ (adj)}
{Relating to or characterized by deep thought.}
{Reflective learning helps students understand their own learning processes.}
{Trong giáo dục chỉ phương pháp học tập có sự suy ngẫm, tự đúc kết.}

\vocabentry{Registration}
{/ˌredʒɪˈstreɪʃn/ (n)}
{The action or process of registering or of being registered.}
{Registration for the summer school starts next month.}
{Đây là một danh từ dùng để chỉ việc đăng ký khóa học.}

\vocabentry{Reliability}
{/rɪˌlaɪəˈbɪləti/ (n)}
{The quality of being trustworthy or performing consistently well.}
{The reliability of standardized tests is a subject of much debate.}
{Trong đánh giá giáo dục chỉ độ tin cậy của bài kiểm tra.}

\vocabentry{Requirement}
{/rɪˈkwaɪərmənt/ (n)}
{A thing that is needed or wanted.}
{One of the requirements for the job is a degree in engineering.}
{Đây là một danh từ dùng để chỉ yêu cầu hoặc điều kiện bắt buộc.}

\vocabentry{Resources}
{/rɪˈsɔːrsɪz/ (n)}
{A stock or supply of money, materials, staff, and other assets.}
{The library provides students with a wide variety of research resources.}
{Đây là một danh từ dùng để chỉ nguồn tài liệu học tập.}

\vocabentry{Review}
{/rɪˈvjuː/ (v)}
{A formal assessment of something with the possibility of change.}
{Students should review their notes regularly to prepare for the exam.}
{Đây là một động từ dùng để chỉ việc ôn tập bài học hoặc đánh giá lại.}

\vocabentry{Revision}
{/rɪˈvɪʒn/ (n)}
{The action of revising.}
{Revision is essential for performing well in examinations.}
{Đây là một danh từ dùng để chỉ việc ôn thi hoặc chỉnh sửa văn bản.}

\vocabentry{Scholar}
{/ˈskɑːlər/ (n)}
{A specialist in a particular branch of study.}
{He is a leading scholar in the field of ancient Greek history.}
{Đây là một danh từ dùng để chỉ học giả hoặc người có học thức sâu rộng.}

\vocabentry{Schooling}
{/ˈskuːlɪŋ/ (n)}
{Education received at school.}
{She received most of her primary schooling in a small village.}
{Đây là một danh từ dùng để chỉ việc đi học hoặc sự giáo dục tại trường.}

\vocabentry{Scientific}
{/ˌsaɪənˈtɪfɪk/ (adj)}
{Based on or characterized by the methods and principles of science.}
{The research uses a rigorous scientific methodology to collect data.}
{Đây là một tính từ dùng để chỉ tính khoa học.}

\vocabentry{Secondary school}
{/ˈsekənderi skuːl/ (n)}
{A school for pupils between the ages of 11 and 18.}
{He teaches geography at a local secondary school.}
{Đây là một danh từ dùng để chỉ trường trung học.}

\vocabentry{Self-assessment}
{/self əˈsesmənt/ (n)}
{Evaluation of oneself or one's actions and attitudes.}
{Self-assessment is a useful tool for personal and professional development.}
{Đây là một danh từ dùng để chỉ việc tự đánh giá bản thân.}

\vocabentry{Self-discipline}
{/self ˈdɪsəplɪn/ (n)}
{The ability to control one's feelings and overcome one's weaknesses.}
{Studying for a degree requires a high level of self-discipline.}
{Đây là một danh từ dùng để chỉ tinh thần tự kỷ luật.}

\vocabentry{Self-study}
{/self ˈstʌdi/ (n)}
{Studying by oneself without direct supervision or instruction.}
{The course material is designed for effective self-study.}
{Đây là một danh từ dùng để chỉ việc tự học.}

\vocabentry{Semester}
{/sɪˈmestər/ (n)}
{A half-year term in a school or college.}
{The fall semester begins in September and ends in December.}
{Đây là một danh từ dùng để chỉ học kỳ.}

\vocabentry{Senior}
{/ˈsiːniər/ (adj)}
{Relating to old people or those at a higher level.}
{She is a senior lecturer in the department of sociology.}
{Trong giáo dục chỉ học sinh năm cuối hoặc cấp cao.}

\vocabentry{Skill}
{/skɪl/ (n)}
{The ability to do something well.}
{Communication is an essential skill for success in any career.}
{Đây là một danh từ dùng để chỉ kỹ năng.}

\vocabentry{Socialization}
{/ˌsoʊʃələˈzeɪʃn/ (n)}
{The process of learning to behave in a way that is acceptable to society.}
{Schools play a crucial role in the socialization of young children.}
{Trong giáo dục chỉ quá trình hòa nhập xã hội.}

\vocabentry{Society}
{/səˈsaɪəti/ (n)}
{The aggregate of people living together in a more or less ordered community.}
{Education is essential for the development and progress of society.}
{Đây là một danh từ dùng để chỉ xã hội.}

\vocabentry{Special needs}
{/ˌspeʃl ˈniːdz/ (n)}
{Particular educational requirements resulting from learning difficulties or disabilities.}
{The school provides extra support for students with special needs.}
{Đây là một danh từ dùng để chỉ các nhu cầu giáo dục đặc biệt.}

\vocabentry{Student}
{/ˈstuːdnt/ (n)}
{A person who is studying at a school or college.}
{Thousands of international students come to London every year.}
{Đây là một danh từ dùng để chỉ sinh viên hoặc học sinh.}

\vocabentry{Study}
{/ˈstʌdi/ (n/v)}
{The devotion of time and attention to acquiring knowledge; to devote time to acquiring knowledge.}
{She is conducting a study on the effects of social media on mental health.}
{Đây là một danh từ/động từ dùng để chỉ việc học tập hoặc nghiên cứu.}

\vocabentry{Subject}
{/ˈsʌbdʒɪkt/ (n)}
{A branch of knowledge studied or taught in a school.}
{Mathematics was always my favorite subject at school.}
{Đây là một danh từ dùng để chỉ môn học.}

\vocabentry{Supervisor}
{/ˈsuːpərvaɪzər/ (n)}
{A person who supervises a person or an activity.}
{I meet with my research supervisor once a week to discuss my progress.}
{Trong giáo dục chỉ giảng viên hướng dẫn luận văn.}

\vocabentry{Supplementary}
{/ˌsʌplɪˈmentri/ (adj)}
{Completing or enhancing something.}
{The teacher provided some supplementary material for students who were struggling.}
{Trong giáo dục chỉ tài liệu hoặc khóa học bổ trợ.}

\vocabentry{Support}
{/səˈpɔːrt/ (n/v)}
{Give assistance to; assistance provided.}
{The university offers extra support for students who have learning difficulties.}
{Đây là một danh từ/động từ dùng để chỉ sự hỗ trợ học tập.}

\vocabentry{Suspending}
{/səˈspendɪŋ/ (v)}
{Temporarily prevent from continuing or being in force or effect.}
{The student was suspended for two weeks for violating the school's rules.}
{Trong giáo dục chỉ việc đình chỉ học.}

\vocabentry{Talent}
{/ˈtælənt/ (n)}
{Natural aptitude or skill.}
{The music school provides a platform for young talents to perform.}
{Đây là một danh từ dùng để chỉ tài năng.}

\vocabentry{Target}
{/ˈtɑːrɡɪt/ (n)}
{A person, object, or place selected as the aim of an attack.}
{The government has set a target of reducing school dropout rates by 10\%.}
{Trong giáo dục chỉ mục tiêu cần đạt được.}

\vocabentry{Teacher}
{/ˈtiːtʃər/ (n)}
{A person who teaches, especially in a school.}
{A good teacher can inspire students to reach their full potential.}
{Đây là một danh từ dùng để chỉ giáo viên.}

\vocabentry{Teaching}
{/ˈtiːtʃɪŋ/ (n)}
{The ideas or principles taught by an authority.}
{Teaching is a rewarding but challenging profession.}
{Đây là một danh từ dùng để chỉ sự giảng dạy hoặc học thuyết.}

\vocabentry{Teamwork}
{/ˈtiːmwɜːrk/ (n)}
{The combined action of a group of people.}
{Group projects are a great way to develop teamwork skills.}
{Đây là một danh từ dùng để chỉ tinh thần làm việc nhóm.}

\vocabentry{Technique}
{/tekˈniːk/ (n)}
{A way of carrying out a particular task.}
{She is trying to improve her painting technique by attending a weekly class.}
{Đây là một danh từ dùng để chỉ kỹ thuật hoặc phương pháp làm việc.}

\vocabentry{Technology}
{/tekˈnɑːlədʒi/ (n)}
{The application of scientific knowledge for practical purposes.}
{Educational technology has made learning more interactive and accessible.}
{Trong giáo dục chỉ công nghệ hỗ trợ giảng dạy.}

\vocabentry{Term}
{/tɜːrm/ (n)}
{A fixed or limited period for which something lasts.}
{The spring term ends in late May, just before the summer holidays.}
{Đây là một danh từ dùng để chỉ học kỳ hoặc thuật ngữ.}

\vocabentry{Test}
{/test/ (n/v)}
{A procedure intended to establish the quality, performance, or reliability of something.}
{We have a math test every Friday afternoon to check our progress.}
{Đây là một danh từ/động từ dùng để chỉ bài kiểm tra.}

\vocabentry{Textbook}
{/ˈtekstbʊk/ (n)}
{A book used as a standard work for the study of a particular subject.}
{You need to buy several new textbooks for the sociology course.}
{Đây là một danh từ dùng để chỉ sách giáo khoa.}

\vocabentry{Theory}
{/ˈθɪəri/ (n)}
{A system of ideas intended to explain something.}
{The course covers both educational theory and practical teaching methods.}
{Đây là một danh từ dùng để chỉ lý thuyết.}

\vocabentry{Thinking}
{/ˈθɪŋkɪŋ/ (n)}
{The process of using one's mind to consider or reason about something.}
{Critical thinking is an essential skill for any university student.}
{Đây là một danh từ dùng để chỉ tư duy hoặc suy nghĩ.}

\vocabentry{Topic}
{/ˈtɑːpɪk/ (n)}
{A matter dealt with in a text, discourse, or conversation.}
{The teacher chose a very interesting topic for the class discussion.}
{Đây là một danh từ dùng để chỉ chủ đề bài học.}

\vocabentry{Trainee}
{/ˌtreɪˈniː/ (n)}
{A person undergoing training for a particular job or profession.}
{The hospital has several trainee doctors working in the emergency department.}
{Đây là một danh từ dùng để chỉ người đang thực tập hoặc được đào tạo nghề.}

\vocabentry{Training}
{/ˈtreɪnɪŋ/ (n)}
{The action of teaching a person or animal a particular skill or type of behavior.}
{All new employees receive thorough training before they start their jobs.}
{Đây là một danh từ dùng để chỉ hoạt động đào tạo.}

\vocabentry{Tutor}
{/ˈtuːtər/ (n/v)}
{A private teacher, typically one who teaches a single pupil or a very small group.}
{I hired a math tutor to help me prepare for my final exams.}
{Đây là một danh từ/động từ dùng để chỉ gia sư hoặc việc dạy kèm.}

\vocabentry{Underpin}
{/ˌʌndərˈpɪn/ (v)}
{Support, justify, or form the basis for.}
{The theory of evolution underpins much of modern biology.}
{Trong giáo dục chỉ việc củng cố kiến thức hoặc làm cơ sở cho lý thuyết.}

\vocabentry{Understand}
{/ˌʌndərˈstænd/ (v)}
{Perceive the intended meaning of.}
{I didn't quite understand the teacher's explanation of the formula.}
{Đây là một động từ dùng để chỉ việc hiểu hoặc nắm bắt kiến thức.}

\vocabentry{Uniform}
{/ˈjuːnɪfɔːrm/ (n)}
{The distinctive clothing worn by members of the same organization.}
{All students are required to wear the school uniform during term time.}
{Đây là một danh từ dùng để chỉ đồng phục trường học.}

\vocabentry{University}
{/ˌjuːnɪˈvɜːrsəti/ (n)}
{An educational institution designed for instruction, examination, or both, of students in many branches of advanced learning.}
{He plans to attend university to study architecture.}
{Đây là một danh từ dùng để chỉ trường đại học.}

\vocabentry{Updating}
{/ˌʌpˈdeɪtɪŋ/ (v)}
{Make something more modern or up to date.}
{Doctors must keep updating their knowledge of new medical treatments.}
{Đây là một động từ dùng để chỉ việc cập nhật kiến thức.}

\vocabentry{Values}
{/ˈvæljuːz/ (n)}
{A person's principles or standards of behavior.}
{Schools play an important role in teaching students social and moral values.}
{Trong giáo dục chỉ các giá trị đạo đức được truyền thụ.}

\vocabentry{Variety}
{/vəˈraɪəti/ (n)}
{The quality or state of being different or diverse.}
{The course offers a wide variety of elective subjects to choose from.}
{Đây là một danh từ dùng để chỉ sự đa dạng trong phương pháp hoặc tài liệu.}

\vocabentry{Versatile}
{/ˈvɜːrsətl/ (adj)}
{Able to adapt or be adapted to many different functions or activities.}
{Education aims to produce versatile individuals who can thrive in any career.}
{Trong giáo dục chỉ người đa tài hoặc phương pháp linh hoạt.}

\vocabentry{Virtual}
{/ˈvɜːrtʃuəl/ (adj)}
{Not physically existing as such but made by software to appear to do so.}
{Virtual classrooms have transformed the way we learn.}
{Đây là một tính từ dùng để chỉ các yếu tố ảo hoặc trực tuyến.}

\vocabentry{Vision}
{/ˈvɪʒn/ (n)}
{The ability to think about or plan the future with imagination or wisdom.}
{The university's vision is to become a world leader in medical research.}
{Trong giáo dục chỉ tầm nhìn chiến lược của một tổ chức.}

\vocabentry{Vital}
{/ˈvaɪtl/ (adj)}
{Absolutely necessary or important.}
{Education is vital for the development and progress of any nation.}
{Đây là một tính từ dùng để chỉ tính chất thiết yếu.}

\vocabentry{Vocabulary}
{/vəˈkæbjəleri/ (n)}
{The body of words used in a particular language.}
{Reading books is an excellent way to expand your vocabulary.}
{Đây là một danh từ dùng để chỉ từ vựng.}

\vocabentry{Vocation}
{/voʊˈkeɪʃn/ (n)}
{A strong feeling of suitability for a particular career or occupation.}
{Teaching is not just a job for her; it is a true vocation.}
{Đây là một danh từ dùng để chỉ thiên hướng nghề nghiệp hoặc công việc như một sứ mệnh.}

\vocabentry{Volume}
{/ˈvɑːljuːm/ (n)}
{A book forming part of a larger work or series.}
{The encyclopedia consists of twenty-four volumes covering various subjects.}
{Đây là một danh từ dùng để chỉ tập sách hoặc khối lượng tài liệu.}

\vocabentry{Well-being}
{/ˌwel ˈbiːɪŋ/ (n)}
{The state of being comfortable, healthy, or happy.}
{Schools should focus on the well-being of their students, not just their grades.}
{Trong giáo dục chỉ sự hạnh phúc và sức khỏe của học sinh.}

\vocabentry{Wisdom}
{/ˈwɪzdəm/ (n)}
{The quality of having experience, knowledge, and good judgment.}
{Older people often possess a great deal of wisdom and life experience.}
{Đây là một danh từ dùng để chỉ sự thông thái hoặc trí khôn.}

\vocabentry{Writing}
{/ˈraɪtɪŋ/ (n)}
{The activity or skill of marking coherent words on paper or a computer.}
{The course aims to improve students' academic writing skills.}
{Đây là một danh từ dùng để chỉ kỹ năng viết.}

\vocabentry{Youth}
{/juːθ/ (n)}
{The period between childhood and adult age.}
{Education is essential for empowering the youth to build a better future.}
{Đây là một danh từ dùng để chỉ tuổi trẻ hoặc thanh thiếu niên.}

\vocabentry{Zeal}
{/ziːl/ (n)}
{Great energy or enthusiasm in pursuit of a cause or an objective.}
{Her zeal for learning foreign languages is truly inspiring.}
{Đây là một danh từ dùng để chỉ lòng nhiệt huyết trong học tập.}

\vocabentry{Zone}
{/zoʊn/ (n)}
{An area or stretch of land having a particular characteristic.}
{The library has a designated quiet zone for students who need to focus.}
{Trong giáo dục chỉ một khu vực học tập hoặc phạm vi kiến thức.}

