% Auto-generated from raw.txt
% Usage:
%   \vocabentry{word}{/ipa/ (pos)}{definition}{example}{note}

\vocabentry{Itinerary}
{/aɪˈtɪnəreri/ (n)}
{A planned route or journey.}
{We had to change our itinerary because of the bad weather.}
{}

\vocabentry{Ecotourism}
{/ˈiːkoʊtʊrɪzəm/ (n)}
{Tourism directed toward exotic, often threatened, natural environments, intended to support conservation efforts.}
{Ecotourism helps to protect the rainforest by providing a source of income for local communities that doesn't involve logging.}
{Đây là một danh từ dùng để chỉ du lịch sinh thái.}

\vocabentry{Accommodation}
{/əˌkɑːməˈdeɪʃn/ (n)}
{A room, group of rooms, or building in which someone may live or stay.}
{Finding affordable accommodation in the city center during peak season is almost impossible.}
{Đây là một danh từ dùng để chỉ chỗ ở, nơi lưu trú cho khách du lịch.}

\vocabentry{Destination}
{/ˌdestɪˈneɪʃn/ (n)}
{The place to which someone or something is going or being sent.}
{What is your final destination?}
{}

\vocabentry{Hospitality}
{/ˌhɑːspɪˈtæləti/ (n)}
{Friendly reception of guests.}
{The local people are famous for their hospitality and generosity.}
{}

\vocabentry{Sightseeing}
{/ˈsaɪtsiːɪŋ/ (n)}
{The activity of visiting places of interest in a particular location.}
{We spent the entire day sightseeing in London, visiting the Tower of London and Big Ben.}
{Đây là một danh từ dùng để chỉ hoạt động tham quan các danh lam thắng cảnh.}

\vocabentry{Breathtaking}
{/ˈbreθteɪkɪŋ/ (adj)}
{Astonishing or awe-inspiring in quality, so as to take one's breath away.}
{The view from the top of the mountain was absolutely breathtaking.}
{Đây là một tính từ dùng để chỉ vẻ đẹp đến ngỡ ngàng, ngoạn mục.}

\vocabentry{Expedition}
{/ˌekspəˈdɪʃn/ (n)}
{A journey or voyage undertaken by a group of people with a particular purpose.}
{He joined an expedition to search for the lost city in the jungle.}
{}

\vocabentry{Inbound}
{/ˈɪnbaʊnd/ (adj)}
{Traveling or transport heading toward a particular place.}
{The government is launching a campaign to boost inbound tourism from European markets.}
{Đây là một tính từ dùng để chỉ khách du lịch từ nước ngoài vào một quốc gia.}

\vocabentry{Outbound}
{/ˈaʊtbaʊnd/ (adj)}
{Traveling or transport heading away from a particular place.}
{Outbound travel has increased significantly as people have more disposable income.}
{Đây là một tính từ dùng để chỉ khách du lịch đi từ trong nước ra nước ngoài.}

\vocabentry{Excursion}
{/ɪkˈskɜːrʒn/ (n)}
{A short journey or trip.}
{The hotel organizes daily excursions to the local sights.}
{}

\vocabentry{Souvenir}
{/ˌsuːvəˈnɪr/ (n)}
{A thing that is kept as a reminder of a person, place, or event.}
{I bought a small model of the Eiffel Tower as a souvenir of my trip to Paris.}
{Đây là một danh từ dùng để chỉ quà lưu niệm.}

\vocabentry{Immersive}
{/ɪˈmɜːrsɪv/ (adj)}
{Providing information or entertainment in a way that makes the person feel completely involved.}
{Staying with a local family provides a more immersive cultural experience than staying in a hotel.}
{Đây là một tính từ dùng để chỉ trải nghiệm mang tính đắm chìm, đi sâu vào thực tế.}

\vocabentry{Heritage}
{/ˈherɪtɪdʒ/ (n)}
{Valued objects and qualities such as cultural traditions and historic buildings passed down from previous generations.}
{This building is part of our national heritage and must be protected.}
{Đây là một danh từ dùng để chỉ di sản văn hóa hoặc truyền thống.}

\vocabentry{Jet lag}
{/ˈdʒet læɡ/ (n)}
{Extreme tiredness and other physical effects felt by a person after a long flight across different time zones.}
{It took me several days to recover from the jet lag after flying from New York to Tokyo.}
{Đây là một cụm danh từ dùng để chỉ sự mệt mỏi do lệch múi giờ sau chuyến bay dài.}

\vocabentry{Stopover}
{/ˈstɑːpoʊvər/ (n)}
{A break in a journey.}
{We had a two-day stopover in Singapore on our way to Australia.}
{Đây là một danh từ dùng để chỉ điểm dừng chân tạm thời trong một chuyến bay dài hoặc hành trình dài.}

\vocabentry{Picturesque}
{/ˌpɪktʃəˈresk/ (adj)}
{Visually attractive, especially in a quaint or charming way.}
{We stayed in a picturesque village in the heart of the English countryside.}
{Đây là một tính từ dùng để chỉ vẻ đẹp cổ kính, đẹp như tranh vẽ.}

\vocabentry{Globetrotter}
{/ˈɡloʊbtrɑːtər/ (n)}
{A person who travels widely.}
{As a professional photographer, he is a real globetrotter, traveling to dozens of countries every year.}
{Đây là một danh từ dùng để chỉ người đi du lịch khắp thế giới.}

\vocabentry{Embark}
{/ɪmˈbɑːrk/ (v)}
{Go on board a ship, aircraft, or other vehicle.}
{We are about to embark on a cruise around the Mediterranean.}
{Đây là một động từ dùng để chỉ việc bắt đầu một chuyến hành trình.}

\vocabentry{Sustainable}
{/səˈsteɪnəbl/ (adj)}
{Able to be maintained at a certain rate or level. (Đây là một tính từ dùng để chỉ tính bền vững (trong du lịch bền vững).).}
{Sustainable tourism aims to minimize the negative impact on the environment and local culture.}
{}

\vocabentry{Amenities}
{/əˈmenəti/ (n)}
{Desirable or useful features of a building or place.}
{The hotel offers a wide range of amenities, including a spa, a fitness center, and a swimming pool.}
{Đây là một danh từ dùng để chỉ các tiện nghi công cộng hoặc tiện ích tại khách sạn.}

\vocabentry{All-inclusive}
{/ˌɔːl ɪnˈkluːsɪv/ (adj)}
{Including everything, especially all the costs of a holiday.}
{We booked an all-inclusive holiday so we wouldn't have to worry about the cost of meals and drinks.}
{Đây là một tính từ dùng để chỉ loại hình du lịch trọn gói mọi chi phí.}

\vocabentry{High season}
{/ˈhaɪ ˌsiːzn/ (n)}
{The time of year when the most people visit a place.}
{Prices for flights and hotels are much higher during the high season.}
{Đây là một cụm danh từ dùng để chỉ mùa cao điểm du lịch.}

\vocabentry{Off the beaten track}
{/ɔːf ðə ˈbiːtn træk/ (idiom)}
{In a place where few people go, far from any main roads and towns.}
{We like to go off the beaten track to discover hidden gems that most tourists miss.}
{Đây là một thành ngữ dùng để chỉ những nơi hẻo lánh, ít người biết đến.}

\vocabentry{Backpack}
{/ˈbækpæk/ (v)}
{Travel or hike carrying one's belongings in a backpack.}
{They spent six months backpacking across Southeast Asia after they finished university.}
{Đây là một động từ dùng để chỉ việc đi du lịch bụi.}

\vocabentry{Voyage}
{/ˈvɔɪɪdʒ/ (n)}
{A long journey involving travel by sea or in space.}
{The Titanic sank on its maiden voyage across the Atlantic.}
{Đây là một danh từ dùng để chỉ một chuyến đi dài bằng đường biển hoặc đường hàng không.}

\vocabentry{Transit}
{/ˈtrænzɪt/ (n)}
{The carrying of people, goods, or materials from one place to another.}
{Passengers in transit do not need a visa to stay in the airport's international zone.}
{Đây là một danh từ dùng để chỉ sự quá cảnh hoặc vận chuyển.}

\vocabentry{Charter}
{/ˈtʃɑːrtər/ (v)}
{Rent a ship, aircraft, or bus for a special use.}
{The group chartered a small plane to take them to the remote island.}
{Đây là một động từ dùng để chỉ việc thuê riêng một phương tiện cho mục đích cụ thể.}

\vocabentry{Disembark}
{/ˌdɪsɪmˈbɑːrk/ (v)}
{Leave a ship, aircraft, or other vehicle.}
{Passengers were asked to disembark from the plane through the rear exit.}
{Đây là một động từ dùng để chỉ việc xuống tàu, máy bay hoặc xe.}

\vocabentry{Hospitality industry}
{/ˌhɑːspɪˈtæləti ˈɪndəstri/ (n)}
{The business of providing food, drink, and accommodation for customers.}
{The hospitality industry has been hit hard by the travel restrictions.}
{Đây là một cụm danh từ dùng để chỉ ngành dịch vụ nhà hàng, khách sạn.}

\vocabentry{Remote}
{/rɪˈmoʊt/ (adj)}
{(Of a place) situated far from the main centers of population; distant.}
{The village is located in a remote part of the mountains, far from the nearest city.}
{Đây là một tính từ dùng để chỉ những nơi xa xôi, hẻo lánh.}

\vocabentry{Landmark}
{/ˈlændmɑːrk/ (n)}
{An object or feature of a landscape or town that is easily seen and recognized from a distance.}
{The Statue of Liberty is a famous landmark in New York City.}
{Đây là một danh từ dùng để chỉ địa danh, công trình biểu tượng.}

\vocabentry{Culture shock}
{/ˈkʌltʃər ʃɑːk/ (n)}
{The feeling of disorientation experienced by someone who is suddenly subjected to an unfamiliar culture.}
{Living in a foreign country can be a great experience, but many people suffer from culture shock at first.}
{Đây là một cụm danh từ dùng để chỉ cú sốc văn hóa.}

\vocabentry{Domestic}
{/dəˈmestɪk/ (adj)}
{Relating to or occurring within a particular country; not foreign or international.}
{Domestic tourism has seen a boost as people choose to explore their own country instead of traveling abroad.}
{Đây là một tính từ dùng để chỉ du lịch nội địa.}

\vocabentry{Exotic}
{/ɪɡˈzɑːtɪk/ (adj)}
{Originating in or characteristic of a distant foreign country.}
{I love trying exotic fruits when I travel to tropical countries.}
{Đây là một tính từ dùng để chỉ vẻ đẹp kỳ lạ, ngoại lai.}

\vocabentry{Infrastructure}
{/ˈɪnfrəstrʌktʃər/ (n)}
{The basic physical and organizational structures and facilities needed for the operation of a society.}
{Developing the tourism infrastructure is essential for attracting more international visitors.}
{Trong du lịch, đây là cơ sở hạ tầng phục vụ du khách.}

\vocabentry{Lush}
{/lʌʃ/ (adj)}
{(Of vegetation) growing luxuriantly.}
{The valley was covered in lush green forests and sparkling waterfalls.}
{Đây là một tính từ dùng để chỉ thảm thực vật tươi tốt, sum suê.}

\vocabentry{Over tourism}
{/ˌoʊvər ˈtʊrɪzəm/ (n)}
{A situation in which too many tourists visit a place, leading to a negative impact on the environment and local quality of life.}
{Many historic cities are struggling to deal with the effects of over tourism.}
{Đây là một cụm danh từ dùng để chỉ tình trạng quá tải du lịch.}

\vocabentry{Recreation}
{/ˌrekriˈeɪʃn/ (n)}
{Activity done for enjoyment when one is not working.}
{The park provides a variety of facilities for recreation, including hiking trails and picnic areas.}
{Đây là một danh từ dùng để chỉ hoạt động giải trí, vui chơi.}

\vocabentry{Trek}
{/trek/ (v)}
{Go on a long arduous journey, typically on foot.}
{We trekked through the jungle for three days to reach the ancient ruins.}
{Đây là một động từ dùng để chỉ việc đi bộ đường dài, leo núi thám hiểm.}

\vocabentry{Wanderlust}
{/ˈwɑːndərlʌst/ (n)}
{A strong desire to travel.}
{His wanderlust has taken him to almost every continent on Earth.}
{Đây là một danh từ dùng để chỉ niềm khao khát mãnh liệt được đi du lịch.}

\vocabentry{Secluded}
{/rɪˈkluːdɪd/ (adj)}
{(Of a place) not seen or visited by many people; sheltered and private.}
{We found a secluded beach where we could relax without being disturbed by other tourists.}
{Đây là một tính từ dùng để chỉ những nơi yên tĩnh, tách biệt.}

\vocabentry{Visa}
{/ˈviːzə/ (n)}
{An endorsement on a passport indicating that the holder is allowed to enter, leave, or stay for a specified period of time in a country.}
{You need to apply for a visa at least a month before your trip.}
{Đây là một danh từ dùng để chỉ thị thực nhập cảnh.}

\vocabentry{Boarding pass}
{/ˈbɔːrdɪŋ pæs/ (n)}
{A pass for boarding an aircraft, given to a passenger when the check-in process is done.}
{Please have your passport and boarding pass ready for inspection.}
{Đây là một cụm danh từ dùng để chỉ thẻ lên máy bay.}

\vocabentry{Customs}
{/ˈkʌstəmz/ (n)}
{The official department that administers and collects the duties levied by a government on imported goods.}
{We had to go through customs after arriving at the international airport.}
{Đây là một danh từ dùng để chỉ cơ quan hải quan.}

\vocabentry{Travel insurance}
{/ˈtrævl ɪnˌʃʊrəns/ (n)}
{Insurance that is intended to cover medical expenses, trip cancellation, lost luggage, and other losses incurred while traveling.}
{It is always a good idea to take out travel insurance before going abroad.}
{Đây là một cụm danh từ dùng để chỉ bảo hiểm du lịch.}

\vocabentry{Guided tour}
{/ˌɡaɪdɪd ˈtʊr/ (n)}
{A tour of a historical site or place of interest with a guide who provides information.}
{We took a guided tour of the Vatican Museums to learn more about the art and history.}
{Đây là một cụm danh từ dùng để chỉ chuyến tham quan có hướng dẫn viên.}

\vocabentry{Peak season}
{/ˈpiːk ˌsiːzn/ (n)}
{The time of year when the most people visit a place.}
{Hotel rates are much higher during the peak season in the summer.}
{Tương đương với High Season - mùa cao điểm.}

\vocabentry{Low season}
{/ˈloʊ ˌsiːzn/ (n)}
{The time of year when the fewest people visit a place.}
{You can find great deals on flights and accommodation during the low season.}
{Đây là một cụm danh từ dùng để chỉ mùa thấp điểm du lịch.}

\vocabentry{Traveler's cheque}
{/ˈtrævələrz tʃek/ (n)}
{A cheque for a fixed amount that can be cashed or used in payment after endorsement by the holder's signature.}
{Traveler's cheques are less common now that ATMs are widely available.}
{Đây là một cụm danh từ dùng để chỉ séc du lịch.}

\vocabentry{Valise}
{/vəˈliːs/ (n)}
{A small traveling bag or suitcase.}
{He carried his essential items in a small leather valise.}
{Đây là một danh từ dùng để chỉ vali hoặc túi xách nhỏ đi du lịch.}

\vocabentry{Wayfarer}
{/ˈweɪferər/ (n)}
{A person who travels on foot.}
{The weary wayfarer stopped at an inn for a meal and a night's rest.}
{Đây là một danh từ dùng để chỉ người lữ hành, đặc biệt là người đi bộ.}

\vocabentry{Voyager}
{/ˈvɔɪɪdʒər/ (n)}
{A person who goes on a long journey, typically by sea.}
{The early voyagers used the stars to navigate across the vast oceans.}
{Đây là một danh từ dùng để chỉ người đi du hành xa.}

\vocabentry{Transcontinental}
{/ˌtrænzkɑːntɪˈnentl/ (adj)}
{(Of a journey, etc.) crossing a continent.}
{The transcontinental railway revolutionized travel across the United States.}
{Đây là một tính từ dùng để chỉ chuyến đi xuyên lục địa.}

\vocabentry{Terminal}
{/ˈtɜːrmɪnl/ (n)}
{A building at an airport where passengers transfer between ground transportation and the facilities that allow them to board and disembark from aircraft.}
{We met our friends at the arrivals terminal.}
{Đây là một danh từ dùng để chỉ nhà ga sân bay.}

\vocabentry{Souvenir shop}
{/suːvəˈnɪr ʃɑːp/ (n)}
{A shop that sells items intended to be kept as a reminder of a place.}
{The airport is full of souvenir shops selling local crafts and food.}
{Đây là một cụm danh từ dùng để chỉ cửa hàng lưu niệm.}

\vocabentry{Self-catering}
{/ˌself ˈkeɪtərɪŋ/ (adj)}
{(Of a holiday or accommodation) offering facilities for you to cook your own meals.}
{We stayed in a self-catering apartment so we could save money by cooking our own meals.}
{Đây là một tính từ dùng để chỉ loại hình du lịch tự phục vụ ăn uống.}

\vocabentry{Resort}
{/rɪˈzɔːrt/ (n)}
{A place that is a popular destination for vacations or recreation.}
{The seaside resort was packed with families enjoying the summer sun.}
{Đây là một danh từ dùng để chỉ khu nghỉ dưỡng.}

\vocabentry{Pilgrimage}
{/ˈpɪlɡrɪmɪdʒ/ (n)}
{A journey to a place associated with someone or something well known or respected.}
{Thousands of people make a pilgrimage to the holy site every year.}
{Đây là một danh từ dùng để chỉ chuyến hành hương.}

\vocabentry{Passage}
{/ˈpæsɪdʒ/ (n)}
{A journey by sea or air.}
{The passage across the Atlantic took ten days by ship.}
{Đây là một danh từ dùng để chỉ chuyến đi hoặc sự đi qua.}

\vocabentry{Overnight}
{/ˌoʊvərˈnaɪt/ (adv)}
{During or for the night.}
{We decided to stay overnight in a small hotel near the station.}
{Đây là một trạng từ dùng để chỉ hành động diễn ra qua đêm.}

\vocabentry{Outpost}
{/ˈaʊtpoʊst/ (n)}
{A small military camp or settlement in a remote area.}
{The scientific outpost is located in a remote part of the Arctic.}
{Đây là một danh từ dùng để chỉ trạm tiền tiêu hoặc khu định cư xa xôi.}

\vocabentry{Mainland}
{/ˈmeɪnlænd/ (n)}
{A large continuous extent of land that includes the greater part of a country or region.}
{We took a ferry from the island back to the mainland.}
{Đây là một danh từ dùng để chỉ đất liền.}

\vocabentry{Luggage}
{/ˈlʌɡɪdʒ/ (n)}
{Suitcases or other bags in which to pack personal belongings for traveling.}
{Please do not leave your luggage unattended at any time.}
{Đây là một danh từ dùng để chỉ hành lý.}

\vocabentry{Inn}
{/ɪn/ (n)}
{A house providing accommodation for travelers.}
{We stayed in a traditional inn that has been in the same family for generations.}
{}

\vocabentry{Hostel}
{/ˈhɑːstl/ (n)}
{Inexpensive food and lodging for travelers.}
{Staying in a hostel is a great way to save money and meet new people.}
{}

\vocabentry{Hitchhike}
{/ˈhɪtʃhaɪk/ (v)}
{Travel by getting free rides from passing vehicles.}
{They decided to hitchhike across Europe to save money on transportation.}
{Đây là một động từ dùng để chỉ việc đi nhờ xe.}

\vocabentry{Guest house}
{/ˈɡest haʊs/ (n)}
{A private house offering accommodation to paying guests.}
{We stayed in a charming guest house near the beach.}
{Đây là một cụm danh từ dùng để chỉ nhà khách.}

\vocabentry{Fare}
{/fer/ (n)}
{The money a passenger on public transport has to pay.}
{Train fares have increased significantly over the last few years.}
{Đây là một danh từ dùng để chỉ giá vé tàu, xe, máy bay.}

\vocabentry{Curfew}
{/ˈkɜːrfjuː/ (n)}
{A regulation requiring people to remain indoors between specified hours.}
{The local government imposed a curfew to ensure the safety of residents and tourists.}
{Thường dùng trong các khu vực an ninh nghiêm ngặt khi đi du lịch.}

\vocabentry{Cuisine}
{/kwɪˈziːn/ (n)}
{A style or method of cooking, especially as characteristic of a particular country, region, or establishment.}
{The local cuisine is famous for its use of fresh seafood and spices.}
{Đây là một danh từ dùng để chỉ ẩm thực.}

\vocabentry{Cruise}
{/kruːz/ (n)}
{A voyage on a ship or boat taken for pleasure or as a holiday.}
{We are going on a ten-day cruise around the Caribbean.}
{Đây là một danh từ dùng để chỉ chuyến du ngoạn bằng tàu thủy.}

\vocabentry{Check-in}
{/ˈtʃek ɪn/ (n)}
{The process of announcing one's arrival at an airport, hotel, etc.}
{Check-in starts two hours before the flight's scheduled departure time.}
{Đây là một danh từ dùng để chỉ việc làm thủ tục nhập cảnh/nhận phòng.}

\vocabentry{Chalet}
{/ʃæˈleɪ/ (n)}
{A wooden house or cottage with overhanging eaves, typical of the Swiss Alps.}
{We rented a cozy chalet for our skiing holiday in the mountains.}
{Đây là một danh từ dùng để chỉ biệt thự nhỏ bằng gỗ, thường ở vùng núi.}

\vocabentry{Bungalow}
{/ˈbʌŋɡəloʊ/ (n)}
{A low house having only one story.}
{We stayed in a beautiful bungalow right on the beach.}
{Đây là một danh từ dùng để chỉ nhà một tầng, thường thấy ở các khu nghỉ dưỡng ven biển.}

\vocabentry{Backpacking}
{/ˈbækpækɪŋ/ (n)}
{The activity of traveling or hiking carrying one's belongings in a backpack.}
{Backpacking is a popular way for young people to see the world on a budget.}
{Đây là một danh từ dùng để chỉ hành trình du lịch bụi.}

\vocabentry{Arrival}
{/əˈraɪvl/ (n)}
{The action or process of arriving.}
{Please check the arrival time of your flight before heading to the airport.}
{Đây là một danh từ dùng để chỉ sự đến nơi.}

\vocabentry{Departure}
{/dɪˈpɑːrtʃər/ (n)}
{The action of leaving, especially to start a journey.}
{The flight's departure was delayed by two hours due to bad weather.}
{Đây là một danh từ dùng để chỉ sự khởi hành.}

\vocabentry{Agritourism}
{/ˌæɡroʊˈtʊrɪzəm/ (n)}
{Tourism in which tourists stay with local people in rural areas abroad.}
{Agritourism is a growing trend as people seek more authentic travel experiences.}
{Đây là một danh từ dùng để chỉ du lịch nông nghiệp.}

\vocabentry{Adventure}
{/ədˈventʃər/ (n)}
{An unusual and exciting, typically hazardous, experience or activity.}
{Climbing Mount Everest was the greatest adventure of his life.}
{Đây là một danh từ dùng để chỉ chuyến phiêu lưu.}

\vocabentry{Aloft}
{/əˈlɔːft/ (adv)}
{Up in the air; overhead.}
{We spent ten hours aloft on our flight to London.}
{Thường dùng để mô tả trạng thái khi ở trên máy bay.}

\vocabentry{Artifact}
{/ˈɑːrtɪfækt/ (n)}
{An object made by a human being, typically one of cultural or historical interest.}
{The museum contains many ancient artifacts discovered in the local area.}
{Đây là một danh từ dùng để chỉ hiện vật văn hóa.}

\vocabentry{Backcountry}
{/ˈbækˌkʌntri/ (n)}
{A remote undeveloped rural area.}
{We spent a week hiking in the backcountry, far from any roads or towns.}
{Đây là một danh từ dùng để chỉ vùng sâu vùng xa, hẻo lánh.}

\vocabentry{Board}
{/bɔːrd/ (v)}
{Get on or into (a ship, aircraft, or other vehicle).}
{Passengers were asked to board the plane in order of their seat numbers.}
{Đây là một động từ dùng để chỉ việc lên tàu/máy bay.}

\vocabentry{Book}
{/bʊk/ (v)}
{Reserve (accommodation, a ticket, etc.) in advance.}
{You should book your hotel well in advance if you are traveling during the peak season.}
{Đây là một động từ dùng để chỉ việc đặt trước.}

\vocabentry{Botany}
{/ˈbɑːtəni/ (n)}
{The scientific study of plants.}
{The botanical garden is a popular destination for tourists interested in plants.}
{Liên quan đến du lịch nghiên cứu thực vật.}

\vocabentry{Budget}
{/ˈbʌdʒɪt/ (n)}
{An estimate of income and expenditure for a set period of time.}
{We are traveling on a tight budget, so we are staying in hostels and eating street food.}
{Đây là một danh từ dùng để chỉ ngân sách chuyến đi.}

\vocabentry{Canyon}
{/ˈkænjən/ (n)}
{A deep gorge, typically one with a river flowing through it.}
{The Grand Canyon is one of the most famous natural landmarks in the world.}
{Đây là một danh từ dùng để chỉ hẻm núi.}

\vocabentry{Cargo}
{/ˈkɑːrɡoʊ/ (n)}
{Goods carried on a ship, aircraft, or motor vehicle.}
{The ship was carrying a large cargo of luxury goods.}
{Liên quan đến vận chuyển hàng hóa du lịch.}

\vocabentry{Caravan}
{/ˈkærəvæn/ (n)}
{A vehicle equipped for living in, typically towed by a car and used for holidays.}
{We spent the summer traveling around Europe in a caravan.}
{Đây là một danh từ dùng để chỉ xe nhà di động.}

\vocabentry{Cathedral}
{/kəˈθiːdrəl/ (n)}
{The principal church of a diocese, with which the bishop is officially associated.}
{The city's cathedral is a masterpiece of Gothic architecture.}
{Đây là một danh từ dùng để chỉ nhà thờ lớn.}

\vocabentry{Cemetery}
{/ˈseməteri/ (n)}
{A burial ground; a graveyard.}
{Highgate Cemetery in London is a popular destination for tourists interested in history.}
{Một số nghĩa trang lịch sử là điểm tham quan du lịch.}

\vocabentry{Chamber}
{/ˈtʃeɪmbər/ (n)}
{A large room used for formal or public events.}
{The cave system contains several large chambers with impressive stalactites.}
{Thường dùng để mô tả các phòng trong cung điện hoặc hang động.}

\vocabentry{Chartered}
{/ˈtʃɑːrtərd/ (adj)}
{(Of a ship, aircraft, or bus) rented for a special use.}
{The group took a chartered bus to the ski resort.}
{}

\vocabentry{Coastline}
{/ˈkoʊstlaɪn/ (n)}
{The land along a coast.}
{The country has a beautiful coastline with many sandy beaches and rocky cliffs.}
{Đây là một danh từ dùng để chỉ đường bờ biển.}

\vocabentry{Concourse}
{/ˈkɑːŋkɔːrs/ (n)}
{A large open area inside or in front of a public building.}
{We waited for our friends in the main concourse of the airport.}
{Đây là một danh từ dùng để chỉ sảnh lớn ở sân bay hoặc nhà ga.}

\vocabentry{Connoisseur}
{/ˌkɑːnəˈsɜːr/ (n)}
{An expert judge in matters of taste.}
{The wine tour is designed for connoisseurs who want to learn more about the local vineyards.}
{Thường dùng cho du khách sành ăn uống - food connoisseur.}

\vocabentry{Continent}
{/ˈkɑːntɪnənt/ (n)}
{Any of the world's main continuous expanses of land.}
{Africa is a vast continent with many different cultures and landscapes.}
{Đây là một danh từ dùng để chỉ lục địa.}

\vocabentry{Cosmopolitan}
{/ˌkɑːzməˈpɑːlɪtən/ (adj)}
{Including people from many different countries and cultures.}
{London is a very cosmopolitan city where you can find food from almost every country in the world.}
{Đây là một tính từ dùng để chỉ tính đa văn hóa của một thành phố.}

\vocabentry{Cruise ship}
{/ˈkruːz ʃɪp/ (n)}
{A large ship that carries people on voyages for pleasure.}
{The cruise ship has several restaurants, a theater, and even a small water park.}
{Đây là một cụm danh từ dùng để chỉ tàu du lịch hạng sang.}

\vocabentry{Day trip}
{/ˈdeɪ trɪp/ (n)}
{A journey or visit completed in one day.}
{We took a day trip to the seaside to escape the heat of the city.}
{Đây là một cụm danh từ dùng để chỉ chuyến đi trong ngày.}

\vocabentry{Delayed}
{/dɪˈleɪd/ (adj)}
{Late.}
{Our flight was delayed by three hours, so we missed our connecting flight.}
{Đây là một tính từ dùng để chỉ sự chậm trễ chuyến bay/tàu.}

\vocabentry{Delicacy}
{/ˈdelɪkəsi/ (n)}
{A choice or expensive food.}
{Caviar is considered a delicacy in many parts of the world.}
{Đây là một danh từ dùng để chỉ món cao lương mỹ vị/đặc sản.}

\vocabentry{Departure lounge}
{/dɪˈpɑːrtʃər laʊndʒ/ (n)}
{A seating area in an airport where passengers wait to board their aircraft.}
{We spent an hour in the departure lounge before our flight was called.}
{Đây là một cụm danh từ dùng để chỉ phòng chờ khởi hành.}

\vocabentry{Dock}
{/dɑːk/ (v)}
{(Of a ship) come into a dock and stay there.}
{The cruise ship will dock in Barcelona tomorrow morning.}
{Đây là một động từ dùng để chỉ việc tàu cập bến.}

\vocabentry{Dormitory}
{/ˈdɔːrmətɔːri/ (n)}
{A large bedroom containing a number of beds in a school or institution.}
{Staying in a dormitory is a cheap way to travel if you don't mind sharing a room.}
{Thường thấy trong các Hostel cho dân du lịch bụi.}

\vocabentry{Double room}
{/ˌdʌbl ˈruːm/ (n)}
{A room for two people.}
{We booked a double room with a view of the sea.}
{Đây là một cụm danh từ dùng để chỉ phòng đôi.}

\vocabentry{Duty-free}
{/ˌduːti ˈfriː/ (adj)}
{(Of goods) exempt from payment of duty.}
{You can buy duty-free perfume and alcohol at the airport.}
{Đây là một tính từ dùng để chỉ hàng miễn thuế.}

\vocabentry{Ecology}
{/iˈkɑːlədʒi/ (n)}
{The branch of biology that deals with the relations of organisms to one another and to their physical surroundings.}
{The island's fragile ecology is being threatened by the increase in tourism.}
{Liên quan đến du lịch sinh thái.}

\vocabentry{Educational}
{/ˌedʒuˈkeɪʃənl/ (adj)}
{Relating to education.}
{The school trip to the museum was very educational for the students.}
{Đây là một tính từ dùng để chỉ loại hình du lịch giáo dục.}

\vocabentry{Embarkation}
{/ˌembɑːrˈkeɪʃn/ (n)}
{The act of getting on a ship or aircraft.}
{Please have your boarding pass ready for embarkation.}
{Đây là một danh từ dùng để chỉ sự lên tàu/máy bay.}

\vocabentry{Enchanting}
{/ɪnˈtʃæntɪŋ/ (adj)}
{Delightfully charming or attractive.}
{We spent an enchanting evening wandering through the old town.}
{Đây là một tính từ dùng để chỉ vẻ đẹp mê hoặc, quyến rũ.}

\vocabentry{En route}
{/ˌɑːn ˈruːt/ (adv)}
{On the way.}
{We stopped for lunch in a small village en route to the capital.}
{Đây là một trạng từ dùng để chỉ đang trên đường đi.}

\vocabentry{Environment}
{/ɪnˈvaɪrənmənt/ (n)}
{The surroundings or conditions in which a person, animal, or plant lives or operates.}
{We must protect the natural environment for future generations of travelers.}
{Quan trọng trong du lịch bền vững.}

\vocabentry{Equipment}
{/ɪˈkwɪpmənt/ (n)}
{The necessary items for a particular purpose. (Đây là một danh từ dùng để chỉ trang thiết bị (ví dụ: đồ leo núi).).}
{You need to have the right equipment before you go hiking in the mountains.}
{}

\vocabentry{Ethnicity}
{/eθˈnɪsəti/ (n)}
{The fact or state of belonging to a social group that has a common national or cultural tradition.}
{The city is a melting pot of different ethnicities and cultures.}
{Liên quan đến du lịch văn hóa.}

\vocabentry{Explorer}
{/ɪkˈsplɔːrər/ (n)}
{A person who explores an unfamiliar area.}
{Marco Polo was one of the most famous explorers in history.}
{Đây là một danh từ dùng để chỉ nhà thám hiểm.}

\vocabentry{Facilities}
{/fəˈsɪləti/ (n)}
{Places, amenities, or pieces of equipment provided for a particular purpose.}
{The campsite has excellent facilities, including a shower block and a small shop.}
{Đây là một danh từ dùng để chỉ cơ sở vật chất.}

\vocabentry{Fair}
{/fer/ (n)}
{A periodic gathering for the sale of goods, often with shows and entertainment.}
{We visited a traditional craft fair and bought some handmade pottery.}
{Thường dùng cho hội chợ du lịch - tourism fair.}

\vocabentry{Farewell}
{/ˌferˈwel/ (n)}
{An act of parting or of marking someone's departure.}
{The hotel organized a farewell dinner for the guests on their last night.}
{Dùng trong các buổi tiệc chia tay khách du lịch.}

\vocabentry{Fasten}
{/ˈfæsn/ (v)}
{Close or do up securely.}
{Please fasten your seatbelt as we are experiencing some turbulence.}
{Thường dùng: fasten your seatbelt - thắt dây an toàn.}

\vocabentry{Ferry}
{/ˈferi/ (n)}
{A boat or ship for conveying passengers and goods, especially over a relatively short distance and as a regular service.}
{We took a ferry across the river to get to the other side of the city.}
{Đây là một danh từ dùng để chỉ phà.}

\vocabentry{Festival}
{/ˈfestɪvl/ (n)}
{A day or period of celebration.}
{The city's annual film festival attracts thousands of tourists every year.}
{Đây là một danh từ dùng để chỉ lễ hội.}

\vocabentry{First class}
{/ˌfɜːrst ˈklæs/ (n)}
{The highest class of travel on a plane, train, etc.}
{Traveling in first class is much more comfortable, but it's also very expensive.}
{Đây là một danh từ dùng để chỉ khoang hạng nhất.}

\vocabentry{Flight}
{/flaɪt/ (n)}
{An act of passing through the air by use of wings or a vehicle.}
{Our flight to London takes about seven hours.}
{Đây là một danh từ dùng để chỉ chuyến bay.}

\vocabentry{Flock}
{/flɑːk/ (v)}
{Congregate or mass in a large group or crowd.}
{Thousands of people flock to the city every year to see the cherry blossoms.}
{Thường dùng: tourists flock to... - du khách đổ xô đến...}

\vocabentry{Folklore}
{/ˈfoʊklɔːr/ (n)}
{The traditional beliefs, customs, and stories of a community, passed through the generations by word of mouth.}
{The local museum has an interesting exhibition on Irish folklore.}
{Liên quan đến du lịch văn hóa.}

\vocabentry{Foreign}
{/ˈfɔːrən/ (adj)}
{Of, from, in, or characteristic of a country or language other than one's own.}
{Learning a foreign language can make your travels much more enjoyable.}
{Đây là một tính từ dùng để chỉ nước ngoài.}

\vocabentry{Fortress}
{/ˈfɔːrtrəs/ (n)}
{A military stronghold, especially a strongly fortified town.}
{The ancient fortress was built to protect the city from invaders.}
{Điểm tham quan lịch sử.}

\vocabentry{Front desk}
{/ˌfrʌnt ˈdesk/ (n)}
{The desk at the entrance of a hotel.}
{Please go to the front desk to pick up your room key.}
{Đây là một cụm danh từ dùng để chỉ quầy lễ tân.}

\vocabentry{Gala}
{/ˈɡeɪlə/ (n)}
{A social occasion with special entertainments or performances.}
{The hotel held a gala dinner to celebrate its 50th anniversary.}
{Thường là các buổi tiệc du lịch lớn.}

\vocabentry{Gallery}
{/ˈɡæləri/ (n)}
{A room or building for the display or sale of works of art.}
{The city has many art galleries featuring the work of local and international artists.}
{Đây là một danh từ dùng để chỉ phòng triển lãm nghệ thuật.}

\vocabentry{Garment}
{/ˈɡɑːrmənt/ (n)}
{An item of clothing.}
{We bought some traditional silk garments at the market.}
{Thường liên quan đến trang phục truyền thống của địa phương.}

\vocabentry{Gaze}
{/ɡeɪz/ (v)}
{Look steadily and intently, especially in admiration, surprise, or thought.}
{We stood on the cliff and gazed out at the vast ocean.}
{Dùng khi du khách ngắm nhìn cảnh đẹp.}

\vocabentry{Gem}
{/dʒem/ (n)}
{A precious or semiprecious stone; something valued for its beauty or perfection.}
{This small island is a real hidden gem that is not yet known to many tourists.}
{Thường dùng: hidden gem - viên ngọc quý tiềm ẩn.}

\vocabentry{Geographic}
{/ˌdʒiːəˈɡræfɪk/ (adj)}
{Relating to geography.}
{The country has a very diverse geographic landscape, with mountains, deserts, and forests.}
{Đây là một tính từ dùng để chỉ các đặc điểm địa lý.}

\vocabentry{Geology}
{/dʒiˈɑːlədʒi/ (n)}
{The science that deals with the earth's physical structure and substance.}
{The park is famous for its unique geology, including ancient volcanic rock formations.}
{Liên quan đến du lịch địa chất.}

\vocabentry{Gorge}
{/ɡɔːrdʒ/ (n)}
{A narrow valley between hills or mountains, typically with steep rocky walls and a stream running through it.}
{We hiked through the gorge and enjoyed the stunning views.}
{Đây là một danh từ dùng để chỉ hẻm núi sâu.}

\vocabentry{Grandeur}
{/ˈɡrændʒər/ (n)}
{Splendor and impressiveness, especially of appearance or style.}
{We were awed by the grandeur of the ancient palace.}
{Đây là một danh từ dùng để chỉ sự vĩ đại, tráng lệ.}

\vocabentry{Greenway}
{/ˈɡriːnweɪ/ (n)}
{A strip of undeveloped land near an urban area, set aside for recreational use or environmental protection.}
{The city has several greenways that are popular with cyclists and walkers.}
{Điểm du lịch dã ngoại.}

\vocabentry{Grotto}
{/ˈɡrɑːtoʊ/ (n)}
{A small picturesque cave, especially an artificial one in a park or garden.}
{The boat took us into a beautiful blue grotto where the water was crystal clear.}
{Đây là một danh từ dùng để chỉ hang động nhỏ.}

\vocabentry{Guided}
{/ˈɡaɪdɪd/ (adj)}
{Directed or supervised by a guide.}
{We took a guided tour of the city to learn more about its history.}
{}

\vocabentry{Gulf}
{/ɡʌlf/ (n)}
{A deep inlet of the sea almost surrounded by land, with a narrow mouth.}
{The country is located on the shores of a beautiful gulf.}
{Đây là một danh từ dùng để chỉ vịnh biển.}

\vocabentry{Habitation}
{/ˌhæbɪˈteɪʃn/ (n)}
{The state or process of living in a particular place.}
{The area has been a site of human habitation for thousands of years.}
{Thường dùng để mô tả các khu định cư cổ.}

\vocabentry{Harbor}
{/ˈhɑːrbər/ (n)}
{A place on the coast where vessels may find shelter.}
{The harbor was full of fishing boats and luxury yachts.}
{Đây là một danh từ dùng để chỉ bến cảng.}

\vocabentry{Haze}
{/heɪz/ (n)}
{A slight obscuration of the lower atmosphere.}
{The heat haze made it difficult to see the mountains in the distance.}
{Thường liên quan đến thời tiết khi đi du lịch.}

\vocabentry{Headquarters}
{/ˈhedkwɔːrtərz/ (n)}
{The premises serving as the center of an organization.}
{The airline's headquarters are located near the international airport.}
{Trong du lịch: Head office của hãng lữ hành.}

\vocabentry{Heath}
{/hiːθ/ (n)}
{An area of open uncultivated land.}
{We spent the afternoon walking across the heath and enjoying the fresh air.}
{Cảnh quan thiên nhiên để tham quan.}

\vocabentry{Highland}
{/ˈhaɪlənd/ (n)}
{An area of high or mountainous land.}
{The Scottish Highlands are famous for their rugged beauty and dramatic landscapes.}
{Đây là một danh từ dùng để chỉ vùng cao/cao nguyên.}

\vocabentry{Hike}
{/haɪk/ (v)}
{Walk for a long distance, especially across country or in the woods.}
{We plan to hike to the top of the mountain tomorrow morning.}
{Đây là một động từ dùng để chỉ việc đi bộ đường dài/leo núi.}

\vocabentry{Hillside}
{/ˈhɪlsaɪd/ (n)}
{The sloping side of a hill.}
{The hillside was covered in colorful wildflowers.}
{Đây là một danh từ dùng để chỉ sườn đồi.}

\vocabentry{Historic}
{/hɪˈstɔːrɪk/ (adj)}
{Famous or important in history.}
{The city has many historic buildings that have been beautifully restored.}
{Đây là một tính từ dùng để chỉ các địa danh có giá trị lịch sử.}

\vocabentry{Holidays}
{/ˈhɑːlədeɪz/ (n)}
{A period of time during which you relax and enjoy yourself away from home.}
{Where are you going for your summer holidays this year?}
{Đây là một danh từ dùng để chỉ kỳ nghỉ.}

\vocabentry{Homestead}
{/ˈhoʊmsted/ (n)}
{A person's or family's residence, which comprises the land, house, and outbuildings.}
{We stayed in a traditional homestead and learned how to make cheese and bread.}
{Du lịch trải nghiệm cuộc sống trang trại.}

\vocabentry{Horizon}
{/həˈraɪzn/ (n)}
{The line at which the earth's surface and the sky appear to meet.}
{We sat on the beach and watched the sun sink below the horizon.}
{Đây là một danh từ dùng để chỉ đường chân trời.}

\vocabentry{Hot spring}
{/ˌhɑːt ˈsprɪŋ/ (n)}
{A spring of naturally hot water.}
{The area is famous for its natural hot springs and mud baths.}
{Đây là một cụm danh từ dùng để chỉ suối nước nóng.}

\vocabentry{Hotel}
{/hoʊˈtel/ (n)}
{An establishment providing accommodation, meals, and other services for travelers and tourists.}
{We booked a five-star hotel with a view of the harbor.}
{Đây là một danh từ dùng để chỉ khách sạn.}

\vocabentry{Houseboat}
{/ˈhaʊsboʊt/ (n)}
{A boat that is or can be used as a home.}
{We spent a week traveling along the river on a houseboat.}
{Đây là một danh từ dùng để chỉ nhà thuyền.}

\vocabentry{Hovercraft}
{/ˈhʌvərkræft/ (n)}
{A vehicle or craft that travels over water or land on a cushion of air.}
{The hovercraft is the fastest way to get across the bay.}
{Phương tiện vận chuyển du khách.}

\vocabentry{Huddle}
{/ˈhʌdl/ (v)}
{Crowd together; nestle closely.}
{We huddled together around the campfire to keep warm.}
{Thường mô tả du khách tụ tập khi thời tiết lạnh.}

\vocabentry{Hygiene}
{/ˈhaɪdʒiːn/ (n)}
{Conditions or practices conducive to maintaining health and preventing disease, especially through cleanliness.}
{You should be careful about food hygiene when traveling to certain countries.}
{Rất quan trọng khi đi du lịch.}

\vocabentry{Identity}
{/aɪˈdentəti/ (n)}
{The fact of being who or what a person or thing is.}
{The city is proud of its unique cultural identity and traditions.}
{Liên quan đến bản sắc văn hóa vùng miền.}

\vocabentry{Idyllic}
{/aɪˈdɪlɪk/ (adj)}
{(Especially of a time or place) like an idyll; extremely happy, peaceful, or picturesque.}
{We spent an idyllic week in a small cottage by the sea.}
{Đây là một tính từ dùng để chỉ vẻ đẹp bình dị, thôn dã.}

\vocabentry{Immigration}
{/ˌɪmɪˈɡreɪʃn/ (n)}
{The action of coming to live permanently in a foreign country.}
{You have to go through immigration and customs after arriving at the international airport.}
{Thủ tục kiểm soát tại sân bay.}

\vocabentry{Indigenous}
{/ɪnˈdɪdʒənəs/ (adj)}
{Originating or occurring naturally in a particular place; native.}
{The museum has a collection of art and artifacts from the indigenous people of the area.}
{Liên quan đến du lịch văn hóa bản địa.}

\vocabentry{Inhabitant}
{/ɪnˈhæbɪtənt/ (n)}
{A person or animal that lives in or occupies a place.}
{The city's inhabitants are very friendly and welcoming to tourists.}
{Đây là một danh từ dùng để chỉ cư dân địa phương.}

\vocabentry{Inland}
{/ˌɪnˈlænd/ (adj)}
{Situated in the interior of a country rather than on the coast.}
{We decided to explore the inland part of the country, far from the crowded beaches.}
{Đây là một tính từ dùng để chỉ các vùng nội địa.}

\vocabentry{Inlet}
{/ˈɪnlet/ (n)}
{A small arm of the sea, a lake, or a river.}
{The coastline is full of small inlets and hidden coves.}
{Cảnh quan thiên nhiên để du lịch tàu thủy.}

\vocabentry{Insulate}
{/ˈɪnsəleɪt/ (v)}
{Protect (something) by interposing material that prevents the loss of heat or the intrusion of sound.}
{The hotel rooms are well insulated, so you can't hear the noise from the street.}
{Thường dùng cho các khu nghỉ dưỡng cao cấp.}

\vocabentry{Interaction}
{/ˌɪntərˈækʃn/ (n)}
{Reciprocal action or influence.}
{Interaction with local people is one of the most rewarding parts of traveling.}
{Du lịch tương tác văn hóa.}

\vocabentry{Intercontinental}
{/ˌɪntərˌkɑːntɪˈnentl/ (adj)}
{Relating to or traveling between continents.}
{Intercontinental flights can be very long and tiring.}
{Đây là một tính từ dùng để chỉ chuyến bay xuyên lục địa.}

\vocabentry{Intrepid}
{/ɪnˈtrepɪd/ (adj)}
{Fearless; adventurous.}
{The intrepid explorer set off alone to cross the desert.}
{Dùng để mô tả các du khách gan dạ, thích mạo hiểm.}

\vocabentry{Island}
{/ˈaɪlənd/ (n)}
{A piece of land surrounded by water.}
{We spent our holiday on a small island in the Pacific.}
{Đây là một danh từ dùng để chỉ hòn đảo.}

\vocabentry{Jet}
{/dʒet/ (n)}
{A rapid stream of liquid or gas forced out of a small opening; an aircraft powered by jet engines.}
{We flew to London on a large commercial jet.}
{Đây là một danh từ dùng để chỉ máy bay phản lực.}

\vocabentry{Journey}
{/ˈdʒɜːrni/ (n)}
{An act of traveling from one place to another.}
{The journey from London to Paris takes about two and a half hours by train.}
{Đây là một danh từ dùng để chỉ chuyến đi/hành trình.}

\vocabentry{Jungle}
{/ˈdʒʌŋɡl/ (n)}
{An area of land overgrown with dense forest and tangled vegetation, typically in the tropics.}
{We trekked through the jungle for three days to reach the ancient ruins.}
{Đây là một danh từ dùng để chỉ rừng nhiệt đới.}

\vocabentry{Kayaking}
{/ˈkaɪækɪŋ/ (n)}
{The activity of traveling in a kayak.}
{Kayaking is a popular activity for tourists visiting the coast.}
{Đây là một danh từ dùng để chỉ môn chèo thuyền kayak.}

\vocabentry{Lagoon}
{/ləˈɡuːn/ (n)}
{A stretch of salt water separated from the sea by a low sandbank or coral reef.}
{The resort is located on the shores of a beautiful blue lagoon.}
{Đây là một danh từ dùng để chỉ đầm phá.}

\vocabentry{Lake}
{/leɪk/ (n)}
{A large body of water surrounded by land.}
{The lake is a popular destination for fishing and boating.}
{Đây là một danh từ dùng để chỉ hồ nước.}

\vocabentry{Landing}
{/ˈlændɪŋ/ (n)}
{An act of coming to land after a voyage or flight.}
{The plane made a smooth landing at the international airport.}
{Đây là một danh từ dùng để chỉ sự hạ cánh.}

\vocabentry{Landscape}
{/ˈlændskeɪp/ (n)}
{All the visible features of an area of countryside or land.}
{The landscape changed from rolling hills to flat plains as we traveled north.}
{Đây là một danh từ dùng để chỉ phong cảnh/cảnh quan.}

\vocabentry{Legend}
{/ˈledʒənd/ (n)}
{A traditional story sometimes popularly regarded as historical but unauthenticated.}
{The ancient ruins are the subject of many local legends.}
{Liên quan đến du lịch văn hóa truyền thuyết địa phương.}

\vocabentry{Leisure}
{/ˈliːʒər/ (n)}
{Use of free time for enjoyment.}
{Travelers use their leisure time to explore new places and relax.}
{Đây là một danh từ dùng để chỉ sự thư giãn/giải trí.}

\vocabentry{Lighthouse}
{/ˈlaɪthaʊs/ (n)}
{A tower or other structure containing a beacon light to warn or guide ships at sea.}
{We climbed to the top of the lighthouse for a great view of the coast.}
{Địa điểm tham quan ven biển.}

\vocabentry{Limestone}
{/ˈlaɪmstoʊn/ (n)}
{A hard sedimentary rock, composed mainly of calcium carbonate or dolomite.}
{The area is famous for its limestone caves and dramatic cliffs.}
{Thường tạo nên các hang động và núi đá vôi đẹp.}

\vocabentry{Locality}
{/loʊˈkæləti/ (n)}
{An area or neighborhood.}
{Each locality has its own unique character and traditions.}
{Đây là một danh từ dùng để chỉ địa phương/khu vực cụ thể.}

\vocabentry{Lodge}
{/lɑːdʒ/ (n)}
{A small house at the gates of a park or in the grounds of a large house, occupied by a gatekeeper, gardener, or other employee.}
{We stayed in a luxury safari lodge in the heart of the national park.}
{Thường là nơi lưu trú ở vùng rừng núi.}

\vocabentry{Lowland}
{/ˈloʊlənd/ (n)}
{Low-lying country.}
{The country's lowlands are primarily used for agriculture.}
{Đây là một danh từ dùng để chỉ vùng đất thấp/đồng bằng.}

\vocabentry{Mainstream}
{/ˈmeɪnstriːm/ (adj)}
{Representing widespread current thought.}
{Mainstream tourism often focuses on popular destinations and all-inclusive resorts.}
{Trong du lịch: du lịch đại chúng, phổ biến.}

\vocabentry{Manifest}
{/ˈmænɪfest/ (n)}
{A document listing a ship's contents, cargo, passengers, and crew.}
{The ship's manifest showed that there were over five hundred passengers on board.}
{Đây là một danh từ dùng để chỉ bản kê khai hành khách/hàng hóa.}

