% Auto-generated from raw.txt
% Usage:
%   \vocabentry{word}{/ipa/ (pos)}{definition}{example}{note}

\vocabentry{Abundance}
{/əˈbʌndəns/ (n)}
{A very large quantity of something.}
{The region is known for its abundance of natural gas and precious metals.}
{Đây là một danh từ dùng để chỉ sự phong phú hoặc sự dồi dào của một nguồn tài nguyên nào đó.}

\vocabentry{Allocation}
{/ˌæləˈkeɪʃn/ (n)}
{The action or process of distributing something.}
{The fair allocation of water resources is a major challenge for the government.}
{Đây là một danh từ dùng để chỉ sự phân bổ tài nguyên cho các mục đích hoặc khu vực khác nhau.}

\vocabentry{Alternative}
{/ɔːlˈtɜːrnətɪv/ (adj)}
{(Of fuel or energy) available as another possibility or hope.}
{Scientists are developing alternative energy sources to reduce our reliance on oil.}
{Đây là một tính từ dùng để chỉ các nguồn thay thế, thường là năng lượng sạch thay cho nhiên liệu hóa thạch.}

\vocabentry{Anthropogenic}
{/ˌænθrəpəˈdʒenɪk/ (adj)}
{Originating in human activity.}
{Anthropogenic climate change is depleting natural water reserves in many regions.}
{Đây là một tính từ dùng để chỉ các tác động hoặc sự thay đổi tài nguyên do con người gây ra.}

\vocabentry{Aquifer}
{/ˈækwɪfər/ (n)}
{A body of permeable rock which can contain or transmit groundwater.}
{Farmers rely on the local aquifer to provide water for their crops during the dry season.}
{Đây là một danh từ dùng để chỉ tầng ngậm nước dưới lòng đất.}

\vocabentry{Arable}
{/ˈærəbl/ (adj)}
{(Of land) used or suitable for growing crops.}
{The country is losing much of its arable land to urbanization and soil erosion.}
{Đây là một tính từ dùng để chỉ đất canh tác được, một nguồn tài nguyên đất quan trọng.}

\vocabentry{Barren}
{/ˈbærən/ (adj)}
{(Of land) too poor to produce much or any vegetation.}
{After years of over-farming, the once fertile fields became barren and dry.}
{Đây là một tính từ dùng để chỉ đất đai cằn cỗi, không thể khai thác nông nghiệp.}

\vocabentry{Biomass}
{/ˈbaɪoʊmæs/ (n)}
{Organic matter used as a fuel, especially in a power station.}
{Biomass is considered a renewable resource because it comes from plants and animals.}
{Đây là một danh từ dùng để chỉ sinh khối, nguồn năng lượng từ chất hữu cơ.}

\vocabentry{By-product}
{/ˈbaɪ ˌprɑːdʌkt/ (n)}
{An incidental or secondary product made in the manufacture of something else.}
{Natural gas is often a by-product of oil extraction.}
{Đây là một danh từ dùng để chỉ tác phẩm phụ hoặc phụ phẩm trong quá trình khai thác tài nguyên.}

\vocabentry{Capital}
{/ˈkæpɪtl/ (n)}
{Wealth in the form of money or other assets owned by a person or organization.}
{Protecting our natural capital is essential for long-term economic stability.}
{Đây là một danh từ, trong ngữ cảnh này thường chỉ "natural capital" - vốn tự nhiên.}

\vocabentry{Circulation}
{/ˌsɜːrkjəˈleɪʃn/ (n)}
{The movement to and fro or around something.}
{Ocean circulation helps distribute heat around the planet.}
{Đây là một danh từ dùng để chỉ sự lưu thông, ví dụ như sự lưu thông của dòng nước hoặc không khí.}

\vocabentry{Commodity}
{/kəˈmɑːdəti/ (n)}
{A raw material or primary agricultural product that can be bought and sold.}
{Oil is the most traded commodity in the global market.}
{Đây là một danh từ dùng để chỉ hàng hóa là nguyên liệu thô.}

\vocabentry{Conservation}
{/ˌkɑːnsərˈveɪʃn/ (n)}
{The protection of natural resources from the damaging effects of human activity.}
{Water conservation is becoming increasingly important due to global warming.}
{Đây là một danh từ dùng để chỉ sự bảo tồn tài nguyên.}

\vocabentry{Consumption}
{/kənˈsʌmpʃn/ (n)}
{The using up of a resource.}
{The high consumption of fossil fuels is contributing to environmental degradation.}
{Đây là một danh từ dùng để chỉ sự tiêu thụ tài nguyên.}

\vocabentry{Crude}
{/kruːd/ (adj)}
{In a natural or raw state; not yet processed or refined. (Đây là một tính từ dùng để chỉ trạng thái thô, chưa qua chế biến (thường dùng cho dầu thô).).}
{The country exports millions of barrels of crude oil every day.}
{}

\vocabentry{Depletion}
{/dɪˈpliːʃn/ (n)}
{Reduction in the number or quantity of something.}
{The depletion of the ozone layer is a serious environmental concern.}
{Đây là một danh từ dùng để chỉ sự cạn kiệt tài nguyên.}

\vocabentry{Deposit}
{/dɪˈpɑːzɪt/ (n)}
{A layer or body of accumulated matter.}
{Geologists have discovered a massive deposit of copper in the mountains.}
{Đây là một danh từ dùng để chỉ các mỏ hoặc trữ lượng khoáng sản.}

\vocabentry{Desalination}
{/ˌdiːˌsælɪˈneɪʃn/ (n)}
{The process of removing salt from seawater.}
{Desalination is a crucial technology for water-scarce regions.}
{Đây là một danh từ dùng để chỉ sự khử muối để tạo ra nước ngọt.}

\vocabentry{Distribution}
{/ˌdɪstrɪˈbjuːʃn/ (n)}
{The action of sharing something out among a number of recipients.}
{The uneven distribution of natural resources often leads to international tension.}
{Đây là một danh từ dùng để chỉ sự phân bổ tài nguyên trên bề mặt Trái Đất.}

\vocabentry{Diversification}
{/daɪˌvɜːrsɪfɪˈkeɪʃn/ (n)}
{The process of becoming more varied.}
{Energy diversification is necessary to ensure a stable power supply.}
{Đây là một danh từ dùng để chỉ sự đa dạng hóa, ví dụ như đa dạng hóa nguồn năng lượng.}

\vocabentry{Efficiency}
{/ɪˈfɪʃnsi/ (n)}
{The state or quality of being efficient.}
{Improving energy efficiency can significantly reduce utility bills.}
{Đây là một danh từ dùng để chỉ hiệu suất khai thác hoặc sử dụng tài nguyên.}

\vocabentry{Encroachment}
{/ɪnˈkroʊtʃmənt/ (n)}
{Intrusion on a person's territory or rights.}
{Urban encroachment is destroying the local forest ecosystem.}
{Đây là một danh từ dùng để chỉ sự xâm lấn của con người vào các vùng tài nguyên tự nhiên.}

\vocabentry{Endowment}
{/ɪnˈdaʊmənt/ (n)}
{A quality or ability possessed or inherited by someone.}
{The country's natural endowment of minerals has driven its economic growth.}
{Đây là một danh từ, trong ngữ cảnh này chỉ "natural endowment" - tài nguyên thiên nhiên sẵn có.}

\vocabentry{Exhaustible}
{/ɪɡˈzɔːstəbl/ (adj)}
{Capable of being used up.}
{Fossil fuels are exhaustible resources that will eventually run out.}
{Đây là một tính từ dùng để chỉ các tài nguyên có thể bị cạn kiệt, không tái tạo được.}

\vocabentry{Extraction}
{/ɪkˈstrækʃn/ (n)}
{The action of taking out something, especially using effort or force. (Đây là một danh từ dùng để chỉ sự khai thác, rút ra (ví dụ: khai thác mỏ).).}
{The extraction of minerals can have a devastating impact on the environment.}
{}

\vocabentry{Fertility}
{/fərˈtɪləti/ (n)}
{The quality of being fertile.}
{Soil fertility can be maintained through natural composting and crop rotation.}
{Đây là một danh từ dùng để chỉ độ màu mỡ của đất.}

\vocabentry{Fossil}
{/ˈfɑːsl/ (n)}
{The remains or impression of a prehistoric organism.}
{Coal is a fossil fuel formed from the remains of ancient plants.}
{Đây là một danh từ dùng để chỉ hóa thạch, nguồn gốc của nhiên liệu hóa thạch.}

\vocabentry{Geothermal}
{/ˌdʒiːoʊˈθɜːrml/ (adj)}
{Relating to or produced by the internal heat of the earth.}
{Geothermal energy provides a constant and renewable source of heat.}
{Đây là một tính từ dùng để chỉ nguồn địa nhiệt.}

\vocabentry{Habitat}
{/ˈhæbɪtæt/ (n)}
{The natural home or environment of an animal or plant.}
{The destruction of natural habitats is a major cause of species extinction.}
{Đây là một danh từ dùng để chỉ môi trường sống tự nhiên, một loại tài nguyên sinh học.}

\vocabentry{Harness}
{/ˈhɑːrnɪs/ (v)}
{Control and make use of (natural resources), especially to produce energy.}
{We need to learn how to harness the power of the wind more effectively.}
{Đây là một động từ dùng để chỉ việc khai thác và tận dụng nguồn năng lượng tự nhiên.}

\vocabentry{Hydrocarbon}
{/ˌhaɪdrəˈkɑːrbən/ (n)}
{A compound of hydrogen and carbon, such as those that are the chief components of petroleum.}
{Hydrocarbons are the primary source of energy for the transport industry.}
{Đây là một danh từ dùng để chỉ hydrocacbon, thành phần chính của dầu mỏ.}

\vocabentry{Hydropower}
{/ˈhaɪdroʊpaʊər/ (n)}
{Electricity produced from the energy of moving water.}
{Hydropower is one of the most widely used forms of renewable energy.}
{Đây là một danh từ dùng để chỉ thủy điện.}

\vocabentry{Inexhaustible}
{/ˌɪnɪɡˈzɔːstəbl/ (adj)}
{Unable to be used up because existing in abundance.}
{Solar energy is an inexhaustible resource that will last for billions of years.}
{Đây là một tính từ dùng để chỉ các nguồn tài nguyên vô tận.}

\vocabentry{Infrastructure}
{/ˈɪnfrəstrʌktʃər/ (n)}
{The basic physical and organizational structures needed for the operation of a society.}
{The country needs to invest in its water infrastructure to prevent shortages.}
{Đây là một danh từ dùng để chỉ cơ sở hạ tầng khai thác tài nguyên.}

\vocabentry{Irrigation}
{/ˌɪrɪˈɡeɪʃn/ (n)}
{The supply of water to land or crops to help growth.}
{Modern irrigation techniques can help save a significant amount of water.}
{Đây là một danh từ dùng để chỉ hệ thống tưới tiêu tài nguyên nước.}

\vocabentry{Jurisdiction}
{/ˌdʒʊrɪsˈdɪkʃn/ (n)}
{The official power to make legal decisions and judgments.}
{The mineral deposits fall under the jurisdiction of the national government.}
{Đây là một danh từ dùng để chỉ quyền tài phán đối với tài nguyên thiên nhiên.}

\vocabentry{Kinetic}
{/kɪˈnetɪk/ (adj)}
{Relating to or resulting from motion.}
{Wind turbines convert the kinetic energy of the wind into electricity.}
{Đây là một tính từ dùng để chỉ động năng, thường nói về năng lượng gió hoặc nước.}

\vocabentry{Lumber}
{/ˈlʌmbər/ (n)}
{Raw wood that has been prepared for use as a building material.}
{The lumber industry is a major part of the local economy.}
{Đây là một danh từ dùng để chỉ gỗ xẻ, tài nguyên rừng.}

\vocabentry{Management}
{/ˈmænɪdʒmənt/ (n)}
{The process of dealing with or controlling things.}
{Sustainable forest management is essential for preserving biodiversity.}
{Đây là một danh từ dùng để chỉ sự quản lý tài nguyên.}

\vocabentry{Metallic}
{/məˈtælɪk/ (adj)}
{Relating to or resembling metal or metals.}
{Iron and aluminum are important metallic resources for construction.}
{Đây là một tính từ dùng để chỉ các tài nguyên thuộc kim loại.}

\vocabentry{Mineral}
{/ˈmɪnərəl/ (n)}
{A solid inorganic substance of natural occurrence.}
{The mountains are rich in minerals like gold, silver, and copper.}
{Đây là một danh từ dùng để chỉ khoáng sản.}

\vocabentry{Natural}
{/ˈnætʃrəl/ (adj)}
{Existing in or caused by nature; not made or caused by humankind.}
{We must protect our natural environment for future generations.}
{Đây là một tính từ dùng để chỉ những gì thuộc về tự nhiên.}

\vocabentry{Non-renewable}
{/ˌnɑːn rɪˈnuːəbl/ (adj)}
{Existing in finite quantity; not capable of being replenished.}
{Fossil fuels are non-renewable and produce harmful greenhouse gases.}
{Đây là một tính từ dùng để chỉ tài nguyên không tái tạo.}

\vocabentry{Nutrient}
{/ˈnuːtriənt/ (n)}
{A substance that provides nourishment essential for growth and the maintenance of life.}
{Fertilizers are used to replace the nutrients that crops take from the soil.}
{Đây là một danh từ dùng để chỉ dinh dưỡng trong đất hoặc nước.}

\vocabentry{Offshore}
{/ˌɔːfˈʃɔːr/ (adj)}
{Situated at sea some distance from the shore.}
{Offshore oil drilling is more expensive and risky than onshore drilling.}
{Đây là một tính từ dùng để chỉ các mỏ tài nguyên ngoài khơi.}

\vocabentry{Ore}
{/ɔːr/ (n)}
{A naturally occurring solid material from which a metal or valuable mineral can be extracted.}
{Iron ore is the primary raw material used to make steel.}
{Đây là một danh từ dùng để chỉ quặng.}

\vocabentry{Outcrop}
{/ˈaʊtkrɑːp/ (n)}
{A rock formation that is visible on the surface.}
{Geologists studied the rocky outcrop to find signs of gold deposits.}
{Đây là một danh từ dùng để chỉ phần đá lộ ra trên mặt đất, giúp nhận biết mỏ khoáng.}

\vocabentry{Over-extraction}
{/ˌoʊvər ɪkˈstrækʃn/ (n)}
{The action of taking too much of a resource.}
{Over-extraction of groundwater can cause the land to sink.}
{Đây là một danh từ dùng để chỉ sự khai thác quá mức.}

\vocabentry{Ownership}
{/ˈoʊnərʃɪp/ (n)}
{The act, state, or right of possessing something.}
{The ownership of the mineral rights is currently being disputed in court.}
{Đây là một danh từ dùng để chỉ quyền sở hữu tài nguyên.}

\vocabentry{Petroleum}
{/pəˈtroʊliəm/ (n)}
{A liquid mixture of hydrocarbons that is present in certain rock strata and can be refined to produce fuels.}
{Petroleum is a versatile resource used to make everything from fuel to plastic.}
{Đây là một danh từ dùng để chỉ dầu mỏ.}

\vocabentry{Pipeline}
{/ˈpaɪplaɪn/ (n)}
{A long pipe, typically underground, for conveying oil, gas, etc. over long distances.}
{The pipeline carries natural gas from the fields to the city.}
{Đây là một danh từ dùng để chỉ đường ống dẫn tài nguyên.}

\vocabentry{Potable}
{/ˈpoʊtəbl/ (adj)}
{Safe to drink; drinkable.}
{Access to potable water is a basic human right that many still lack.}
{Đây là một tính từ dùng để chỉ nước có thể uống được.}

\vocabentry{Precious}
{/ˈpreʃəs/ (adj)}
{(Of a distinct object or substance) of great value; not to be wasted or treated carelessly.}
{Gold and diamonds are precious resources that have been valued for centuries.}
{Đây là một tính từ dùng để chỉ các tài nguyên quý hiếm.}

\vocabentry{Preservation}
{/ˌprezərˈveɪʃn/ (n)}
{The action of preserving something.}
{The preservation of ancient forests is vital for protecting rare species.}
{Đây là một danh từ dùng để chỉ sự giữ gìn nguyên trạng tài nguyên.}

\vocabentry{Primary}
{/ˈpraɪmeri/ (adj)}
{Of chief importance; principal.}
{Agriculture is a primary industry that relies heavily on soil and water.}
{Đây là một tính từ dùng để chỉ các tài nguyên sơ cấp hoặc hoạt động khai thác chính.}

\vocabentry{Processing}
{/ˈprɑːsesɪŋ/ (n)}
{Perform a series of mechanical or chemical operations on (something) in order to change or preserve it.}
{The processing of iron ore into steel requires a lot of energy.}
{Đây là một danh từ dùng để chỉ sự chế biến tài nguyên thô.}

\vocabentry{Production}
{/prəˈdʌkʃn/ (n)}
{The action of making or manufacturing from components or raw materials.}
{The country has increased its production of natural gas to meet rising demand.}
{Đây là một danh từ dùng để chỉ sự sản xuất/khai thác tài nguyên.}

\vocabentry{Prospecting}
{/ˈprɑːspektɪŋ/ (n)}
{Search for mineral deposits in a place, especially by means of experimental drilling.}
{Prospecting for oil in the Arctic is a controversial issue.}
{Đây là một danh từ dùng để chỉ sự thăm dò khoáng sản.}

\vocabentry{Purification}
{/ˌpjʊrɪfɪˈkeɪʃn/ (n)}
{The removal of contaminants from something.}
{Water purification is necessary to make river water safe for drinking.}
{Đây là một danh từ dùng để chỉ sự làm sạch hoặc thanh lọc tài nguyên.}

\vocabentry{Quarry}
{/ˈkwɔːri/ (n)}
{A place, typically a large, deep pit, from which stone or other materials are or have been extracted.}
{The old quarry has been turned into a nature reserve.}
{Đây là một danh từ dùng để chỉ mỏ đá.}

\vocabentry{Rare}
{/rer/ (adj)}
{(Of an event, situation, or condition) not occurring very often.}
{Rare earth elements are essential for making high-tech electronics.}
{Đây là một tính từ dùng để chỉ các tài nguyên hiếm.}

\vocabentry{Raw}
{/rɔː/ (adj)}
{(Of a material or substance) in its natural state; not yet processed or refined.}
{The factory imports raw materials like cotton and rubber from overseas.}
{Đây là một tính từ dùng để chỉ nguyên liệu ở trạng thái thô.}

\vocabentry{Refining}
{/rɪˈfaɪnɪŋ/ (n)}
{Remove impurities or unwanted elements from (a substance), typically as part of an industrial process. (Đây là một danh từ dùng để chỉ sự tinh chế (ví dụ: tinh chế dầu).).}
{The refining process turns crude oil into gasoline and other products.}
{}

\vocabentry{Regulation}
{/ˌreɡjuˈleɪʃn/ (n)}
{A rule or directive made and maintained by an authority.}
{Strict government regulations are needed to prevent the over-extraction of water.}
{Đây là một danh từ dùng để chỉ sự quy định hoặc luật lệ về tài nguyên.}

\vocabentry{Replenish}
{/rɪˈplenɪʃ/ (v)}
{Fill (something) up again.}
{It takes time for nature to replenish the groundwater after a long drought.}
{Đây là một động từ dùng để chỉ việc làm đầy lại hoặc tái tạo tài nguyên.}

\vocabentry{Reserve}
{/rɪˈzɜːrv/ (n)}
{A supply of a commodity not needed for immediate use but available if required.}
{The country has large oil reserves that could last for decades.}
{Đây là một danh từ dùng để chỉ trữ lượng tài nguyên.}

\vocabentry{Residue}
{/ˈrezɪduː/ (n)}
{A small amount of something that remains after the main part has gone or been taken or used.}
{The mining process leaves behind a toxic residue that can pollute local water.}
{Đây là một danh từ dùng để chỉ dư lượng hoặc bã rác sau khai thác.}

\vocabentry{Resilient}
{/rɪˈzɪliənt/ (adj)}
{(Of a person or animal) able to withstand or recover quickly from difficult conditions.}
{We need to build a more resilient energy system that can withstand natural disasters.}
{Đây là một tính từ dùng để chỉ khả năng phục hồi của hệ tài nguyên.}

\vocabentry{Resource}
{/ˈriːsɔːrs/ (n)}
{A stock or supply of money, materials, staff, and other assets that can be drawn on by a person or organization in order to function effectively.}
{Time is our most valuable resource, and we should not waste it.}
{Đây là một danh từ dùng để chỉ nguồn lực hoặc tài nguyên.}

\vocabentry{Restoration}
{/ˌrestəˈreɪʃn/ (n)}
{The action of returning something to a former owner, place, or condition.}
{The restoration of the wetlands has helped to improve local water quality.}
{Đây là một danh từ dùng để chỉ sự phục hồi lại tài nguyên hoặc môi trường.}

\vocabentry{Revenue}
{/ˈrevənuː/ (n)}
{Income, especially when of a company or organization and of a substantial nature.}
{The government relies on the revenue from oil exports to fund public services.}
{Đây là một danh từ dùng để chỉ doanh thu từ việc bán tài nguyên.}

\vocabentry{Royalty}
{/ˈrɔɪəlti/ (n)}
{A sum of money paid to a patentee for the use of a patent or to an author for each copy of a book sold.}
{Mining companies must pay a royalty to the government for every ton of coal extracted.}
{Đây là một danh từ dùng để chỉ phí bản quyền hoặc phí khai thác tài nguyên trả cho nhà nước.}

\vocabentry{Run-off}
{/ˈrʌn ɔːf/ (n)}
{The draining away of water from the surface of an area of land.}
{Rainwater run-off from the fields can carry fertilizers into the river.}
{Đây là một danh từ dùng để chỉ dòng chảy tràn, có thể mang theo tài nguyên hoặc ô nhiễm.}

\vocabentry{Salinity}
{/səˈlɪnəti/ (n)}
{The quality or degree of being saline.}
{High soil salinity can make it impossible for crops to grow.}
{Đây là một danh từ dùng để chỉ độ mặn của đất hoặc nước.}

\vocabentry{Scarcity}
{/ˈskersəti/ (n)}
{The state of being scarce or in short supply; shortage.}
{Water scarcity is a major problem in many parts of the world.}
{Đây là một danh từ dùng để chỉ sự khan hiếm tài nguyên.}

\vocabentry{Secular}
{/ˈsekjələr/ (adj)}
{Not connected with religious or spiritual matters.}
{There is a secular trend towards the use of renewable energy.}
{Đây là một tính từ, trong ngữ cảnh này dùng để chỉ các xu hướng tài nguyên dài hạn - "secular trends".}

\vocabentry{Sediment}
{/ˈsedɪmənt/ (n)}
{Matter that settles to the bottom of a liquid.}
{The river carries a lot of sediment, which forms a fertile delta at its mouth.}
{Đây là một danh từ dùng để chỉ trầm tích.}

\vocabentry{Seismic}
{/ˈsaɪzmɪk/ (adj)}
{Relating to earthquakes or other vibrations of the earth and its crust.}
{Seismic surveys are used to find oil and gas deposits deep underground.}
{Đây là một tính từ dùng để chỉ các hoạt động địa chấn dùng trong thăm dò tài nguyên.}

\vocabentry{Sequestration}
{/ˌsiːkwəˈstreɪʃn/ (n)}
{The action of taking legal possession of assets until a debt has been paid or other claims have been met. (Đây là một danh từ dùng để chỉ sự cô lập (ví dụ: cô lập carbon).).}
{Carbon sequestration is a key technology for reducing greenhouse gas emissions.}
{}

\vocabentry{Shortage}
{/ˈʃɔːrtɪdʒ/ (n)}
{A state or situation in which something needed cannot be obtained in sufficient amounts.}
{The city is facing a serious water shortage after a dry winter.}
{Đây là một danh từ dùng để chỉ sự thiếu hụt tài nguyên.}

\vocabentry{Solar}
{/ˈsoʊlər/ (adj)}
{Relating to or determined by the sun.}
{Solar panels are an increasingly popular way to generate electricity.}
{Đây là một tính từ dùng để chỉ những gì thuộc về mặt trời.}

\vocabentry{Source}
{/sɔːrs/ (n)}
{A place, person, or thing from which something comes or can be obtained.}
{The river is the main source of water for the entire city.}
{Đây là một danh từ dùng để chỉ nguồn tài nguyên.}

\vocabentry{Speculation}
{/ˌspekjuˈleɪʃn/ (n)}
{The forming of a theory or conjecture without firm evidence.}
{Speculation in the oil market can cause prices to rise and fall rapidly.}
{Đây là một danh từ dùng để chỉ sự đầu cơ tài nguyên trên thị trường.}

\vocabentry{Stability}
{/stəˈbɪləti/ (n)}
{The state of being stable.}
{Political stability is essential for attracting investment in the mining sector.}
{Đây là một danh từ dùng để chỉ sự ổn định của nguồn cung tài nguyên.}

\vocabentry{Stakeholder}
{/ˈsteɪkhoʊldər/ (n)}
{A person with an interest or concern in something, especially a business.}
{All stakeholders must be consulted before any new mining project begins.}
{Đây là một danh từ dùng để chỉ các bên liên quan trong quản lý tài nguyên.}

\vocabentry{Statute}
{/ˈstætʃuːt/ (n)}
{A written law passed by a legislative body.}
{The protection of natural resources is enshrined in national statute.}
{Đây là một danh từ dùng để chỉ các điều luật về quản lý tài nguyên.}

\vocabentry{Stock}
{/stɑːk/ (n)}
{The goods or merchandise kept on the premises of a business or warehouse and available for sale or distribution.}
{The global stock of fish is being depleted by over-fishing.}
{Đây là một danh từ dùng để chỉ trữ lượng hoặc kho tài nguyên.}

\vocabentry{Subsoil}
{/ˈsʌbsɔɪl/ (n)}
{The soil lying immediately under the surface soil.}
{The subsoil is often rich in minerals but poor in organic matter.}
{Đây là một danh từ dùng để chỉ tầng đất dưới.}

\vocabentry{Subsurface}
{/sʌbˈsɜːrfɪs/ (adj)}
{Situated or occurring below the surface.}
{Subsurface mining is more dangerous and expensive than surface mining.}
{Đây là một tính từ dùng để chỉ các mỏ tài nguyên dưới bề mặt.}

\vocabentry{Sulphur}
{/ˈsʌlfər/ (n)}
{A yellow non-metallic element.}
{Sulphur is used in the production of fertilizers and chemicals.}
{Đây là một danh từ dùng để chỉ lưu huỳnh, một tài nguyên khoáng phi kim.}

\vocabentry{Supply}
{/səˈplaɪ/ (n)}
{A stock of a resource from which a person or place can be provided with the amount needed.}
{The supply of natural gas was cut off due to a technical problem.}
{Đây là một danh từ dùng để chỉ nguồn cung tài nguyên.}

\vocabentry{Surface}
{/ˈsɜːrfɪs/ (n)}
{The outside part or upper layer of something.}
{Open-pit mining involves removing the surface layer of the earth.}
{Đây là một danh từ dùng để chỉ bề mặt Trái Đất nơi khai thác tài nguyên.}

\vocabentry{Surplus}
{/ˈsɜːrpləs/ (n)}
{An amount of something left over when requirements have been met; an excess of production or supply over demand.}
{The country has a surplus of wheat this year and can export it to other nations.}
{Đây là một danh từ dùng để chỉ sự dư thừa tài nguyên.}

\vocabentry{Sustainability}
{/səˌsteɪnəˈbɪləti/ (n)}
{The ability to be maintained at a certain rate or level.}
{Sustainability is the key principle of modern resource management.}
{Đây là một danh từ dùng để chỉ sự bền vững tài nguyên.}

\vocabentry{Symmetry}
{/ˈsɪmətri/ (n)}
{The quality of being made up of exactly similar parts facing each other or around an axis.}
{There is a lack of symmetry in the global distribution of energy resources.}
{Đây là một danh từ, ít dùng trực tiếp nhưng có thể chỉ sự cân đối trong phân bổ tài nguyên.}

\vocabentry{Tailings}
{/ˈteɪlɪŋz/ (n)}
{The residue of an ore that remains after the valuable part has been extracted.}
{Tailings ponds can be a major source of water pollution if they are not managed properly.}
{Đây là một danh từ dùng để chỉ rác thải quặng sau khai thác.}

\vocabentry{Tantalum}
{/ˈtæntələm/ (n)}
{A rare, hard, blue-gray, lustrous transition metal.}
{Tantalum is essential for making the capacitors used in smartphones.}
{Đây là một danh từ dùng để chỉ tantali, một tài nguyên khoáng hiếm cho công nghệ.}

\vocabentry{Tariff}
{/ˈtærɪf/ (n)}
{A tax or duty to be paid on a particular class of imports or exports.}
{The government has imposed a tariff on imported steel to protect the local industry.}
{Đây là một danh từ dùng để chỉ thuế quan đối với tài nguyên xuất nhập khẩu.}

\vocabentry{Technique}
{/tekˈniːk/ (n)}
{A way of carrying out a particular task, especially the execution or performance of an artistic work or a scientific procedure.}
{New drilling techniques have made it possible to extract oil from deeper locations.}
{Đây là một danh từ dùng để chỉ kỹ thuật khai thác.}

\vocabentry{Tectonics}
{/tekˈtɑːnɪks/ (n)}
{The study of the earth's structural features.}
{Understanding plate tectonics is essential for finding mineral deposits.}
{Đây là một danh từ dùng để chỉ các quá trình kiến tạo tạo ra tài nguyên mỏ.}

\vocabentry{Template}
{/ˈtemplet/ (n)}
{A shaped piece of metal, wood, card, plastic, or other material used as a pattern for processes.}
{This project could serve as a template for future resource management programs.}
{Đây là một danh từ, có thể dùng để chỉ khuôn mẫu quản lý tài nguyên.}

\vocabentry{Tenant}
{/ˈtenənt/ (n)}
{A person who occupies land or property rented from a landlord.}
{The mining company is a tenant of the government-owned land.}
{Đây là một danh từ dùng để chỉ người thuê đất khai thác tài nguyên.}

\vocabentry{Tenure}
{/ˈtenjər/ (n)}
{The conditions under which land or buildings are held or occupied.}
{Land tenure systems can have a significant impact on agricultural productivity.}
{Đây là một danh từ dùng để chỉ quyền sử dụng đất hoặc tài nguyên.}

\vocabentry{Terminal}
{/ˈtɜːrmɪnl/ (n)}
{A point where something ends or is distributed.}
{The oil terminal is located at the edge of the harbor.}
{Đây là một danh từ dùng để chỉ cảng hoặc trạm trung chuyển tài nguyên.}

\vocabentry{Terrestrial}
{/təˈrestriəl/ (adj)}
{On or relating to the earth.}
{Terrestrial ecosystems provide us with food, fiber, and fuel.}
{Đây là một tính từ dùng để chỉ các tài nguyên trên cạn.}

\vocabentry{Territory}
{/ˈterətɔːri/ (n)}
{An area of land under the jurisdiction of a ruler or state.}
{The disputed territory is believed to contain large deposits of natural gas.}
{Đây là một danh từ dùng để chỉ lãnh thổ chứa tài nguyên.}

\vocabentry{Thermal}
{/ˈθɜːrml/ (adj)}
{Relating to heat. (Đây là một tính từ dùng để chỉ thuộc về nhiệt (ví dụ: nhiệt năng).).}
{Thermal power plants generate electricity by burning fossil fuels.}
{}

\vocabentry{Threshold}
{/ˈθreʃhoʊld/ (n)}
{The magnitude or intensity that must be exceeded for a certain reaction.}
{We must not exceed the ecological threshold of our water resources.}
{Đây là một danh từ dùng để chỉ ngưỡng khai thác an toàn.}

\vocabentry{Tidal}
{/ˈtaɪdl/ (adj)}
{Relating to or affected by tides.}
{Tidal energy is a predictable and renewable source of power.}
{Đây là một tính từ dùng để chỉ thuộc về thủy triều.}

\vocabentry{Timber}
{/ˈtɪmbər/ (n)}
{Wood prepared for use in building and carpentry.}
{Timber is one of the most important renewable resources in the world.}
{Đây là một danh từ dùng để chỉ gỗ nguyên liệu.}

\vocabentry{Tin}
{/tɪn/ (n)}
{A silvery-white, soft, malleable transition metal.}
{Tin is used to coat other metals to prevent corrosion.}
{Đây là một danh từ dùng để chỉ thiếc, một tài nguyên khoáng kim loại.}

\vocabentry{Topography}
{/təˈpɑːɡrəfi/ (n)}
{The arrangement of the natural and artificial physical features of an area.}
{The rugged topography of the region makes mining difficult and expensive.}
{Đây là một danh từ dùng để chỉ địa hình, ảnh hưởng đến việc khai thác tài nguyên.}

\vocabentry{Topsoil}
{/ˈtɑːpsɔɪl/ (n)}
{The top layer of soil.}
{Erosion is washing away the valuable topsoil, making the land less productive.}
{Đây là một danh từ dùng để chỉ tầng đất mặt màu mỡ.}

\vocabentry{Total}
{/ˈtoʊtl/ (adj)}
{Comprising the whole number or amount.}
{The total amount of fresh water on Earth is limited.}
{Đây là một tính từ dùng để chỉ tổng lượng tài nguyên.}

\vocabentry{Toxic}
{/ˈtɑːksɪk/ (adj)}
{Poisonous.}
{Mining activities can release toxic chemicals into the air and water.}
{Đây là một tính từ dùng để chỉ các chất độc hại phát sinh từ khai thác tài nguyên.}

\vocabentry{Toxin}
{/ˈtɑːksɪn/ (n)}
{A poison of plant or animal origin.}
{Some industrial processes produce toxins that are difficult to remove from the environment.}
{Đây là một danh từ dùng để chỉ độc tố.}

\vocabentry{Trace}
{/treɪs/ (n)}
{A very small amount of something.}
{There are traces of gold in the riverbed, but not enough to be worth mining.}
{Đây là một danh từ dùng để chỉ vết hoặc lượng nhỏ tài nguyên quý.}

\vocabentry{Traditional}
{/trəˈdɪʃənl/ (adj)}
{Existing in or as part of a tradition; long-established. (Đây là một tính từ dùng để chỉ các nguồn tài nguyên truyền thống (than, dầu).).}
{We need to move away from traditional fossil fuels and towards renewable energy.}
{}

\vocabentry{Transit}
{/ˈtrænzɪt/ (n)}
{The carrying of people or things from one place to another.}
{The country earns money from the transit of oil through its pipelines.}
{Đây là một danh từ dùng để chỉ sự vận chuyển tài nguyên.}

\vocabentry{Transmission}
{/trænzˈmɪʃn/ (n)}
{The action or process of transmitting something.}
{High-voltage lines are used for the transmission of electricity over long distances.}
{Đây là một danh từ dùng để chỉ sự truyền tải năng lượng tài nguyên.}

\vocabentry{Transparency}
{/trænsˈpærənsi/ (n)}
{The condition of being transparent.}
{Transparency in the mining sector is essential for preventing corruption.}
{Đây là một danh từ dùng để chỉ sự minh bạch trong quản lý tài nguyên.}

\vocabentry{Transportation}
{/ˌtrænspərˈteɪʃn/ (n)}
{The action of transporting someone or something or the process of being transported.}
{The transportation of coal is a major expense for the power company.}
{Đây là một danh từ dùng để chỉ sự vận chuyển tài nguyên.}

\vocabentry{Treasury}
{/ˈtreʒəri/ (n)}
{The funds or revenue of a government, corporation, or institution.}
{The revenue from mineral exports has boosted the national treasury.}
{Đây là một danh từ dùng để chỉ ngân khố nhà nước nhận tiền từ tài nguyên.}

\vocabentry{Trend}
{/trend/ (n)}
{A general direction in which something is developing or changing.}
{The current trend is towards cleaner and more sustainable energy sources.}
{Đây là một danh từ dùng để chỉ xu hướng tiêu thụ hoặc giá cả tài nguyên.}

\vocabentry{Trial}
{/ˈtraɪəl/ (n)}
{A formal examination of evidence by a judge.}
{The company is conducting trials of a new gold extraction technique.}
{Đây là một danh từ, trong kỹ thuật chỉ sự thử nghiệm khai thác.}

\vocabentry{Tributary}
{/ˈtrɪbjuteri/ (n)}
{A river or stream flowing into a larger river or lake.}
{Pollution in a small tributary can affect the water quality of the main river.}
{Đây là một danh từ dùng để chỉ chi lưu sông, một phần của tài nguyên nước.}

\vocabentry{Tropical}
{/ˈtrɑːpɪkl/ (adj)}
{Of, typical of, or peculiar to the tropics.}
{Tropical rainforests are rich in biodiversity and valuable timber.}
{Đây là một tính từ dùng để chỉ các vùng tài nguyên nhiệt đới.}

\vocabentry{Trough}
{/trɔːf/ (n)}
{A long, narrow open container for animals to eat or drink out of.}
{The price of oil reached a trough during the global economic crisis.}
{Đây là một danh từ, trong kinh tế chỉ mức giá thấp nhất của tài nguyên.}

\vocabentry{Turbine}
{/ˈtɜːrbaɪn/ (n)}
{A machine for producing continuous power in which a wheel or rotor is made to revolve by a fast-moving flow of water, steam, gas, air, or other fluid.}
{Wind turbines are a familiar sight in many coastal areas.}
{Đây là một danh từ dùng để chỉ tua-bin khai thác năng lượng.}

\vocabentry{Turbulent}
{/ˈtɜːrbjələnt/ (adj)}
{Characterized by conflict, disorder, or confusion; not controlled or calm.}
{The global energy market has been very turbulent in recent years.}
{Đây là một tính từ dùng để chỉ dòng chảy rối hoặc sự bất ổn của thị trường tài nguyên.}

\vocabentry{Turf}
{/tɜːrf/ (n)}
{Grass and the surface layer of earth held together by its roots.}
{In some parts of Ireland, turf is still used as a fuel for heating homes.}
{Đây là một danh từ dùng để chỉ lớp đất cỏ hoặc than bùn khai thác.}

\vocabentry{Typical}
{/ˈtɪpɪkl/ (adj)}
{Having the distinctive qualities of a particular type of person or thing.}
{Coal is a typical example of a non-renewable and polluting resource.}
{Đây là một tính từ dùng để chỉ các đặc tính điển hình của một loại tài nguyên.}

\vocabentry{Ubiquitous}
{/juːˈbɪkwɪtəs/ (adj)}
{Present, appearing, or found everywhere.}
{Solar energy is ubiquitous, but it is not always easy to harness effectively.}
{Đây là một tính từ dùng để chỉ những tài nguyên có mặt ở khắp mọi nơi.}

\vocabentry{Ultimate}
{/ˈʌltɪmət/ (adj)}
{Being or happening at the end of a process; final.}
{The ultimate source of all energy on Earth is the sun.}
{Đây là một tính từ dùng để chỉ giới hạn cuối cùng của tài nguyên.}

\vocabentry{Underground}
{/ˌʌndərˈɡraʊnd/ (adj)}
{Beneath the surface of the ground.}
{Underground mining is often more dangerous than open-pit mining.}
{Đây là một tính từ dùng để chỉ các mỏ tài nguyên dưới lòng đất.}

\vocabentry{Underlying}
{/ˌʌndərˈlaɪɪŋ/ (adj)}
{Lying or situated under something.}
{The underlying rock structure determines where mineral deposits are found.}
{Đây là một tính từ dùng để chỉ lớp địa chất nằm dưới.}

\vocabentry{Undertake}
{/ˌʌndərˈteɪk/ (v)}
{Commit oneself to and begin (an enterprise or responsibility); take on.}
{The company plans to undertake a major exploration project in the desert.}
{Đây là một động từ dùng để chỉ việc bắt đầu một dự án khai thác.}

\vocabentry{Uneven}
{/ʌnˈiːvn/ (adj)}
{Not level or smooth.}
{The uneven distribution of oil resources has a huge impact on global politics.}
{Đây là một tính từ dùng để chỉ sự phân bổ không đồng đều.}

\vocabentry{Unfavorable}
{/ʌnˈfeɪvərəbl/ (adj)}
{Expressing or showing a lack of approval or support.}
{Unfavorable weather conditions can disrupt the production of solar energy.}
{Đây là một tính từ dùng để chỉ các điều kiện khai thác không thuận lợi.}

\vocabentry{Uniqueness}
{/juˈniːknəs/ (n)}
{The quality of being particularly remarkable, special, or unusual.}
{The uniqueness of the local ecosystem makes it a valuable natural resource.}
{Đây là một danh từ dùng để chỉ tính độc nhất của một nguồn tài nguyên.}

\vocabentry{Universal}
{/ˌjuːnɪˈvɜːrsl/ (adj)}
{Of, affecting, or done by all people or things in the world or in a particular group; applicable to all cases.}
{Access to clean water is a universal human need.}
{Đây là một tính từ dùng để chỉ các vấn đề tài nguyên mang tính toàn cầu.}

\vocabentry{Unlimited}
{/ʌnˈlɪmɪtɪd/ (adj)}
{Not limited or restricted in terms of number, quantity, or extent.}
{We must move towards energy sources that are virtually unlimited, like solar power.}
{Đây là một tính từ dùng để chỉ các tài nguyên dường như vô hạn.}

\vocabentry{Unprecedented}
{/ʌnˈpresɪdentɪd/ (adj)}
{Never done or known before.}
{The current rate of resource depletion is unprecedented in human history.}
{Đây là một tính từ dùng để chỉ các sự kiện chưa từng có tiền lệ về tài nguyên.}

\vocabentry{Unregulated}
{/ʌnˈreɡjuleɪtɪd/ (adj)}
{Not controlled or supervised by regulations or laws.}
{Unregulated mining can lead to severe environmental damage and social conflict.}
{Đây là một tính từ dùng để chỉ việc khai thác không có sự kiểm soát của pháp luật.}

\vocabentry{Unstable}
{/ʌnˈsteɪbl/ (adj)}
{Prone to change, fail, or give way; not stable.}
{The price of oil is often unstable due to political events in the Middle East.}
{Đây là một tính từ dùng để chỉ sự bất ổn của thị trường hoặc địa chất.}

\vocabentry{Untapped}
{/ˌʌnˈtæpt/ (adj)}
{(Of a resource) not yet exploited or used.}
{The country has vast untapped reserves of natural gas.}
{Đây là một tính từ dùng để chỉ các tài nguyên chưa được khai phá.}

\vocabentry{Upstream}
{/ˌʌpˈstriːm/ (adj)}
{Moving or situated in the direction from which a stream or river flows.}
{Upstream activities in the oil industry include exploration and drilling.}
{Đây là một tính từ dùng để chỉ các hoạt động ở giai đoạn đầu của khai thác tài nguyên.}

\vocabentry{Uranium}
{/juˈreɪniəm/ (n)}
{A dense gray radioactive metal used as a fuel in nuclear reactors.}
{Uranium is a powerful but controversial source of energy.}
{Đây là một danh từ dùng để chỉ urani, tài nguyên năng lượng hạt nhân.}

\vocabentry{Urban}
{/ˈɜːrbən/ (adj)}
{In, relating to, or characteristic of a city or town.}
{Urban areas consume a disproportionate share of the world's resources.}
{Đây là một tính từ dùng để chỉ khu vực đô thị tiêu thụ nhiều tài nguyên.}

\vocabentry{Utility}
{/juːˈtɪləti/ (n)}
{The state of being useful, profitable, or beneficial.}
{The price of utilities like electricity and water is rising in many countries.}
{Đây là một danh từ dùng để chỉ các dịch vụ công ích như điện, nước.}

\vocabentry{Utilization}
{/ˌjuːtələˈzeɪʃn/ (n)}
{The action of making practical and effective use of something.}
{We need to improve our utilization of renewable energy sources.}
{Đây là một danh từ dùng để chỉ sự tận dụng tài nguyên.}

\vocabentry{Valuable}
{/ˈvæljuəbl/ (adj)}
{Worth a great deal of money.}
{Diamond is one of the most valuable minerals in the world.}
{Đây là một tính từ dùng để chỉ các tài nguyên có giá trị cao.}

\vocabentry{Value}
{/ˈvæljuː/ (n)}
{The regard that something is held to deserve; the importance, worth, or usefulness of something.}
{The economic value of our natural resources is huge.}
{Đây là một danh từ dùng để chỉ giá trị kinh tế hoặc sinh thái.}

\vocabentry{Vanish}
{/ˈvænɪʃ/ (v)}
{Disappear suddenly and completely.}
{Many species are vanishing due to the destruction of their natural habitats.}
{Đây là một động từ dùng để chỉ việc tài nguyên biến mất hoặc cạn kiệt.}

\vocabentry{Variable}
{/ˈveriəbl/ (adj)}
{Not consistent or having a fixed pattern; liable to change.}
{The output of wind and solar energy is variable and depends on the weather.}
{Đây là một tính từ dùng để chỉ tính biến động của nguồn cung hoặc giá cả.}

\vocabentry{Vast}
{/væst/ (adj)}
{Of very great extent or quantity; immense.}
{Russia has vast reserves of timber and natural gas.}
{Đây là một tính từ dùng để chỉ lượng tài nguyên khổng lồ.}

\vocabentry{Vegetation}
{/ˌvedʒəˈteɪʃn/ (n)}
{Plants considered collectively, especially those found in a particular area or habitat.}
{The natural vegetation helps to prevent soil erosion and maintain water quality.}
{Đây là một danh từ dùng để chỉ thảm thực vật.}

\vocabentry{Vehicle}
{/ˈviːəkl/ (n)}
{A thing used for transporting people or goods.}
{Electric vehicles are a key part of the transition to a low-carbon economy.}
{Đây là một danh từ, có thể là phương tiện tiêu thụ hoặc vận chuyển tài nguyên.}

\vocabentry{Vein}
{/veɪn/ (n)}
{A distinct sheet-like body of crystallized minerals within a rock.}
{Miners discovered a rich vein of gold running through the rock.}
{Đây là một danh từ dùng để chỉ mạch khoáng sản.}

\vocabentry{Vent}
{/vent/ (n)}
{An opening that allows air, gas, or liquid to pass out of or into a confined space.}
{Geothermal energy is harnessed by drilling into hot underground vents.}
{Đây là một danh từ dùng để chỉ lỗ thông hơi địa nhiệt hoặc núi lửa.}

\vocabentry{Venture}
{/ˈventʃər/ (n)}
{A risky or daring journey or undertaking.}
{The new offshore oil field is a joint venture between two major companies.}
{Đây là một danh từ dùng để chỉ các liên doanh khai thác tài nguyên.}

\vocabentry{Versatile}
{/ˈvɜːrsətl/ (adj)}
{Able to adapt or be adapted to many different functions or activities.}
{Natural gas is a versatile fuel that can be used for heating, cooking, and generating electricity.}
{Đây là một tính từ dùng để chỉ tính đa năng của tài nguyên.}

\vocabentry{Vessel}
{/ˈvesl/ (n)}
{A ship or large boat.}
{The oil was transported to the refinery in a large tanker vessel.}
{Đây là một danh từ dùng để chỉ tàu vận chuyển tài nguyên như dầu hoặc quặng.}

\vocabentry{Vest}
{/vest/ (v)}
{Confer or bestow (power, authority, or property) on someone.}
{The ownership of all minerals is vested in the state.}
{Đây là một động từ, thường nói về quyền sở hữu tài nguyên được trao cho nhà nước.}

\vocabentry{Viability}
{/ˌvaɪəˈbɪləti/ (n)}
{Ability to work successfully.}
{The low price of coal has threatened the viability of many mines.}
{Đây là một danh từ dùng để chỉ tính khả thi kinh tế của một dự án khai thác.}

\vocabentry{Viable}
{/ˈvaɪəbl/ (adj)}
{Capable of working successfully; feasible.}
{Solar energy has become a viable alternative to fossil fuels in many regions.}
{Đây là một tính từ dùng để chỉ tính khả thi.}

\vocabentry{Vibrant}
{/ˈvaɪbrənt/ (adj)}
{Full of energy and life.}
{The country has a vibrant mining sector that attracts international investors.}
{Đây là một tính từ dùng để chỉ thị trường tài nguyên sôi động.}

\vocabentry{Vibration}
{/vaɪˈbreɪʃn/ (n)}
{An instance of vibrating.}
{Seismic vibrations are used to map underground geological structures.}
{Đây là một danh từ dùng trong kỹ thuật thăm dò địa chấn.}

\vocabentry{Vigor}
{/ˈvɪɡər/ (n)}
{Physical strength and good health.}
{The government is promoting the development of the energy sector with great vigor.}
{Đây là một danh từ, có thể chỉ sự phát triển mạnh mẽ của ngành tài nguyên.}

\vocabentry{Virtual}
{/ˈvɜːrtʃuəl/ (adj)}
{Almost or nearly as described, but not completely or according to strict definition.}
{The export of agricultural products is also a virtual export of water.}
{Đây là một tính từ, ví dụ "virtual water" - lượng nước ẩn trong sản phẩm.}

\vocabentry{Visibility}
{/ˌvɪzəˈbɪləti/ (n)}
{The state of being able to see or be seen.}
{Improved visibility in the supply chain helps to ensure that minerals are sourced ethically.}
{Đây là một danh từ dùng để chỉ tính minh bạch hoặc khả năng quan sát mỏ.}

\vocabentry{Vital}
{/ˈvaɪtl/ (adj)}
{Absolutely necessary or important; essential.}
{Natural resources are vital for the economic development of any nation.}
{Đây là một tính từ dùng để chỉ tầm quan trọng sống còn của tài nguyên.}

\vocabentry{Volatility}
{/ˌvɑːləˈtɪləti/ (n)}
{Liability to change rapidly and unpredictably, especially for the worse.}
{Price volatility in the oil market makes it difficult for companies to plan for the future.}
{Đây là một danh từ dùng để chỉ sự biến động mạnh của giá cả tài nguyên.}

\vocabentry{Volume}
{/ˈvɑːljuːm/ (n)}
{The amount of space that a substance or object occupies.}
{The volume of timber harvested each year must be carefully controlled.}
{Đây là một danh từ dùng để chỉ khối lượng hoặc thể tích tài nguyên.}

\vocabentry{Voluntary}
{/ˈvɑːlənteri/ (adj)}
{Done, given, or acting of one's own free will.}
{Many companies have made voluntary commitments to reduce their carbon footprint.}
{Đây là một tính từ dùng để chỉ các cam kết bảo vệ tài nguyên tự nguyện.}

\vocabentry{Vulnerability}
{/ˌvʌlnərəˈbɪləti/ (n)}
{The quality or state of being exposed to the possibility of being attacked or harmed.}
{The country's vulnerability to energy price shocks is a major concern.}
{Đây là một danh từ dùng để chỉ sự dễ bị tổn thương của nguồn cung tài nguyên.}

\vocabentry{Waste}
{/weɪst/ (n)}
{Material, substance, or by-product that is eliminated or discarded as no longer useful or required.}
{Industrial waste must be treated properly to avoid polluting the environment.}
{Đây là một danh từ dùng để chỉ rác thải từ khai thác.}

\vocabentry{Wastewater}
{/ˈweɪstwaːtər/ (n)}
{Water that has been used in the home, in a business, or as part of an industrial process.}
{Wastewater from the mine contains heavy metals that are harmful to fish.}
{Đây là một danh từ dùng để chỉ nước thải.}

\vocabentry{Water}
{/ˈwɔːtər/ (n)}
{A colorless, transparent, odorless, tasteless liquid that forms the seas, lakes, rivers, and rain and is the basis of the fluids of living organisms.}
{Water is our most essential natural resource, and we must not waste it.}
{Đây là một danh từ dùng để chỉ tài nguyên nước.}

\vocabentry{Watercourse}
{/ˈwɔːtərkɔːrs/ (n)}
{A river, canal, or other route for water to flow along.}
{The protection of watercourses from pollution is a priority for the government.}
{Đây là một danh từ dùng để chỉ dòng sông hoặc kênh dẫn nước.}

\vocabentry{Waterfront}
{/ˈwɔːtərfrʌnt/ (n)}
{A part of a town or city that adjoins a river, lake, or sea.}
{The development of the waterfront has increased the value of the local land.}
{Đây là một danh từ dùng để chỉ khu vực ven mặt nước.}

\vocabentry{Watershed}
{/ˈwɔːtərʃed/ (n)}
{An area or ridge of land that separates waters flowing to different rivers, basins, or seas.}
{Protecting the watershed is essential for ensuring a clean and reliable water supply.}
{Đây là một danh từ dùng để chỉ lưu vực sông hoặc bước ngoặt trong quản lý tài nguyên.}

\vocabentry{Wave}
{/weɪv/ (n)}
{A long body of water curling into an arched form and breaking on the shore.}
{Wave energy is a promising but still relatively undeveloped source of renewable power.}
{Đây là một danh từ dùng để chỉ năng lượng sóng biển.}

\vocabentry{Wealth}
{/welθ/ (n)}
{An abundance of valuable possessions or money.}
{The discovery of gold brought great wealth to the region.}
{Đây là một danh từ dùng để chỉ sự giàu có từ tài nguyên.}

\vocabentry{Weapon}
{/ˈwepən/ (n)}
{A thing designed or used for inflicting bodily harm or physical damage.}
{Energy supplies have sometimes been used as a political weapon.}
{Đây là một danh từ, có thể chỉ việc dùng tài nguyên làm "vũ khí" kinh tế.}

\vocabentry{Weather}
{/ˈweðər/ (n)}
{The state of the atmosphere at a place and time.}
{Extreme weather events like droughts can severely affect water resources.}
{Đây là một danh từ dùng để chỉ thời tiết ảnh hưởng đến tài nguyên.}

\vocabentry{Weigh}
{/weɪ/ (v)}
{Find out how heavy (someone or something) is.}
{Every truckload of ore is weighed before it leaves the mine.}
{Đây là một động từ dùng để chỉ việc cân quặng hoặc tài nguyên khai thác.}

\vocabentry{Welfare}
{/ˈwelfer/ (n)}
{The health, happiness, and fortunes of a person or group.}
{The government aims to use the revenue from natural resources to improve public welfare.}
{Đây là một danh từ dùng để chỉ phúc lợi từ việc khai thác tài nguyên.}

\vocabentry{Well}
{/wel/ (n)}
{A shaft sunk into the ground to obtain water, oil, or gas.}
{The company has drilled several new oil wells in the North Sea.}
{Đây là một danh từ dùng để chỉ giếng nước, giếng dầu.}

\vocabentry{Western}
{/ˈwestərn/ (adj)}
{Living in or originating from the West.}
{Western nations have traditionally been the largest consumers of the world's resources.}
{Đây là một tính từ dùng để chỉ các quốc gia phương Tây tiêu thụ nhiều tài nguyên.}

\vocabentry{Wetland}
{/ˈwetlænd/ (n)}
{Land consisting of marshes or swamps; saturated land.}
{Wetlands are important ecosystems that provide habitat for many species.}
{Đây là một danh từ dùng để chỉ vùng đất ngập nước.}

\vocabentry{Wholesale}
{/ˈhoʊlseɪl/ (adj)}
{Done on a large scale; extensive.}
{Wholesale destruction of the rainforest must be stopped.}
{Đây là một tính từ dùng để chỉ việc bán sỉ tài nguyên hoặc khai thác quy mô lớn.}

\vocabentry{Widespread}
{/ˌwaɪdˈspred/ (adj)}
{Found or distributed over a large area or number of people.}
{There is widespread concern about the impact of mining on the environment.}
{Đây là một tính từ dùng để chỉ sự phổ biến của tài nguyên hoặc ô nhiễm.}

\vocabentry{Wild}
{/waɪld/ (adj)}
{(Of an animal or plant) living or growing in the natural environment; not domesticated or cultivated.}
{Wild animals and plants are an important part of our natural heritage.}
{Đây là một tính từ dùng để chỉ các tài nguyên hoang dã.}

\vocabentry{Wildlife}
{/ˈwaɪldlaɪf/ (n)}
{Wild animals collectively; the native fauna of a region.}
{The project aims to protect the local wildlife and their natural habitat.}
{Đây là một danh từ dùng để chỉ động vật hoang dã.}

\vocabentry{Wind}
{/wɪnd/ (n)}
{The perceptible natural movement of the air.}
{Wind power is a clean and increasingly cost-effective way to generate electricity.}
{Đây là một danh từ dùng để chỉ tài nguyên năng lượng gió.}

\vocabentry{Windmill}
{/ˈwɪndmɪl/ (n)}
{A building with sails or vanes that turn in the wind and generate power.}
{Traditional windmills were used for grinding grain or pumping water.}
{Đây là một danh từ dùng để chỉ cối xay gió hoặc tua-bin gió.}

\vocabentry{Wood}
{/wʊd/ (n)}
{The hard fibrous material that forms the main substance of the trunk or branches of a tree or shrub.}
{Wood is a versatile natural resource used for building, furniture, and fuel.}
{Đây là một danh từ dùng để chỉ gỗ, tài nguyên rừng.}

\vocabentry{Workforce}
{/ˈwɜːrkfɔːrs/ (n)}
{The people engaged in or available for work.}
{The mining industry provides employment for a large workforce.}
{Đây là một danh từ dùng để chỉ lực lượng lao động trong ngành tài nguyên.}

\vocabentry{Zinc}
{/zɪŋk/ (n)}
{A silvery-white metal that is a constituent of several alloys.}
{Zinc is used to galvanize steel to prevent it from rusting.}
{Đây là một danh từ dùng để chỉ kẽm, tài nguyên khoáng kim loại.}

