% Auto-generated from raw.txt
% Usage:
%   \vocabentry{word}{/ipa/ (pos)}{definition}{example}{note}

\vocabentry{Automation}
{/ˌɔːtəˈmeɪʃn/ (n)}
{The use of largely automatic equipment in a system of manufacturing or other production process.}
{Automation in the car industry has led to increased efficiency but fewer manual jobs.}
{Đây là một danh từ dùng để chỉ sự tự động hóa, giúp giảm thiểu sự can thiệp của con người.}

\vocabentry{Artificial Intelligence}
{/ˌɑːrtɪfɪʃl ɪnˈtelɪdʒəns/ (n)}
{The theory and development of computer systems able to perform tasks that normally require human intelligence. (Đây là một cụm danh từ dùng để chỉ trí tuệ nhân tạo (AI).).}
{Artificial intelligence is being used to improve medical diagnoses and personalize online shopping.}
{}

\vocabentry{Algorithm}
{/ˈælɡərɪðəm/ (n)}
{A process or set of rules to be followed in calculations or other problem-solving operations, especially by a computer.}
{Social media platforms use complex algorithms to determine which posts you see first.}
{Đây là một danh từ dùng để chỉ thuật toán.}

\vocabentry{Bandwidth}
{/ˈbændwɪtθ/ (n)}
{The transmission capacity of a computer network or other telecommunication system.}
{High-definition video streaming requires a significant amount of bandwidth.}
{Đây là một danh từ dùng để chỉ băng thông, độ rộng của dải tần số để truyền tải dữ liệu.}

\vocabentry{Biotechnology}
{/ˌbaɪoʊtekˈnɑːlədʒi/ (n)}
{The exploitation of biological processes for industrial and other purposes.}
{Biotechnology has made it possible to develop crops that are resistant to pests and diseases.}
{Đây là một danh từ dùng để chỉ công nghệ sinh học.}

\vocabentry{Blockchain}
{/ˈblɑːktʃeɪn/ (n)}
{A system in which a record of transactions made in bitcoin or another cryptocurrency are maintained across several computers.}
{Blockchain technology provides a secure and transparent way to record financial transactions.}
{Đây là một danh từ dùng để chỉ công nghệ chuỗi khối.}

\vocabentry{Cloud computing}
{/ˌklaʊd kəmˈpjuːtɪŋ/ (n)}
{The practice of using a network of remote servers hosted on the internet to store, manage, and process data.}
{Cloud computing allows businesses to access powerful software without having to maintain their own servers.}
{Đây là một cụm danh từ dùng để chỉ điện toán đám mây.}

\vocabentry{Cybersecurity}
{/ˌsaɪbər sɪˈkjʊrəti/ (n)}
{The state of being protected against the criminal or unauthorized use of electronic data.}
{Cybersecurity is a top priority for banks and other financial institutions.}
{Đây là một danh từ dùng để chỉ an ninh mạng.}

\vocabentry{Data mining}
{/ˈdeɪtə ˌmaɪnɪŋ/ (n)}
{The practice of examining large databases in order to generate new information.}
{Companies use data mining to identify patterns in consumer behavior and improve their marketing.}
{Đây là một cụm danh từ dùng để chỉ khai thác dữ liệu.}

\vocabentry{Digitalization}
{/ˌdɪdʒɪtələˈzeɪʃn/ (n)}
{The conversion of text, pictures, or sound into a digital form that can be processed by a computer.}
{The digitalization of medical records has made it easier for doctors to share information.}
{Đây là một danh từ dùng để chỉ sự kỹ thuật số hóa.}

\vocabentry{Encryption}
{/ɪnˈkrɪpʃn/ (n)}
{The process of converting information or data into a code, especially to prevent unauthorized access.}
{End-to-end encryption ensures that only the sender and receiver can read a message.}
{Đây là một danh từ dùng để chỉ sự mã hóa dữ liệu.}

\vocabentry{Gadget}
{/ˈɡædʒɪt/ (n)}
{A small mechanical or electronic device or tool, especially an ingenious or novel one.}
{This kitchen gadget makes it easy to chop vegetables in seconds.}
{Đây là một danh từ dùng để chỉ các thiết bị điện tử nhỏ gọn, tiện ích.}

\vocabentry{Hardware}
{/ˈhɑːrdwer/ (n)}
{The physical tools, machinery, and equipment of a computer system.}
{The computer's hardware includes the monitor, keyboard, and central processing unit.}
{Đây là một danh từ dùng để chỉ phần cứng máy tính.}

\vocabentry{Innovation}
{/ˌɪnəˈveɪʃn/ (n)}
{The action or process of innovating; a new method, idea, product, etc.}
{Technological innovation is the primary driver of economic growth.}
{Đây là một danh từ dùng để chỉ sự đổi mới, cách tân công nghệ.}

\vocabentry{Interface}
{/ˈɪntərfeɪs/ (n)}
{A device or program enabling a user to communicate with a computer.}
{The software has a user-friendly interface that is easy to navigate.}
{Đây là một danh từ dùng để chỉ giao diện người dùng.}

\vocabentry{Internet of Things}
{/ˌɪntərnet əv ˈθɪŋz/ (n)}
{The interconnection via the internet of computing devices embedded in everyday objects. (Đây là một cụm danh từ dùng để chỉ vạn vật kết nối (IoT).).}
{The Internet of Things allows you to control your home's lights and thermostat from your smartphone.}
{}

\vocabentry{Nanotechnology}
{/ˌnænoʊtekˈnɑːlədʒi/ (n)}
{The branch of technology that deals with dimensions and tolerances of less than 100 nanometers.}
{Nanotechnology is being used to develop new materials that are stronger and lighter than steel.}
{Đây là một danh từ dùng để chỉ công nghệ nano.}

\vocabentry{Obsolescence}
{/ˌɑːbsəˈlesns/ (n)}
{The process of becoming obsolete or outdated and no longer used.}
{Planned obsolescence encourages consumers to buy the latest version of a product.}
{Đây là một danh từ dùng để chỉ sự lỗi thời, thường dùng trong cụm "planned obsolescence" - lỗi thời có kế hoạch.}

\vocabentry{Privacy}
{/ˈpraɪvəsi/ (n)}
{The state or condition of being free from being observed or disturbed by other people.}
{Many people are concerned about their online privacy in the age of big data.}
{Đây là một danh từ dùng để chỉ quyền riêng tư trong không gian mạng.}

\vocabentry{Software}
{/ˈsɔːftwer/ (n)}
{The programs and other operating information used by a computer.}
{You need to install the latest software updates to protect your computer from viruses.}
{Đây là một danh từ dùng để chỉ phần mềm máy tính.}

\vocabentry{Telecommunications}
{/ˌtelikəˌmjuːnɪˈkeɪʃnz/ (n)}
{Communication over a distance by cable, telegraph, telephone, or broadcasting.}
{The telecommunications industry has undergone a massive transformation in recent years.}
{Đây là một danh từ dùng để chỉ viễn thông.}

\vocabentry{User-friendly}
{/ˌjuːzər ˈfrendli/ (adj)}
{(Of a machine or system) easy to use or understand.}
{The new operating system is much more user-friendly than the previous one.}
{Đây là một tính từ dùng để chỉ tính dễ sử dụng của một thiết bị hoặc chương trình.}

\vocabentry{Virtual Reality}
{/ˌvɜːrtʃuəl riˈæləti/ (n)}
{The computer-generated simulation of a three-dimensional image or environment. (Đây là một cụm danh từ dùng để chỉ thực tế ảo (VR).).}
{Virtual reality is being used in flight simulators to train pilots.}
{}

\vocabentry{Wireless}
{/ˈwaɪərləs/ (adj)}
{Using radio, microwaves, etc. (as opposed to wires or cables) to transmit signals.}
{Wireless charging is becoming more popular for smartphones.}
{Đây là một tính từ dùng để chỉ công nghệ không dây.}

\vocabentry{Aggregator}
{/ˈæɡrɪɡeɪtər/ (n)}
{A website or program that collects related items of content and displays them or links to them.}
{A news aggregator allows you to see headlines from many different sources in one place.}
{Đây là một danh từ dùng để chỉ bộ tổng hợp thông tin.}

\vocabentry{Analytics}
{/ˌænəˈlɪtɪks/ (n)}
{The systematic computational analysis of data or statistics.}
{Web analytics help businesses understand how people are using their website.}
{Đây là một danh từ dùng để chỉ việc phân tích dữ liệu.}

\vocabentry{App}
{/æp/ (n)}
{An application, especially as downloaded by a user to a mobile device.}
{There is an app for almost everything, from ordering food to tracking your fitness.}
{Đây là một danh từ viết tắt dùng để chỉ ứng dụng di động.}

\vocabentry{Archive}
{/ˈɑːrkaɪv/ (v)}
{To store (data or information) in an archive.}
{You should archive your old emails to keep your inbox organized.}
{Đây là một động từ dùng để chỉ việc lưu trữ dữ liệu lâu dài.}

\vocabentry{Authentication}
{/ɔːˌθentɪˈkeɪʃn/ (n)}
{The process or action of proving or showing something to be true, genuine, or valid.}
{Two-factor authentication provides an extra layer of security for your online accounts.}
{Đây là một danh từ dùng để chỉ sự xác thực người dùng.}

\vocabentry{Avatar}
{/ˈævətɑːr/ (n)}
{An icon or figure representing a particular person in a video game, internet forum, etc.}
{You can customize your avatar with different clothes and accessories.}
{Đây là một danh từ dùng để chỉ hình ảnh đại diện kỹ thuật số.}

\vocabentry{Back up}
{/ˌbæk ˈʌp/ (v)}
{To make a spare copy of (data or a disk).}
{It's important to back up your files regularly in case your computer crashes.}
{Đây là một cụm động từ dùng để chỉ việc sao lưu dữ liệu.}

\vocabentry{Beta}
{/ˈbeɪtə/ (n)}
{A version of a piece of software of a piece of software that is made available for testing.}
{The game is currently in beta, so there might be a few bugs.}
{Đây là một danh từ dùng để chỉ phiên bản thử nghiệm của phần mềm.}

\vocabentry{Biometrics}
{/ˌbaɪoʊˈmetrɪks/ (n)}
{The measurement and statistical analysis of people's unique physical and behavioral characteristics.}
{Biometrics like fingerprint and facial recognition are used to unlock smartphones.}
{Đây là một danh từ dùng để chỉ công nghệ sinh trắc học.}

\vocabentry{Blog}
{/blɔːɡ/ (n)}
{A regularly updated website or web page, typically one run by an individual or small group.}
{She writes a popular travel blog where she shares photos and stories from her trips.}
{Đây là một danh từ dùng để chỉ nhật ký trực tuyến.}

\vocabentry{Bluetooth}
{/ˈbluːtuːθ/ (n)}
{A standard for the short-range wireless interconnection of mobile phones, computers, and other electronic devices.}
{You can connect your phone to your car's speakers using Bluetooth.}
{Đây là một danh từ dùng để chỉ công nghệ kết nối không dây tầm ngắn.}

\vocabentry{Boot}
{/buːt/ (v)}
{To start (a computer) and load its operating system.}
{My computer takes a long time to boot up in the morning.}
{Đây là một động từ dùng để chỉ việc khởi động máy tính.}

\vocabentry{Browser}
{/ˈbraʊzər/ (n)}
{A program with a graphical user interface for displaying HTML files.}
{Google Chrome and Safari are two of the most popular web browsers.}
{Đây là một danh từ dùng để chỉ trình duyệt web.}

\vocabentry{Bug}
{/bʌɡ/ (n)}
{An error, flaw, failure or fault in a computer program or system that causes it to produce an incorrect or unexpected result.}
{The software developers are working hard to fix the bugs in the latest update.}
{Đây là một danh từ dùng để chỉ lỗi phần mềm.}

\vocabentry{Cache}
{/kæʃ/ (n)}
{An auxiliary memory from which high-speed retrieval is possible.}
{Clearing your browser's cache can sometimes help fix loading issues.}
{Đây là một danh từ dùng để chỉ bộ nhớ đệm.}

\vocabentry{Chatbot}
{/ˈtʃætbɑːt/ (n)}
{A computer program designed to simulate conversation with human users, especially over the internet.}
{Many websites use chatbots to provide instant customer support.}
{Đây là một danh từ dùng để chỉ chương trình trò chuyện tự động.}

\vocabentry{Chip}
{/tʃɪp/ (n)}
{A tiny wafer of semiconducting material used to make an integrated circuit.}
{The new processor chip is much faster and more efficient than the old one.}
{Đây là một danh từ dùng để chỉ chip bán dẫn/vi mạch.}

\vocabentry{Coder}
{/ˈkoʊdər/ (n)}
{A person who writes computer code; a programmer.}
{He works as a coder for a large tech company in Silicon Valley.}
{Đây là một danh từ dùng để chỉ người lập trình/viết mã.}

\vocabentry{Compatibility}
{/kəmˌpætəˈbɪləti/ (n)}
{The ability of two systems to coexist or cooperate without modification.}
{You should check the software's compatibility with your operating system before installing it.}
{Đây là một danh từ dùng để chỉ tính tương thích.}

\vocabentry{Configuration}
{/kənˌfɪɡjəˈreɪʃn/ (n)}
{An arrangement of parts or elements in a particular form, figure, or combination.}
{The technician helped me with the initial configuration of my new router.}
{Đây là một danh từ dùng để chỉ sự cấu hình hệ thống.}

\vocabentry{Connectivity}
{/ˌkɑːnekˈtɪvəti/ (n)}
{The state or quality of being connected or connective.}
{The hotel provides free Wi-Fi connectivity in all guest rooms.}
{Đây là một danh từ dùng để chỉ khả năng kết nối mạng.}

\vocabentry{Cookie}
{/ˈkʊki/ (n)}
{A packet of data sent by an internet server to a browser, which is returned by the browser each time it subsequently accesses the same server.}
{Some websites use cookies to remember your login information and preferences.}
{Đây là một danh từ dùng để chỉ tệp lưu thông tin trình duyệt.}

\vocabentry{CPU}
{/ˌsiː piː ˈjuː/ (n)}
{Central Processing Unit; the primary component of a computer that acts as its "control center."}
{The CPU is often referred to as the brain of the computer.}
{Đây là một danh từ viết tắt dùng để chỉ bộ vi xử lý trung tâm.}

\vocabentry{Cracker}
{/ˈkrækər/ (n)}
{A person who breaks into a computer system, especially for criminal purposes.}
{Crackers can steal sensitive information like credit card numbers and passwords.}
{Đây là một danh từ dùng để chỉ kẻ bẻ khóa máy tính với mục đích xấu.}

\vocabentry{Cryptocurrency}
{/ˈkrɪptoʊkɜːrənsi/ (n)}
{A digital currency in which transactions are verified and records maintained by a decentralized system using cryptography.}
{Bitcoin is the most well-known cryptocurrency in the world.}
{Đây là một danh từ dùng để chỉ tiền mã hóa.}

\vocabentry{Dashboard}
{/ˈdæʃbɔːrd/ (n)}
{A graphical user interface which provides at-a-glance views of key performance indicators.}
{The company's sales dashboard allows managers to track performance in real-time.}
{Đây là một danh từ dùng để chỉ bảng điều khiển kỹ thuật số.}

\vocabentry{Database}
{/ˈdeɪtəbeɪs/ (n)}
{A structured set of data held in a computer, especially one that is accessible in various ways.}
{The customer database contains names, addresses, and purchase history.}
{Đây là một danh từ dùng để chỉ cơ sở dữ liệu.}

\vocabentry{Debug}
{/ˌdiːˈbʌɡ/ (v)}
{To identify and remove errors from computer hardware or software.}
{It took the programmers several days to debug the new software.}
{Đây là một động từ dùng để chỉ việc tìm và sửa lỗi chương trình.}

\vocabentry{Decryption}
{/diːˈkrɪpʃn/ (n)}
{The process of taking encoded or encrypted text and converting it back into text that you or the computer can read.}
{You need a special key for the decryption of the sensitive files.}
{Đây là một danh từ dùng để chỉ sự giải mã.}

\vocabentry{Default}
{/dɪˈfɔːlt/ (n)}
{A preselected option adopted by a computer program or other mechanism when no alternative is specified.}
{The default font for this program is Arial.}
{Đây là một danh từ dùng để chỉ giá trị/thiết lập mặc định.}

\vocabentry{Desktop}
{/ˈdesktɑːp/ (n)}
{The working area of a computer screen.}
{You can keep shortcuts to your most frequently used programs on your desktop.}
{Đây là một danh từ dùng để chỉ màn hình nền máy tính hoặc máy tính để bàn.}

\vocabentry{Device}
{/dɪˈvaɪs/ (n)}
{A thing made or adapted for a particular purpose, especially a piece of mechanical or electronic equipment.}
{Smartphones and tablets are examples of mobile devices.}
{Đây là một danh từ dùng để chỉ thiết bị điện tử.}

\vocabentry{Digital}
{/ˈdɪdʒɪtl/ (adj)}
{Relating to or using signals or information represented by discrete values (usually digits).}
{We live in a digital world where everything is connected.}
{Đây là một tính từ dùng để chỉ kỹ thuật số.}

\vocabentry{Directory}
{/dəˈrektəri/ (n)}
{A file that consists solely of a set of other files.}
{You can organize your files into different directories to make them easier to find.}
{Đây là một danh từ dùng để chỉ thư mục chứa tệp tin.}

\vocabentry{Domain}
{/doʊˈmeɪn/ (n)}
{A distinct subset of the internet with addresses sharing a common suffix.}
{The company's domain name is [www.example.com](https://www.example.com).}
{Đây là một danh từ dùng để chỉ tên miền website.}

\vocabentry{Download}
{/ˌdaʊnˈloʊd/ (v)}
{To copy (data) from one computer system to another, typically over the internet.}
{You can download the latest version of the app from the App Store.}
{Đây là một động từ dùng để chỉ việc tải dữ liệu xuống.}

\vocabentry{Drag and drop}
{/ˌdræɡ ənd ˈdrɑːp/ (v)}
{To move (an icon or other image) to another part of the screen using a mouse.}
{You can easily organize your files by dragging and dropping them into different folders.}
{Đây là một cụm động từ dùng để chỉ việc kéo và thả chuột.}

\vocabentry{Driver}
{/ˈdraɪvər/ (n)}
{A program that controls the operation of a particular type of device attached to a computer.}
{You need to install the correct printer driver before you can print your documents.}
{Đây là một danh từ dùng để chỉ chương trình điều khiển thiết bị.}

\vocabentry{E-commerce}
{/ˈiː kɑːmɜːrs/ (n)}
{Commercial transactions conducted electronically on the internet.}
{E-commerce has grown significantly over the last decade.}
{Đây là một danh từ dùng để chỉ thương mại điện tử.}

\vocabentry{Environment}
{/ɪnˈvaɪrənmənt/ (n)}
{The surroundings or conditions in which a person, animal, or plant lives or operates; in technology, the total of all software and hardware.}
{The new software was tested in a virtual environment before being released.}
{Đây là một danh từ, trong công nghệ dùng để chỉ môi trường hệ thống.}

\vocabentry{Ethernet}
{/ˈiːθərnet/ (n)}
{A system for connecting a number of computer systems to form a local area network (LAN).}
{An Ethernet cable provides a faster and more stable internet connection than Wi-Fi.}
{Đây là một danh từ dùng để chỉ mạng cục bộ có dây.}

\vocabentry{Executable}
{/ɪɡˈzekjətəbl/ (adj)}
{(Of a file) containing a program that can be run by a computer.}
{Be careful when opening executable files from unknown sources.}
{Đây là một tính từ dùng để chỉ tệp tin có thể thực thi/chạy được.}

\vocabentry{Export}
{/ɪkˈspɔːrt/ (v)}
{To transfer (data) in a format that can be used by other programs.}
{You can export your contact list as a CSV file.}
{Đây là một động từ dùng để chỉ việc xuất dữ liệu.}

\vocabentry{Fiber optics}
{/ˌfaɪbər ˈɑːptɪks/ (n)}
{The use of thin flexible fibers of glass or other transparent solids to transmit light signals.}
{Fiber optics allow for extremely high-speed internet connections.}
{Đây là một danh từ dùng để chỉ cáp quang.}

\vocabentry{File}
{/faɪl/ (n)}
{A collection of data, programs, etc. stored in a computer's memory or on a storage device.}
{You should give your files clear and descriptive names.}
{Đây là một danh từ dùng để chỉ tệp tin.}

\vocabentry{Firewall}
{/ˈfaɪərwɔːl/ (n)}
{A part of a computer system or network which is designed to block unauthorized access while permitting outward communication.}
{A firewall helps protect your computer from hackers and viruses.}
{Đây là một danh từ dùng để chỉ tường lửa bảo mật.}

\vocabentry{Flash drive}
{/ˈflæʃ draɪv/ (n)}
{A small electronic device containing flash memory that is used for storing data or transferring it to or from a computer.}
{I always keep a backup of my important files on a flash drive.}
{Đây là một danh từ dùng để chỉ ổ nhớ USB.}

\vocabentry{Folder}
{/ˈfoʊldər/ (n)}
{An icon on a computer screen that can be used to store a group of related files together.}
{You can create a new folder for each of your projects.}
{Đây là một danh từ dùng để chỉ thư mục máy tính.}

\vocabentry{Font}
{/fɑːnt/ (n)}
{A set of type of a particular face and size.}
{You can change the font and size of your text to make it more readable.}
{Đây là một danh từ dùng để chỉ phông chữ.}

\vocabentry{Format}
{/ˈfɔːrmæt/ (v)}
{To arrange or put into a format.}
{You need to format a new hard drive before you can use it to store data.}
{Đây là một động từ dùng để chỉ việc định dạng văn bản hoặc đĩa.}

\vocabentry{Freeware}
{/ˈfriːwer/ (n)}
{Software that is available free of charge.}
{Many useful programs are available as freeware on the internet.}
{Đây là một danh từ dùng để chỉ phần mềm miễn phí.}

\vocabentry{Full-stack}
{/ˌfʊl ˈstæk/ (adj)}
{Relating to or being a computer programmer who is able to work with both the front end and back end of a website or application. (Đây là một tính từ dùng để chỉ lập trình viên toàn năng (làm được cả giao diện và hệ thống).).}
{The company is looking to hire a full-stack developer.}
{}

\vocabentry{Gateway}
{/ˈɡeɪtweɪ/ (n)}
{A device used to connect two different networks.}
{The router acts as a gateway between your local network and the internet.}
{Đây là một danh từ dùng để chỉ cổng kết nối mạng.}

\vocabentry{Gigabyte}
{/ˈɡɪɡəbaɪt/ (n)}
{A unit of information equal to one billion bytes. (Đây là một danh từ dùng để chỉ đơn vị đo dung lượng bộ nhớ (GB).).}
{My new smartphone has 128 gigabytes of storage space.}
{}

\vocabentry{Glitch}
{/ɡlɪtʃ/ (n)}
{A sudden, usually temporary malfunction or fault of equipment.}
{A minor glitch in the system caused the website to crash for a few minutes.}
{Đây là một danh từ dùng để chỉ sự cố kỹ thuật nhỏ/tạm thời.}

\vocabentry{Graphic}
{/ˈɡræfɪk/ (adj)}
{Relating to visual art, especially drawing, design, and lettering.}
{The new game has amazing graphic quality.}
{Đây là một tính từ dùng để chỉ đồ họa.}

\vocabentry{Hacker}
{/ˈhækər/ (n)}
{A person who uses computers to gain unauthorized access to data.}
{Hackers can gain access to your personal information if you don't use strong passwords.}
{Đây là một danh từ dùng để chỉ tin tặc.}

\vocabentry{Hard drive}
{/ˈhɑːrd draɪv/ (n)}
{A high-capacity, self-contained storage device containing a read-write mechanism plus one or more hard disks.}
{My hard drive is almost full, so I need to delete some old files.}
{Đây là một danh từ dùng để chỉ ổ cứng máy tính.}

\vocabentry{Hashtag}
{/ˈhæʃtæɡ/ (n)}
{A word or phrase preceded by a hash sign (\#), used on social media websites and apps to identify digital content on a specific topic.}
{You can use hashtags to make your posts easier to find.}
{Đây là một danh từ dùng để chỉ thẻ gắn chủ đề trên mạng xã hội.}

\vocabentry{Header}
{/ˈhedər/ (n)}
{A line or block of text appearing at the top of each page of a book or document.}
{You can put your name and the date in the header of your document.}
{Đây là một danh từ dùng để chỉ phần đầu trang.}

\vocabentry{Hi-tech}
{/ˌhaɪ ˈtek/ (adj)}
{High technology; very modern technology.}
{The new hospital is equipped with the latest hi-tech medical equipment.}
{Đây là một tính từ dùng để chỉ công nghệ cao.}

\vocabentry{Homepage}
{/ˈhoʊmpeɪdʒ/ (n)}
{The introductory page of a website.}
{You can find a link to the contact page on the homepage.}
{Đây là một danh từ dùng để chỉ trang chủ của một website.}

\vocabentry{Host}
{/hoʊst/ (v)}
{To store (a website or other data) on a server so that it can be accessed over the internet.}
{The company hosts its website on a secure server.}
{Đây là một động từ dùng để chỉ việc lưu trữ/cung cấp dịch vụ máy chủ.}

\vocabentry{Hotspot}
{/ˈhɑːtspɑːt/ (n)}
{A public place where a wireless signal is made available so that people can connect to the internet.}
{You can find a Wi-Fi hotspot in most cafes and libraries.}
{Đây là một danh từ dùng để chỉ điểm phát Wi-Fi công cộng.}

\vocabentry{HTML}
{/ˌeɪtʃ tiː em ˈel/ (n)}
{HyperText Markup Language; the standard markup language for documents designed to be displayed in a web browser.}
{You need to know HTML to create your own website.}
{Đây là một danh từ viết tắt dùng để chỉ ngôn ngữ đánh dấu siêu văn bản để tạo website.}

\vocabentry{Icon}
{/ˈaɪkɑːn/ (n)}
{A small symbol or picture on a computer screen that represents a program or file.}
{You can open a program by clicking on its icon.}
{Đây là một danh từ dùng để chỉ biểu tượng trên màn hình.}

\vocabentry{Import}
{/ɪmˈpɔːrt/ (v)}
{To bring (data) into one program from another.}
{You can import your photos from your phone to your computer.}
{Đây là một động từ dùng để chỉ việc nhập dữ liệu vào một chương trình.}

\vocabentry{Inbox}
{/ˈɪnbɑːks/ (n)}
{An electronic folder in which emails received by an individual are held.}
{I have several unread emails in my inbox.}
{Đây là một danh từ dùng để chỉ hộp thư đến.}

\vocabentry{Input}
{/ˈɪnpʊt/ (n)}
{Data or signals entered into a computer.}
{You can use the keyboard and mouse to provide input to the computer.}
{Đây là một danh từ dùng để chỉ dữ liệu đầu vào.}

\vocabentry{Install}
{/ɪnˈstɔːl/ (v)}
{To place (software) in a computer so that it is ready to be used.}
{You need to install the software before you can use it.}
{Đây là một động từ dùng để chỉ việc cài đặt phần mềm.}

\vocabentry{Interactive}
{/ˌɪntərˈæktɪv/ (adj)}
{(Of a computer or program) allowing a two-way flow of information between it and a user.}
{The website has several interactive features, such as quizzes and games.}
{Đây là một tính từ dùng để chỉ tính tương tác.}

\vocabentry{Intranet}
{/ˈɪntrənet/ (n)}
{A local or restricted communications network, especially a private network created using World Wide Web software.}
{Employees can access company news and documents on the intranet.}
{Đây là một danh từ dùng để chỉ mạng nội bộ.}

\vocabentry{IP address}
{/ˌaɪ ˈpiː ədres/ (n)}
{Internet Protocol address; a unique string of numbers separated by periods that identifies each computer using the Internet Protocol to communicate over a network.}
{Every device connected to the internet has its own IP address.}
{Đây là một cụm danh từ dùng để chỉ địa chỉ giao thức internet.}

\vocabentry{IT}
{/ˌaɪ ˈtiː/ (n)}
{Information Technology; the study or use of systems (especially computers and telecommunications) for storing, retrieving, and sending information.}
{He works in the IT department of a large company.}
{Đây là một danh từ viết tắt dùng để chỉ công nghệ thông tin.}

\vocabentry{Java}
{/ˈdʒɑːvə/ (n)}
{A high-level, class-based, object-oriented programming language.}
{Many websites and apps are built using Java.}
{Đây là một danh từ dùng để chỉ một ngôn ngữ lập trình phổ biến.}

\vocabentry{Keyboard}
{/ˈkiːbɔːrd/ (n)}
{A set of keys used to type text and enter commands into a computer.}
{I prefer using a physical keyboard to a touchscreen.}
{Đây là một danh từ dùng để chỉ bàn phím máy tính.}

\vocabentry{Keyword}
{/ˈkiːwɜːrd/ (n)}
{A word or concept of great significance; in technology, a word used to find information in a search engine.}
{You should use specific keywords to get the best search results.}
{Đây là một danh từ dùng để chỉ từ khóa tìm kiếm.}

\vocabentry{Laptop}
{/ˈlæptɑːp/ (n)}
{A computer that is portable and suitable for use while traveling.}
{I take my laptop with me whenever I go to a coffee shop.}
{Đây là một danh từ dùng để chỉ máy tính xách tay.}

\vocabentry{Link}
{/lɪŋk/ (n)}
{A reference to data that the reader can directly follow by clicking or hovering.}
{You can click on the link to go to the company's website.}
{Đây là một danh từ dùng để chỉ đường liên kết web.}

\vocabentry{Load}
{/loʊd/ (v)}
{To transfer (a program or data) into a computer's memory.}
{The webpage is taking a long time to load.}
{Đây là một động từ dùng để chỉ việc tải chương trình hoặc dữ liệu.}

\vocabentry{Log in}
{/ˌlɔːɡ ˈɪn/ (v)}
{To enter a username and password to gain access to a computer system or website.}
{You need to log in to your account to see your messages.}
{Đây là một cụm động từ dùng để chỉ việc đăng nhập.}

\vocabentry{Log out}
{/ˌlɔːɡ ˈaʊt/ (v)}
{To end a session on a computer system or website.}
{It's important to log out of your accounts when using a public computer.}
{Đây là một cụm động từ dùng để chỉ việc đăng xuất.}

\vocabentry{Malware}
{/ˈmælwer/ (n)}
{Software which is specifically designed to disrupt, damage, or gain unauthorized access to a computer system.}
{You can protect your computer from malware by installing antivirus software.}
{Đây là một danh từ dùng để chỉ phần mềm độc hại.}

\vocabentry{Menu}
{/ˈmenjuː/ (n)}
{A list of commands or options, especially one displayed on a screen.}
{You can find the print command in the file menu.}
{Đây là một danh từ dùng để chỉ thực đơn lựa chọn trên màn hình.}

\vocabentry{Microchip}
{/ˈmaɪkroʊtʃɪp/ (n)}
{A tiny wafer of semiconducting material used to make an integrated circuit.}
{Microchips are used in almost all electronic devices, from computers to cars.}
{Đây là một danh từ dùng để chỉ vi mạch/chíp siêu nhỏ.}

\vocabentry{Modem}
{/ˈmoʊdəm/ (n)}
{A combined device for modulation and demodulation, for example, between the digital data of a computer and the analog signal of a telephone line.}
{You need a modem to connect your computer to the internet.}
{Đây là một danh từ dùng để chỉ thiết bị điều chế và giải điều chế mạng.}

\vocabentry{Monitor}
{/ˈmɑːnɪtər/ (n)}
{A screen used to display visual output from a computer.}
{I have two monitors so I can work on multiple tasks at the same time.}
{Đây là một danh từ dùng để chỉ màn hình máy tính.}

\vocabentry{Motherboard}
{/ˈmʌðərbɔːrd/ (n)}
{A printed circuit board containing the principal components of a computer or other device.}
{The motherboard is the main circuit board of the computer.}
{Đây là một danh từ dùng để chỉ bảng mạch chủ.}

\vocabentry{Mouse}
{/maʊs/ (n)}
{A small handheld device that is used to control the cursor on a computer screen.}
{You can click, double-click, and drag with the mouse.}
{Đây là một danh từ dùng để chỉ chuột máy tính.}

\vocabentry{Multitasking}
{/ˌmʌltiˈtæskɪŋ/ (n)}
{(Of a computer) carrying out more than one program or task at the same time.}
{Modern operating systems are very good at multitasking.}
{Đây là một danh từ dùng để chỉ khả năng đa nhiệm.}

\vocabentry{Network}
{/ˈnetwɜːrk/ (n)}
{A group of two or more computer systems linked together.}
{The office has a local network that allows employees to share files and printers.}
{Đây là một danh từ dùng để chỉ mạng lưới máy tính.}

\vocabentry{Offline}
{/ˌɔːfˈlaɪn/ (adj)}
{Not connected to a computer or the internet.}
{You can work on your document offline and upload it later.}
{Đây là một tính từ dùng để chỉ trạng thái ngoại tuyến.}

\vocabentry{Online}
{/ˌɔːnˈlaɪn/ (adj)}
{Connected to a computer or the internet.}
{You can find almost any information you need online.}
{Đây là một tính từ dùng để chỉ trạng thái trực tuyến.}

\vocabentry{Open source}
{/ˌoʊpən ˈsɔːrs/ (adj)}
{Relating to software for which the original source code is made freely available and may be redistributed and modified.}
{Linux is a popular open source operating system.}
{Đây là một tính từ dùng để chỉ phần mềm mã nguồn mở.}

\vocabentry{Operating system}
{/ˈɑːpəreɪtɪŋ ˌsɪstəm/ (n)}
{The software that supports a computer's basic functions, such as scheduling tasks, executing applications, and controlling peripherals. (Đây là một cụm danh từ dùng để chỉ hệ điều hành (Windows, macOS, v.v.).).}
{Windows 11 is the latest operating system from Microsoft.}
{}

\vocabentry{Output}
{/ˈaʊtpʊt/ (n)}
{Data or signals sent from a computer.}
{The monitor and printer are examples of output devices.}
{Đây là một danh từ dùng để chỉ dữ liệu đầu ra.}

\vocabentry{Password}
{/ˈpæswɜːrd/ (n)}
{A secret word or phrase that is used to gain access to a computer system or website.}
{You should use a strong password that is difficult for others to guess.}
{Đây là một danh từ dùng để chỉ mật khẩu.}

\vocabentry{Paste}
{/peɪst/ (v)}
{To insert (copied or cut text or an image) into a document.}
{You can copy text from one document and paste it into another.}
{Đây là một động từ dùng để chỉ việc dán dữ liệu.}

\vocabentry{PDF}
{/ˌpiː diː ˈef/ (n)}
{Portable Document Format; a file format that provides an electronic image of text or text and graphics that looks like a printed document.}
{You can save your document as a PDF to ensure that it looks the same on all devices.}
{Đây là một danh từ viết tắt dùng để chỉ định dạng tệp tài liệu di động.}

\vocabentry{Peripheral}
{/pəˈrɪfərəl/ (n)}
{A device that is connected to a computer but is not part of the main computer system.}
{Printers, scanners, and external hard drives are examples of peripherals.}
{Đây là một danh từ dùng để chỉ thiết bị ngoại vi.}

\vocabentry{Phishing}
{/ˈfɪʃɪŋ/ (n)}
{The fraudulent practice of sending emails purporting to be from reputable companies in order to induce individuals to reveal personal information.}
{Be careful of phishing emails that ask for your bank details.}
{Đây là một danh từ dùng để chỉ hành vi lừa đảo qua mạng để lấy thông tin cá nhân.}

\vocabentry{Pixel}
{/ˈpɪksl/ (n)}
{A minute area of illumination on a display screen, one of many from which an image is composed.}
{The more pixels a screen has, the sharper the image will be.}
{Đây là một danh từ dùng để chỉ điểm ảnh.}

\vocabentry{Platform}
{/ˈplætfɔːrm/ (n)}
{A standard for a computer system's hardware and software.}
{The app is available on multiple platforms, including iOS and Android.}
{Đây là một danh từ dùng để chỉ nền tảng công nghệ.}

\vocabentry{Plugin}
{/ˈplʌɡɪn/ (n)}
{A software component that adds a specific feature to an existing computer program.}
{You can install a plugin to add more features to your web browser.}
{Đây là một danh từ dùng để chỉ phần mềm bổ trợ.}

\vocabentry{Pop-up}
{/ˈpɑːp ʌp/ (n)}
{A graphical user interface area, usually a small window, that suddenly appears in the foreground of the visual interface.}
{Some websites use pop-ups to display advertisements.}
{Đây là một danh từ dùng để chỉ cửa sổ bật lên tự động.}

\vocabentry{Portal}
{/ˈpɔːrtl/ (n)}
{A website or service that offers a broad array of resources and services, such as email, forums, search engines, and online shopping malls.}
{The university has a student portal where you can check your grades and schedule.}
{Đây là một danh từ dùng để chỉ cổng thông tin điện tử.}

\vocabentry{Post}
{/poʊst/ (v)}
{To publish (content) on a website or social media platform.}
{She posts updates about her life on social media every day.}
{Đây là một động từ dùng để chỉ việc đăng bài viết.}

\vocabentry{Processor}
{/ˈprɑːsesər/ (n)}
{A machine that processes something; in technology, the central processing unit (CPU).}
{The new laptop has a very fast processor.}
{Đây là một danh từ dùng để chỉ bộ vi xử lý.}

\vocabentry{Program}
{/ˈproʊɡræm/ (n)}
{A set of instructions that a computer uses to perform a specific task.}
{You can use a word processing program to write an essay.}
{Đây là một danh từ dùng để chỉ chương trình máy tính.}

\vocabentry{Programmer}
{/ˈproʊɡræmər/ (n)}
{A person who writes computer programs.}
{She is a very talented programmer who knows several different languages.}
{Đây là một danh từ dùng để chỉ lập trình viên.}

\vocabentry{Protocol}
{/ˈproʊtəkɑːl/ (n)}
{A set of rules governing the exchange or transmission of data between devices.}
{HTTP is the protocol used for transmitting web pages over the internet.}
{Đây là một danh từ dùng để chỉ giao thức mạng.}

\vocabentry{RAM}
{/ræm/ (n)}
{Random Access Memory; a type of computer memory that can be accessed randomly.}
{My computer needs more RAM to run heavy software smoothly.}
{Đây là một danh từ viết tắt dùng để chỉ bộ nhớ truy cập ngẫu nhiên.}

\vocabentry{Resolution}
{/ˌrezəˈluːʃn/ (n)}
{The number of pixels per unit of area in an image or on a display screen.}
{The new monitor has a very high resolution.}
{Đây là một danh từ dùng để chỉ độ phân giải màn hình.}

\vocabentry{Router}
{/ˈraʊtər/ (n)}
{A device which forwards data packets between computer networks.}
{The router is connected to the modem to provide Wi-Fi to the whole house.}
{Đây là một danh từ dùng để chỉ bộ định tuyến mạng.}

\vocabentry{Save}
{/seɪv/ (v)}
{To store (data) in a computer's memory or on a storage device.}
{Don't forget to save your work before you close the program.}
{Đây là một động từ dùng để chỉ việc lưu dữ liệu.}

\vocabentry{Scanner}
{/ˈskænər/ (n)}
{A device that scans documents and converts them into digital data.}
{You can use a scanner to make a digital copy of your old photos.}
{Đây là một danh từ dùng để chỉ máy quét tài liệu.}

\vocabentry{Screenshot}
{/ˈskriːnʃɑːt/ (n)}
{An image of the data displayed on the screen of a computer or mobile device.}
{You can take a screenshot to show someone what is on your screen.}
{Đây là một danh từ dùng để chỉ ảnh chụp màn hình.}

\vocabentry{Search engine}
{/ˈsɜːrtʃ ˌendʒɪn/ (n)}
{A program that searches for and identifies items in a database that correspond to keywords or characters specified by the user.}
{Google is the most popular search engine in the world.}
{Đây là một cụm danh từ dùng để chỉ công cụ tìm kiếm.}

\vocabentry{Security}
{/sɪˈkjʊrəti/ (n)}
{The state of being free from danger or threat.}
{You should take security precautions to protect your personal information online.}
{Đây là một danh từ dùng để chỉ sự bảo mật.}

\vocabentry{Server}
{/ˈsɜːrvər/ (n)}
{A computer or computer program which manages access to a centralized resource or service in a network.}
{The company's website is hosted on a secure server.}
{Đây là một danh từ dùng để chỉ máy chủ.}

\vocabentry{Settings}
{/ˈsetɪŋz/ (n)}
{The options available in a program that allow you to customize it to your liking.}
{You can change the language and theme of the app in the settings.}
{Đây là một danh từ dùng để chỉ các thiết lập/cài đặt.}

\vocabentry{Shareware}
{/ˈʃerwer/ (n)}
{Software that is available free of charge for a trial period, after which a fee is often required.}
{You can download a shareware version of the program to see if you like it.}
{Đây là một danh từ dùng để chỉ phần mềm dùng thử.}

\vocabentry{Shortcut}
{/ˈʃɔːrtkʌt/ (n)}
{An icon that provides quick access to a program or file.}
{You can create a shortcut on your desktop for your favorite game.}
{Đây là một danh từ dùng để chỉ biểu tượng truy cập nhanh.}

\vocabentry{Sign in}
{/ˌsaɪn ˈɪn/ (v)}
{To log in to a computer system or website.}
{You need to sign in to your account to use the service.}
{Đây là một cụm động từ dùng để chỉ việc đăng nhập.}

\vocabentry{Smartphone}
{/ˈsmɑːrtfoʊn/ (n)}
{A mobile phone that performs many of the functions of a computer.}
{I use my smartphone for almost everything, from checking my email to taking photos.}
{Đây là một danh từ dùng để chỉ điện thoại thông minh.}

\vocabentry{Social media}
{/ˌsoʊʃl ˈmiːdiə/ (n)}
{Websites and apps that enable users to create and share content or to participate in social networking.}
{Many people use social media to stay in touch with their friends and family.}
{Đây là một cụm danh từ dùng để chỉ mạng xã hội.}

\vocabentry{Spam}
{/spæm/ (n)}
{Irrelevant or inappropriate messages sent on the internet to a large number of recipients.}
{You should be careful about opening emails from unknown senders as they might be spam.}
{Đây là một danh từ dùng để chỉ thư rác.}

\vocabentry{Speaker}
{/ˈspiːkər/ (n)}
{A device that converts electrical signals into sound.}
{I have a pair of high-quality speakers for my computer.}
{Đây là một danh từ dùng để chỉ loa máy tính/thiết bị phát âm thanh.}

\vocabentry{Storage}
{/ˈstɔːrɪdʒ/ (n)}
{The action or method of storing something for future use.}
{My phone has 64 gigabytes of internal storage.}
{Đây là một danh từ dùng để chỉ dung lượng lưu trữ.}

\vocabentry{Streaming}
{/ˈstriːmɪŋ/ (n)}
{A method of transmitting or receiving data (especially video and audio material) over a computer network as a steady, continuous flow.}
{Video streaming services like Netflix have become very popular.}
{Đây là một danh từ dùng để chỉ truyền phát dữ liệu trực tuyến.}

\vocabentry{Sync}
{/sɪŋk/ (v)}
{To synchronize (data or devices).}
{You can sync your phone with your computer to keep your contacts and calendar up to date.}
{Đây là một động từ dùng để chỉ việc đồng bộ hóa dữ liệu.}

\vocabentry{Tablet}
{/ˈtæblət/ (n)}
{A portable computer that has a touchscreen as its primary input device.}
{I use my tablet for reading e-books and watching movies.}
{Đây là một danh từ dùng để chỉ máy tính bảng.}

\vocabentry{Tag}
{/tæɡ/ (v)}
{To label (something) with a tag; in social media, to identify (someone) in a post or photo.}
{You can tag your friends in your photos so they can see them.}
{Đây là một động từ dùng để chỉ việc gắn thẻ tên.}

\vocabentry{Template}
{/ˈtemplet/ (n)}
{A pre-designed document or file that can be used as a starting point for a new project.}
{You can use a resume template to make it easier to write your own.}
{Đây là một danh từ dùng để chỉ biểu mẫu/tài liệu mẫu.}

\vocabentry{Terminal}
{/ˈtɜːrmɪnl/ (n)}
{A device at which a user enters data for a computer system and which displays the received output.}
{The scientist entered the data into the terminal.}
{Đây là một danh từ dùng để chỉ thiết bị đầu cuối/màn hình nhập liệu.}

\vocabentry{Text}
{/tekst/ (v)}
{To send a text message to someone.}
{I'll text you the details later.}
{Đây là một động từ dùng để chỉ việc nhắn tin văn bản.}

\vocabentry{Thread}
{/θred/ (n)}
{A series of related messages on an online forum or social media platform.}
{You can follow the thread to see what other people are saying about the topic.}
{Đây là một danh từ dùng để chỉ luồng thảo luận trực tuyến.}

\vocabentry{Thumbnail}
{/ˈθʌmneɪl/ (n)}
{A small picture that represents a larger image or video.}
{You can click on the thumbnail to see the full-size image.}
{Đây là một danh từ dùng để chỉ hình ảnh thu nhỏ.}

\vocabentry{Timeline}
{/ˈtaɪmlaɪn/ (n)}
{A display of a list of events in chronological order.}
{You can see your friends' latest posts on your timeline.}
{Đây là một danh từ dùng để chỉ dòng thời gian trên mạng xã hội.}

\vocabentry{Toolbar}
{/ˈtuːlbɑːr/ (n)}
{A strip of icons used to perform certain functions in a computer program.}
{You can customize the toolbar to include your most frequently used commands.}
{Đây là một danh từ dùng để chỉ thanh công cụ.}

\vocabentry{Touchscreen}
{/ˈtʌtʃskriːn/ (n)}
{A display screen that is sensitive to the touch of a finger or stylus.}
{Most modern smartphones and tablets have touchscreens.}
{Đây là một danh từ dùng để chỉ màn hình cảm ứng.}

\vocabentry{Traffic}
{/ˈtræfɪk/ (n)}
{The amount of data moving across a network at a given point of time.}
{The website is experiencing a lot of traffic today.}
{Đây là một danh từ dùng để chỉ lưu lượng truy cập mạng.}

\vocabentry{Update}
{/ˌʌpˈdeɪt/ (v)}
{To make (something) more modern or up to date.}
{You should update your apps regularly to get the latest features and security fixes.}
{Đây là một động từ dùng để chỉ việc cập nhật phần mềm.}

\vocabentry{Upload}
{/ˌʌpˈloʊd/ (v)}
{To transfer (data) from one computer system to another, typically to a larger computer or a server.}
{You can upload your photos to social media to share them with your friends.}
{Đây là một động từ dùng để chỉ việc tải dữ liệu lên mạng.}

\vocabentry{URL}
{/ˌjuː ɑːr ˈel/ (n)}
{Uniform Resource Locator; the address of a web page.}
{You can type the URL of the website into your browser's address bar.}
{Đây là một danh từ viết tắt dùng để chỉ địa chỉ website.}

\vocabentry{User}
{/ˈjuːzər/ (n)}
{A person who uses a computer or other device.}
{The software is designed to be easy for the end user.}
{Đây là một danh từ dùng để chỉ người dùng.}

\vocabentry{Username}
{/ˈjuːzərneɪm/ (n)}
{A unique name used by a person to log in to a computer system or website.}
{You should choose a username that is easy for you to remember.}
{Đây là một danh từ dùng để chỉ tên đăng nhập.}

\vocabentry{Virus}
{/ˈvaɪrəs/ (n)}
{A piece of code which is capable of copying itself and typically has a detrimental effect, such as corrupting the system or destroying data.}
{Antivirus software can help protect your computer from viruses.}
{Đây là một danh từ dùng để chỉ vi-rút máy tính.}

\vocabentry{Webcam}
{/ˈwebkæm/ (n)}
{A video camera that inputs to a computer connected to the internet, so that its images can be viewed by internet users.}
{I use my webcam for video calls with my friends and family.}
{Đây là một danh từ dùng để chỉ camera máy tính.}

\vocabentry{Website}
{/ˈwebsaɪt/ (n)}
{A set of related web pages located under a single domain name.}
{You can find a lot of information on the university's website.}
{Đây là một danh từ dùng để chỉ trang web.}

\vocabentry{Wi-Fi}
{/ˈwaɪ faɪ/ (n)}
{A facility allowing computers, smartphones, or other devices to connect to the internet or communicate with one another wirelessly within a particular area.}
{Most cafes and libraries provide free Wi-Fi.}
{Đây là một danh từ dùng để chỉ mạng không dây.}

\vocabentry{Window}
{/ˈwɪndoʊ/ (n)}
{A rectangular area on a computer screen that displays a program or file.}
{You can open multiple windows at the same time and switch between them.}
{Đây là một danh từ dùng để chỉ cửa sổ chương trình máy tính.}

\vocabentry{Workstation}
{/ˈwɜːrksteɪʃn/ (n)}
{A powerful computer used for specialized tasks, such as graphic design or video editing.}
{He has a powerful workstation for his work as an architect.}
{Đây là một danh từ dùng để chỉ máy trạm/máy tính cấu hình cao.}

\vocabentry{Zip file}
{/ˈzɪp faɪl/ (n)}
{A file that contains one or more files that have been compressed to reduce their size.}
{You can send a group of files as a zip file to make them easier to email.}
{Đây là một cụm danh từ dùng để chỉ tệp nén.}

\vocabentry{Adapt}
{/əˈdæpt/ (v)}
{Make (something) suitable for a new use or purpose; modify.}
{The old factory was adapted for use as a modern office space.}
{Đây là một động từ dùng để chỉ việc thích nghi hoặc điều chỉnh thiết bị cho mục đích mới.}

\vocabentry{Advance}
{/ədˈvæns/ (n)}
{A development or improvement.}
{Technological advances have made life much easier in many ways.}
{Đây là một danh từ dùng để chỉ sự tiến bộ công nghệ.}

\vocabentry{Analyze}
{/ˈænəlaɪz/ (v)}
{Examine methodically and in detail.}
{Scientists are analyzing the data to find a solution to the problem.}
{Đây là một động từ dùng để chỉ việc phân tích dữ liệu.}

\vocabentry{Application}
{/ˌæplɪˈkeɪʃn/ (n)}
{A program or piece of software designed and written to fulfill a particular purpose of the user.}
{Word processing and spreadsheet programs are examples of applications.}
{Đây là một danh từ dùng để chỉ ứng dụng phần mềm.}

\vocabentry{Automated}
{/ˈɔːtəmeɪtɪd/ (adj)}
{Operated by largely automatic equipment.}
{The factory has an automated production line.}
{Đây là một tính từ dùng để chỉ tính tự động hóa.}

\vocabentry{Breakthrough}
{/ˈbreɪkθruː/ (n)}
{A sudden, dramatic, and important discovery or development.}
{The scientists have made a major breakthrough in cancer research.}
{Đây là một danh từ dùng để chỉ một bước đột phá lớn.}

\vocabentry{Complex}
{/ˈkɑːmpleks/ (adj)}
{Consisting of many different and connected parts.}
{Modern computers are extremely complex machines.}
{Đây là một tính từ dùng để chỉ tính phức tạp của hệ thống.}

\vocabentry{Component}
{/kəmˈpoʊnənt/ (n)}
{A part or element of a larger whole.}
{Each component of the machine must be in good working order.}
{Đây là một danh từ dùng để chỉ linh kiện/thành phần hệ thống.}

\vocabentry{Compute}
{/kəmˈpjuːt/ (v)}
{Reckon or calculate (a figure or amount).}
{The computer can compute large amounts of data in seconds.}
{Đây là một động từ dùng để chỉ việc tính toán bằng máy tính.}

\vocabentry{Concept}
{/ˈkɑːnsept/ (n)}
{An abstract idea; a general notion.}
{The concept of virtual reality has been around for decades.}
{Đây là một danh từ dùng để chỉ khái niệm công nghệ mới.}

\vocabentry{Cutting-edge}
{/ˌkʌtɪŋ ˈedʒ/ (adj)}
{Highly advanced; innovative or pioneering.}
{The new laboratory is equipped with cutting-edge technology.}
{Đây là một tính từ dùng để chỉ công nghệ tiên tiến nhất.}

\vocabentry{Data}
{/ˈdeɪtə/ (n)}
{Facts and statistics collected together for reference or analysis.}
{The company uses data to improve its products and services.}
{Đây là một danh từ dùng để chỉ dữ liệu.}

\vocabentry{Design}
{/dɪˈzaɪn/ (v)}
{Decide upon the look and functioning of (a building, garment, or other object), typically by making a detailed drawing of it.}
{The new smartphone was designed to be both stylish and functional.}
{Đây là một động từ dùng để chỉ việc thiết kế.}

\vocabentry{Efficiency}
{/ɪˈfɪʃnsi/ (n)}
{The state or quality of being efficient.}
{New technology has improved the efficiency of the manufacturing process.}
{Đây là một danh từ dùng để chỉ hiệu suất/năng suất công việc.}

\vocabentry{Electronic}
{/ɪˌlekˈtrɑːnɪk/ (adj)}
{(Of a device) having or operating by means of many small components, especially microchips and transistors.}
{Smartphones and laptops are examples of electronic devices.}
{Đây là một tính từ dùng để chỉ thuộc về điện tử.}

\vocabentry{Engineer}
{/ˌendʒɪˈnɪr/ (n)}
{A person who designs, builds, or maintains engines, machines, or public works.}
{He works as a software engineer for a tech company.}
{Đây là một danh từ dùng để chỉ kỹ sư.}

\vocabentry{Equipment}
{/ɪˈkwɪpmənt/ (n)}
{The necessary items for a particular purpose.}
{The lab is equipped with the latest scientific equipment.}
{Đây là một danh từ dùng để chỉ thiết bị công nghệ.}

\vocabentry{Experiment}
{/ɪkˈsperɪmənt/ (n)}
{A scientific procedure undertaken to make a discovery, test a hypothesis, or demonstrate a known fact.}
{The scientists are conducting an experiment to test their theory.}
{Đây là một danh từ dùng để chỉ thí nghiệm.}

\vocabentry{Function}
{/ˈfʌŋkʃn/ (n)}
{An activity or purpose natural to or intended for a person or thing.}
{The new app has several useful functions, such as a task manager and a calendar.}
{Đây là một danh từ dùng để chỉ chức năng/tính năng.}

\vocabentry{Future}
{/ˈfjuːtʃər/ (n)}
{The time or a period of time following the moment of speaking or writing.}
{Technology will continue to play a major role in our future.}
{Đây là một danh từ dùng để chỉ tương lai công nghệ.}

