% Auto-generated from raw.txt
% Usage:
%   \vocabentry{word}{/ipa/ (pos)}{definition}{example}{note}

\vocabentry{Sustainability}
{/səˌsteɪnəˈbɪləti/ (n)}
{The ability to be maintained at a certain rate or level without depleting natural resources.}
{The government is promoting sustainability in agriculture to ensure long-term food security.}
{Đây là một danh từ dùng để chỉ sự bền vững, khả năng duy trì sự cân bằng sinh thái mà không làm cạn kiệt tài nguyên.}

\vocabentry{Biodiversity}
{/ˌbaɪoʊdaɪˈvɜːrsəti/ (n)}
{The variety of plant and animal life in the world or in a particular habitat.}
{The construction of the dam will seriously threaten the biodiversity of the river.}
{Đây là một danh từ dùng để chỉ sự đa dạng sinh học, yếu tố sống còn để duy trì hệ sinh thái khỏe mạnh.}

\vocabentry{Conservation}
{/ˌkɑːnsərˈveɪʃn/ (n)}
{The protection of plants and animals and natural areas from the damaging effects of human activity.}
{Energy conservation is one of the most effective ways to reduce your carbon footprint.}
{Đây là một danh từ dùng để chỉ sự bảo tồn thiên nhiên và tài nguyên.}

\vocabentry{Contamination}
{/kənˌtæmɪˈneɪʃn/ (n)}
{The process of making something dirty or poisonous by adding chemicals or waste.}
{The environmental report highlighted the high levels of chemical contamination in the soil.}
{Đây là một danh từ dùng để chỉ sự làm bẩn, sự nhiễm độc hoặc nhiễm khuẩn môi trường.}

\vocabentry{Ecosystem}
{/ˈiːkoʊsɪstəm/ (n)}
{A biological community of interacting organisms and their physical environment.}
{Pollution can disrupt the delicate balance of a local ecosystem.}
{Đây là một danh từ dùng để chỉ hệ sinh thái, nơi các sinh vật tương tác với nhau và với môi trường.}

\vocabentry{Deforestation}
{/ˌdiːˌfɔːrɪˈsteɪʃn/ (n)}
{The action of clearing a wide area of trees.}
{Deforestation in the Amazon has reached an alarming rate in recent years.}
{Đây là một danh từ dùng để chỉ nạn phá rừng, một nguyên nhân chính gây ra biến đổi khí hậu.}

\vocabentry{Greenhouse effect}
{/ˈɡriːnhaʊs ɪˈfekt/ (n)}
{The trapping of the sun's warmth in a planet's lower atmosphere.}
{Carbon dioxide is the main gas responsible for the greenhouse effect.}
{Đây là một cụm danh từ dùng để chỉ hiệu ứng nhà kính, hiện tượng làm trái đất nóng lên.}

\vocabentry{Renewable}
{/rɪˈnuːəbl/ (adj)}
{(Of a natural resource or source of energy) not depleted when used.}
{Many countries are shifting from fossil fuels to renewable energy sources like wind and solar.}
{Đây là một tính từ dùng để chỉ năng lượng tái tạo, có thể phục hồi tự nhiên.}

\vocabentry{Degradation}
{/ˌdeɡrəˈdeɪʃn/ (n)}
{The process of something becoming worse or being destroyed.}
{Environmental degradation is often a direct result of rapid industrialization.}
{Đây là một danh từ dùng để chỉ sự suy thoái, thường dùng để mô tả sự giảm sút chất lượng môi trường.}

\vocabentry{Emissions}
{/iˈmɪʃnz/ (n)}
{The production and discharge of something, especially gas or radiation.}
{The new laws aim to drastically reduce carbon emissions from heavy industry.}
{Đây là một danh từ dùng để chỉ khí thải, đặc biệt là khí gây ô nhiễm từ nhà máy hoặc xe cộ.}

\vocabentry{Habitat}
{/ˈhæbɪtæt/ (n)}
{The natural home or environment of an animal, plant, or other organism.}
{Loss of habitat is the single biggest threat to wildlife today.}
{Đây là một danh từ dùng để chỉ môi trường sống tự nhiên của sinh vật.}

\vocabentry{Pollutant}
{/pəˈluːtənt/ (n)}
{A substance that pollutes something, especially water or the atmosphere.}
{Sulfur dioxide is a major air pollutant produced by coal-burning power plants.}
{Đây là một danh từ dùng để chỉ chất gây ô nhiễm.}

\vocabentry{Mitigation}
{/ˌmɪtɪˈɡeɪʃn/ (n)}
{The action of reducing the severity, seriousness, or painfulness of something.}
{The city is implementing flood mitigation measures to protect coastal homes.}
{Đây là một danh từ dùng để chỉ sự giảm nhẹ hoặc làm dịu các tác động tiêu cực của môi trường.}

\vocabentry{Desertification}
{/dɪˌzɜːrtɪfɪˈkeɪʃn/ (n)}
{The process by which fertile land becomes desert.}
{Overgrazing and poor irrigation are the primary causes of desertification.}
{Đây là một danh từ dùng để chỉ quá trình sa mạc hóa.}

\vocabentry{Biofuel}
{/ˈbaɪoʊfjuːəl/ (n)}
{A fuel derived immediately from living matter.}
{Biofuel production can provide a cleaner alternative to traditional fossil fuels.}
{Đây là một danh từ dùng để chỉ nhiên liệu sinh học, giải pháp thay thế cho xăng dầu.}

\vocabentry{Geothermal}
{/ˌdʒiːoʊˈθɜːrml/ (adj)}
{Relating to or produced by the internal heat of the earth.}
{Geothermal energy is a sustainable way to heat homes in volcanic regions.}
{Đây là một tính từ dùng để chỉ địa nhiệt, nguồn năng lượng từ lòng đất.}

\vocabentry{Disposable}
{/dɪˈspoʊzəbl/ (adj)}
{Intended to be used once and then thrown away.}
{We must reduce our reliance on disposable plastic products to save the oceans.}
{Đây là một tính từ dùng để chỉ các sản phẩm dùng một lần, gây gánh nặng rác thải.}

\vocabentry{Precipitation}
{/prɪˌsɪpɪˈteɪʃn/ (n)}
{Rain, snow, sleet, or hail that falls to the ground.}
{Higher temperatures are changing global patterns of precipitation.}
{Đây là một danh từ dùng để chỉ lượng mưa hoặc các hình thức giáng thủy.}

\vocabentry{Catastrophic}
{/ˌkætəˈstrɑːfɪk/ (adj)}
{Involving or causing sudden great damage or suffering.}
{Climate change could lead to catastrophic wildfires across the globe.}
{Đây là một tính từ dùng để chỉ thảm họa hoặc hậu quả cực kỳ nghiêm trọng.}

\vocabentry{Endangered}
{/ɪnˈdeɪndʒərd/ (adj)}
{(Of a species) at serious risk of extinction.}
{The giant panda is one of the world's most famous endangered species.}
{Đây là một tính từ dùng để chỉ các loài động thực vật có nguy cơ tuyệt chủng.}

\vocabentry{Afforestation}
{/əˌfɔːrɪˈsteɪʃn/ (n)}
{The establishment of a forest or stand of trees in an area where there was no previous tree cover.}
{Afforestation is a key strategy to combat climate change.}
{Đây là một danh từ dùng để chỉ việc trồng rừng trên diện tích chưa từng có rừng.}

\vocabentry{Agroecology}
{/ˌæɡroʊiˈkɑːlə dʒi/ (n)}
{Ecological processes applied to agricultural production systems.}
{Farmers are turning to agroecology to maintain soil health.}
{Đây là một danh từ dùng để chỉ ngành sinh thái nông nghiệp.}

\vocabentry{Aquifer}
{/ˈækwɪfər/ (n)}
{A body of permeable rock which can contain or transmit groundwater.}
{The city's water supply comes from a deep underground aquifer.}
{Đây là một danh từ dùng để chỉ tầng ngậm nước dưới lòng đất.}

\vocabentry{Atmosphere}
{/ˈætməsfɪr/ (n)}
{The envelope of gases surrounding the earth.}
{Industrial gases are accumulating in the Earth's atmosphere.}
{Đây là một danh từ dùng để chỉ bầu khí quyển.}

\vocabentry{Biodegradable}
{/ˌbaɪoʊdɪˈɡreɪdəbl/ (adj)}
{Capable of being decomposed by bacteria or living organisms.}
{Biodegradable waste can be turned into compost for gardening.}
{Đây là một tính từ dùng để chỉ chất có khả năng tự phân hủy sinh học.}

\vocabentry{Biosphere}
{/ˈbaɪoʊsfɪr/ (n)}
{The regions of the earth occupied by living organisms.}
{Every living thing on Earth is part of the global biosphere.}
{Đây là một danh từ dùng để chỉ sinh quyển.}

\vocabentry{Carbon footprint}
{/ˌkɑːrbən ˈfʊtprɪnt/ (n)}
{The total amount of greenhouse gases produced by human activities.}
{Flying long distances significantly increases your carbon footprint.}
{Đây là một cụm danh từ dùng để chỉ dấu chân carbon.}

\vocabentry{Carbon sequestration}
{/ˌkɑːrbən ˌsiːkwəˈstreɪʃn/ (n)}
{The natural or artificial process by which CO2 is removed from the atmosphere.}
{Peatlands are highly effective for natural carbon sequestration.}
{Đây là một cụm danh từ dùng để chỉ quá trình cô lập carbon.}

\vocabentry{Climate change}
{/ˈklaɪmət tʃeɪndʒ/ (n)}
{A change in global or regional climate patterns.}
{Climate change is a global crisis that requires urgent action.}
{Đây là một cụm danh từ dùng để chỉ biến đổi khí hậu.}

\vocabentry{Coastal erosion}
{/ˈkoʊstl ɪˈroʊʒn/ (n)}
{The loss of land along the coastline due to waves or tides.}
{Planting mangroves is a natural way to prevent coastal erosion.}
{Đây là một cụm danh từ dùng để chỉ sự xói mòn bờ biển.}

\vocabentry{Compost}
{/ˈkɑːmpoʊst/ (v)}
{To make vegetable matter into a decayed mixture used for fertilizing.}
{Instead of throwing away food scraps, you should compost them.}
{Đây là một động từ dùng để chỉ việc ủ phân hữu cơ.}

\vocabentry{Drought}
{/draʊt/ (n)}
{A prolonged period of abnormally low rainfall.}
{The agriculture sector suffered greatly during the recent drought.}
{Đây là một danh từ dùng để chỉ nạn hạn hán.}

\vocabentry{Eco-friendly}
{/ˌiːkoʊ ˈfrendli/ (adj)}
{Not harmful to the environment.}
{Eco-friendly vehicles help reduce urban air pollution.}
{Đây là một tính từ dùng để chỉ sự thân thiện với môi trường.}

\vocabentry{Effluent}
{/ˈefluənt/ (n)}
{Liquid waste or sewage discharged into a river or the sea.}
{The factory must treat its effluent before releasing it into the river.}
{Đây là một danh từ dùng để chỉ nước thải chảy ra môi trường.}

\vocabentry{Environmentalism}
{/ɪnˌvaɪrənˈmentəlɪzəm/ (n)}
{Advocacy of the preservation of the natural environment.}
{Growing awareness of pollution has boosted international environmentalism.}
{Đây là một danh từ dùng để chỉ chủ nghĩa môi trường.}

\vocabentry{Extinction}
{/ɪkˈstɪŋkʃn/ (n)}
{The state of a species being or becoming extinct.}
{Many island species are currently facing the threat of extinction.}
{Đây là một danh từ dùng để chỉ sự tuyệt chủng.}

\vocabentry{Fertilizer}
{/ˈfɜːrtəlaɪzər/ (n)}
{A substance added to soil to increase its fertility.}
{Synthetic fertilizers can harm the soil's natural microorganisms.}
{Đây là một danh từ dùng để chỉ phân bón.}

\vocabentry{Fossil fuel}
{/ˈfɑːsl fjuːəl/ (n)}
{A natural fuel such as coal or gas formed from living remains.}
{Burning fossil fuels is the primary cause of carbon emissions.}
{Đây là một cụm danh từ dùng để chỉ nhiên liệu hóa thạch.}

\vocabentry{Glacier}
{/ˈɡleɪʃər/ (n)}
{A slowly moving mass of ice.}
{Many glaciers in the Himalayas are melting rapidly.}
{Đây là một danh từ dùng để chỉ sông băng.}

\vocabentry{Global warming}
{/ˌɡloʊbl ˈwɔːrmɪŋ/ (n)}
{A gradual increase in the overall temperature of the earth's atmosphere.}
{Global warming is leading to more frequent heatwaves.}
{Đây là một cụm danh từ dùng để chỉ sự nóng lên toàn cầu.}

\vocabentry{Hazardous waste}
{/ˌhæzərdəs ˈweɪst/ (n)}
{Waste that has substantial threats to public health.}
{Hazardous waste requires specialized handling and disposal.}
{Đây là một cụm danh từ dùng để chỉ rác thải nguy hại.}

\vocabentry{Hydroponics}
{/ˌhaɪdrəˈpɑːnɪks/ (n)}
{Growing plants in nutrient-rich liquid without soil.}
{Hydroponics is an efficient way to grow food in cities.}
{Đây là một danh từ dùng để chỉ phương pháp thủy canh.}

\vocabentry{Hydropower}
{/ˈhaɪdroʊpaʊər/ (n)}
{Electricity produced from the energy of moving water.}
{Hydropower is a stable source of renewable energy.}
{Đây là một danh từ dùng để chỉ thủy điện.}

\vocabentry{Incineration}
{/ɪnˌsɪnəˈreɪʃn/ (n)}
{The destruction of waste material by burning.}
{Incineration is often used when landfill space is limited.}
{Đây là một danh từ dùng để chỉ sự thiêu hủy rác.}

\vocabentry{Landfill}
{/ˈlændfɪl/ (n)}
{A place to dispose of waste material by burying it.}
{We need to reduce waste to prevent landfills from overflowing.}
{Đây là một danh từ dùng để chỉ bãi chôn lấp rác.}

\vocabentry{Marine}
{/məˈriːn/ (adj)}
{Relating to or found in the sea.}
{Marine ecosystems are being damaged by plastic pollution.}
{Đây là một tính từ dùng để chỉ các yếu tố thuộc về biển.}

\vocabentry{Natural resource}
{/ˌnætʃrəl ˈriːsɔːrs/ (n)}
{Materials such as minerals and forests that occur in nature.}
{Countries must manage their natural resources sustainably.}
{Đây là một cụm danh từ dùng để chỉ tài nguyên thiên nhiên.}

\vocabentry{Non-renewable}
{/ˌnɑːn rɪˈnuːəbl/ (adj)}
{Existing in finite quantity; not capable of being replenished.}
{Oil is a non-renewable resource that we will eventually run out of.}
{Đây là một tính từ dùng để chỉ tài nguyên không tái tạo.}

\vocabentry{Ozone layer}
{/ˈoʊzoʊn leɪər/ (n)}
{A layer in the stratosphere that absorbs UV radiation.}
{Protecting the ozone layer is vital for human health.}
{Đây là một cụm danh từ dùng để chỉ tầng ozone.}

\vocabentry{Pesticide}
{/ˈpestɪsaɪd/ (n)}
{A substance used for destroying insects harmful to plants.}
{Pesticide runoff can contaminate nearby water sources.}
{Đây là một danh từ dùng để chỉ thuốc trừ sâu.}

\vocabentry{Preservation}
{/ˌprezərˈveɪʃn/ (n)}
{The action of keeping something in its present state.}
{The preservation of national parks is a priority for the government.}
{Đây là một danh từ dùng để chỉ sự bảo tồn nguyên trạng.}

\vocabentry{Pristine}
{/ˈprɪstiːn/ (adj)}
{In its original condition; unspoiled.}
{The island is famous for its pristine beaches and clear water.}
{Đây là một tính từ dùng để chỉ trạng thái nguyên sơ, sạch sẽ.}

\vocabentry{Recycle}
{/ˌriːˈsaɪkl/ (v)}
{To convert waste into reusable material.}
{You can recycle aluminum cans and glass bottles at the center.}
{Đây là một động từ dùng để chỉ việc tái chế.}

\vocabentry{Reforestation}
{/ˌriːˌfɔːrɪˈsteɪʃn/ (n)}
{The process of replanting an area with trees.}
{Reforestation helps restore habitats for endangered animals.}
{Đây là một danh từ dùng để chỉ sự tái trồng rừng.}

\vocabentry{Sanitation}
{/ˌsænɪˈteɪʃn/ (n)}
{Conditions relating to public health and sewage disposal.}
{Poor sanitation can lead to the spread of waterborne diseases.}
{Đây là một danh từ dùng để chỉ hệ thống vệ sinh.}

\vocabentry{Sewage}
{/ˈsuːɪdʒ/ (n)}
{Waste water and excrement conveyed in sewers.}
{Cities must have effective systems to treat their sewage.}
{Đây là một danh từ dùng để chỉ nước thải sinh hoạt.}

\vocabentry{Smog}
{/smɑːɡ/ (n)}
{Fog combined with smoke and atmospheric pollutants.}
{Heavy smog in the city can cause respiratory problems.}
{Đây là một danh từ dùng để chỉ sương khói ô nhiễm.}

\vocabentry{Solar energy}
{/ˌsoʊlər ˈenərdʒi/ (n)}
{Radiant light and heat from the Sun harnessed using technology.}
{Solar energy is a clean alternative to coal power.}
{Đây là một cụm danh từ dùng để chỉ năng lượng mặt trời.}

\vocabentry{Tidal power}
{/ˈtaɪdl paʊər/ (n)}
{Hydropower that converts the energy of tides into power.}
{Tidal power is a predictable and clean energy source.}
{Đây là một cụm danh từ dùng để chỉ năng lượng thủy triều.}

\vocabentry{Toxic}
{/ˈtɑːksɪk/ (adj)}
{Poisonous or harmful to the environment.}
{Toxic waste must be disposed of with extreme care.}
{Đây là một tính từ dùng để chỉ tính độc hại.}

\vocabentry{Urbanization}
{/ˌɜːrbənəˈzeɪʃn/ (n)}
{The process of making an area more city-like.}
{Rapid urbanization often destroys local green spaces.}
{Đây là một danh từ dùng để chỉ sự đô thị hóa.}

\vocabentry{Waste management}
{/ˈweɪst ˌmænɪdʒmənt/ (n)}
{Dealing with waste material from collection to disposal.}
{Proper waste management is essential for a clean city.}
{Đây là một cụm danh từ dùng để chỉ việc quản lý rác thải.}

\vocabentry{Wildlife}
{/ˈwaɪldlaɪf/ (n)}
{Wild animals collectively; native fauna.}
{The national park protects a wide variety of wildlife.}
{Đây là một danh từ dùng để chỉ động vật hoang dã.}

\vocabentry{Wind farm}
{/ˈwɪnd fɑːrm/ (n)}
{An area with a group of energy-producing wind turbines.}
{A new wind farm was built off the coast to generate power.}
{Đây là một cụm danh từ dùng để chỉ cánh đồng điện gió.}

\vocabentry{Acid rain}
{/ˌæsɪd ˈreɪn/ (n)}
{Rainfall made acidic by atmospheric pollution.}
{Acid rain can damage historical stone buildings.}
{Đây là một cụm danh từ dùng để chỉ mưa axit.}

\vocabentry{Alleviate}
{/əˈliːvieɪt/ (v)}
{To make a problem less severe.}
{Planting trees can help alleviate the heat island effect.}
{Đây là một động từ dùng để chỉ việc giảm nhẹ tác động tiêu cực.}

\vocabentry{Balance}
{/ˈbæləns/ (n)}
{A condition in which elements are in correct proportions.}
{We must maintain the delicate balance of nature.}
{Đây là một danh từ dùng để chỉ sự cân bằng sinh thái.}

\vocabentry{Cap-and-trade}
{/ˌkæp ən ˈtreɪd/ (n)}
{A system for controlling carbon emissions by market limits.}
{Cap-and-trade encourages companies to reduce their pollution.}
{Đây là một cụm danh từ dùng để chỉ hệ thống mua bán hạn ngạch phát thải.}

\vocabentry{Combustion}
{/kəmˈbʌstʃən/ (n)}
{The process of burning something.}
{The combustion of wood releases particulate matter into the air.}
{Đây là một danh từ dùng để chỉ sự đốt cháy.}

\vocabentry{Deplete}
{/dɪˈpliːt/ (v)}
{To use up the supply or resources of something.}
{Industrial fishing has depleted fish stocks in many oceans.}
{Đây là một động từ dùng để chỉ việc làm cạn kiệt tài nguyên.}

\vocabentry{Discharge}
{/dɪsˈtʃɑːrdʒ/ (v)}
{To allow liquid or gas to flow out from a source.}
{It is illegal to discharge toxic waste into the sea.}
{Đây là một động từ dùng để chỉ việc xả thải.}

\vocabentry{Disposal}
{/dɪˈspoʊzl/ (n)}
{The action of getting rid of something.}
{The safe disposal of batteries is important for the soil.}
{Đây là một danh từ dùng để chỉ việc xử lý/vứt bỏ rác.}

\vocabentry{Ecological}
{/ˌiːkəˈlɑːdʒɪkl/ (adj)}
{Relating to the relation of organisms to surroundings.}
{The forest fire caused an ecological disaster in the region.}
{Đây là một tính từ dùng để chỉ các yếu tố sinh thái.}

\vocabentry{Fumes}
{/fjuːmz/ (n)}
{Strong-smelling or dangerous gas or smoke.}
{Car exhaust fumes are a major source of air pollution.}
{Đây là một danh từ dùng để chỉ khói độc.}

\vocabentry{Geosphere}
{/ˈdʒiːəsfɪr/ (n)}
{The solid part of the earth consisting of the crust and outer mantle.}
{Volcanic eruptions transfer matter from the geosphere to the atmosphere.}
{Đây là một danh từ dùng để chỉ địa quyển.}

\vocabentry{Heatwave}
{/ˈhiːtweɪv/ (n)}
{A prolonged period of abnormally hot weather.}
{During the heatwave, water consumption increased significantly.}
{Đây là một danh từ dùng để chỉ đợt nắng nóng.}

\vocabentry{Inorganic}
{/ˌɪnɔːrˈɡænɪk/ (adj)}
{Not consisting of or deriving from living matter.}
{Inorganic fertilizers are used to speed up crop growth.}
{Đây là một tính từ dùng để chỉ các chất vô cơ.}

\vocabentry{Interdependence}
{/ˌɪntərdɪˈpendəns/ (n)}
{Dependence of two or more things on each other.}
{Ecosystems thrive on the interdependence of various species.}
{Đây là một danh từ dùng để chỉ sự phụ thuộc lẫn nhau.}

\vocabentry{Irreplaceable}
{/ˌɪrɪˈpleɪsəbl/ (adj)}
{Too valuable or rare to be replaced.}
{Ancient forests are irreplaceable parts of our natural heritage.}
{Đây là một tính từ dùng để chỉ sự không thể thay thế.}

\vocabentry{Litter}
{/ˈlɪtər/ (v)}
{To make a place untidy with rubbish.}
{People who litter on the beach should be heavily fined.}
{Đây là một động từ dùng để chỉ việc xả rác bừa bãi.}

\vocabentry{Low-carbon}
{/ˌloʊ ˈkɑːrbən/ (adj)}
{Resulting in small net release of CO2.}
{Transitioning to a low-carbon economy is necessary for the future.}
{Đây là một tính từ dùng để chỉ hoạt động ít phát thải carbon.}

\vocabentry{Meteorology}
{/ˌmiːtiəˈrɑːlədʒi/ (n)}
{The science of the processes of the atmosphere.}
{Meteorology helps us prepare for severe weather events.}
{Đây là một danh từ dùng để chỉ khí tượng học.}

\vocabentry{Natural disaster}
{/ˌnætʃrəl dɪˈzæstər/ (n)}
{A natural event such as a flood that causes damage.}
{Earthquake is a natural disaster that is hard to predict.}
{Đây là một cụm danh từ dùng để chỉ thiên tai.}

\vocabentry{Noxious}
{/ˈnɑːkʃəs/ (adj)}
{Harmful, poisonous, or very unpleasant.}
{The fire released noxious gases into the surrounding air.}
{Đây là một tính từ dùng để chỉ tính độc hại.}

\vocabentry{Offsetting}
{/ˈɔːfsetɪŋ/ (n)}
{Counteracting something by having an equal opposite effect.}
{Carbon offsetting involves planting trees to balance CO2 emissions.}
{Đây là một danh từ dùng để chỉ sự bù đắp carbon.}

\vocabentry{Overfishing}
{/ˌoʊvərˈfɪʃɪŋ/ (n)}
{Depleting the stock of fish by too much fishing.}
{Overfishing threatens the food security of coastal nations.}
{Đây là một danh từ dùng để chỉ nạn đánh bắt cá quá mức.}

\vocabentry{Overpopulation}
{/ˌoʊvərpɑːpjuˈleɪʃn/ (n)}
{Being populated with excessively large numbers.}
{Overpopulation puts pressure on urban infrastructure.}
{Đây là một danh từ dùng để chỉ tình trạng bùng nổ dân số.}

\vocabentry{Permafrost}
{/ˈpɜːrməfrɔːst/ (n)}
{A subsurface layer of soil that remains frozen year-round.}
{Thawing permafrost releases methane into the atmosphere.}
{Đây là một danh từ dùng để chỉ tầng đất đóng băng vĩnh cửu.}

\vocabentry{Plastics}
{/ˈplæstɪks/ (n)}
{Synthetic materials made from organic polymers.}
{Single-use plastics are being banned in many countries.}
{Đây là một danh từ dùng để chỉ các loại nhựa.}

\vocabentry{Poaching}
{/ˈpoʊtʃɪŋ/ (n)}
{Illegal hunting or capturing of wild animals.}
{Poaching has pushed the black rhino to the brink of extinction.}
{Đây là một danh từ dùng để chỉ nạn săn bắn trộm.}

\vocabentry{Precautionary}
{/prɪˈkɔːʃəneri/ (adj)}
{Done as a precaution to prevent harm.}
{We must take a precautionary approach to genetic engineering.}
{Đây là một tính từ dùng để chỉ tính phòng ngừa.}

\vocabentry{Predator}
{/ˈpredətər/ (n)}
{An animal that naturally preys on others.}
{Wolves are important predators in the forest ecosystem.}
{Đây là một danh từ dùng để chỉ kẻ săn mồi.}

\vocabentry{Purify}
{/ˈpjʊrɪfaɪ/ (v)}
{To remove contaminants from something.}
{Advanced filters can purify contaminated river water.}
{Đây là một động từ dùng để chỉ việc làm sạch/thanh lọc.}

\vocabentry{Radiation}
{/ˌreɪdiˈeɪʃn/ (n)}
{The emission of energy as electromagnetic waves.}
{High levels of radiation can be harmful to living cells.}
{Đây là một danh từ dùng để chỉ sự bức xạ/phóng xạ.}

\vocabentry{Raw material}
{/ˌrɔː məˈtɪriəl/ (n)}
{Basic material from which a product is made.}
{Wood is the primary raw material for the paper industry.}
{Đây là một cụm danh từ dùng để chỉ nguyên liệu thô.}

\vocabentry{Refurbish}
{/ˌriːˈfɜːrbɪʃ/ (v)}
{To renovate and redecorate something for reuse.}
{Refurbishing old furniture is an eco-friendly choice.}
{Đây là một động từ dùng để chỉ việc tân trang để tái sử dụng.}

\vocabentry{Regulation}
{/ˌreɡjuˈleɪʃn/ (n)}
{A rule or directive maintained by an authority.}
{Environmental regulations help limit industrial pollution.}
{Đây là một danh từ dùng để chỉ các quy định/luật lệ.}

\vocabentry{Resilient}
{/rɪˈzɪliənt/ (adj)}
{Able to withstand or recover quickly from difficulties.}
{We need to build resilient cities to face climate change.}
{Đây là một tính từ dùng để chỉ khả năng phục hồi.}

\vocabentry{Run-off}
{/ˈrʌn ɔːf/ (n)}
{The draining away of water from the surface of an area of land.}
{Stormwater run-off in cities often carries pollutants into rivers.}
{Đây là một danh từ dùng để chỉ dòng chảy tràn.}

\vocabentry{Scarcity}
{/ˈskersəti/ (n)}
{The state of being in short supply.}
{Water scarcity is a major issue in many desert regions.}
{Đây là một danh từ dùng để chỉ sự khan hiếm.}

\vocabentry{Shortage}
{/ˈʃɔːrtɪdʒ/ (n)}
{A situation in which something needed cannot be obtained.}
{There is a critical shortage of clean water in the refugee camp.}
{Đây là một danh từ dùng để chỉ sự thiếu hụt.}

\vocabentry{Soil erosion}
{/ˈsɔɪl ɪˌroʊʒn/ (n)}
{The displacement of the upper layer of soil.}
{Heavy rain without trees leads to rapid soil erosion.}
{Đây là một cụm danh từ dùng để chỉ sự xói mòn đất.}

\vocabentry{Species}
{/ˈspiːʃiːz/ (n)}
{A group of living organisms consisting of similar individuals.}
{Many species of insects are yet to be discovered.}
{Đây là một danh từ dùng để chỉ loài.}

\vocabentry{Sustainable}
{/səˈsteɪnəbl/ (adj)}
{Able to be maintained at a certain level.}
{Wind energy is a sustainable alternative to coal.}
{Đây là một tính từ dùng để chỉ tính bền vững.}

\vocabentry{Terrestrial}
{/təˈrestriəl/ (adj)}
{Relating to the earth; on or relating to land.}
{Terrestrial animals are often affected by land development.}
{Đây là một tính từ dùng để chỉ các sinh vật trên cạn.}

\vocabentry{Threatened}
{/ˈθretnd/ (adj)}
{Likely to become endangered in the future.}
{The polar bear is currently a threatened species due to ice loss.}
{Đây là một tính từ dùng để chỉ các loài đang bị đe dọa.}

\vocabentry{Toxin}
{/ˈtɑːksɪn/ (n)}
{A poison of plant or animal origin.}
{Some plants produce toxins to protect themselves from insects.}
{Đây là một danh từ dùng để chỉ độc tố.}

\vocabentry{Vegetation}
{/ˌvedʒəˈteɪʃn/ (n)}
{Plants considered collectively; plant life.}
{The dense vegetation of the jungle makes travel difficult.}
{Đây là một danh từ dùng để chỉ thảm thực vật.}

\vocabentry{Water cycle}
{/ˈwɔːtər ˌsaɪkl/ (n)}
{The cycle of water circulating between oceans and land.}
{Evaporation and precipitation are key parts of the water cycle.}
{Đây là một cụm danh từ dùng để chỉ chu kỳ nước.}

\vocabentry{Zero-emission}
{/ˌzɪroʊ iˈmɪʃn/ (adj)}
{Releasing no pollutants into the atmosphere.}
{Zero-emission vehicles are essential for net-zero goals.}
{Đây là một tính từ dùng để chỉ tính không phát thải.}

\vocabentry{Accumulate}
{/əˈkjuːmjəleɪt/ (v)}
{To gather or build up over time.}
{Heavy metals can accumulate in the tissues of fish.}
{Đây là một động từ dùng để chỉ sự tích tụ chất độc.}

\vocabentry{Acidify}
{/əˈsɪdɪfaɪ/ (v)}
{To make or become acid.}
{Increased CO2 levels cause the oceans to acidify.}
{Đây là một động từ dùng để chỉ sự axit hóa, thường dùng cho đại dương.}

\vocabentry{Adaptation}
{/ˌædæpˈteɪʃn/ (n)}
{The process of changing to suit a new environment.}
{Coastal cities need adaptation plans for rising sea levels.}
{Đây là một danh từ dùng để chỉ sự thích nghi.}

\vocabentry{Aerobic}
{/erˈoʊbɪk/ (adj)}
{Requiring free oxygen for life or processes.}
{Composting is an aerobic process that needs air to work.}
{Đây là một tính từ dùng để chỉ tính hiếu khí.}

\vocabentry{Anaerobic}
{/ˌænerˈoʊbɪk/ (adj)}
{Relating to or requiring an absence of free oxygen.}
{Methane is produced by anaerobic bacteria in landfills.}
{Đây là một tính từ dùng để chỉ tính kỵ khí.}

\vocabentry{Anthropogenic}
{/ˌænθrəpəˈdʒenɪk/ (adj)}
{Originating in human activity.}
{Most scientists agree that climate change is anthropogenic.}
{Đây là một tính từ dùng để chỉ các tác động do con người gây ra.}

\vocabentry{Arable}
{/ˈærəbl/ (adj)}
{(Of land) used or suitable for growing crops.}
{The amount of arable land is decreasing due to urbanization.}
{Đây là một tính từ dùng để chỉ đất canh tác.}

\vocabentry{Arid}
{/ˈærɪd/ (adj)}
{(Of land or climate) having little or no rain.}
{Arid regions require special irrigation systems for farming.}
{Đây là một tính từ dùng để chỉ tính khô hạn.}

\vocabentry{Biomass}
{/ˈbaɪoʊmæs/ (n)}
{Organic matter used as a fuel, especially in agriculture.}
{Biomass energy can be produced from wood and agricultural waste.}
{Đây là một danh từ dùng để chỉ sinh khối.}

\vocabentry{Bioremediation}
{/ˌbaɪoʊrɪˌmiːdiˈeɪʃn/ (n)}
{The use of microbes to clean up contaminated soil or water.}
{Bioremediation was used to clean up the oil spill on the beach.}
{Đây là một danh từ dùng để chỉ công nghệ xử lý sinh học.}

\vocabentry{Captivity}
{/kæpˈtɪvəti/ (n)}
{The condition of being imprisoned or confined.}
{Breeding animals in captivity can help prevent extinction.}
{Đây là một danh từ dùng để chỉ sự nuôi nhốt.}

\vocabentry{Carbon neutral}
{/ˌkɑːrbən ˈnuːtrəl/ (adj)}
{Making no net release of CO2 to the atmosphere.}
{The company aims to become carbon neutral by 2030.}
{Đây là một tính từ dùng để chỉ tính trung hòa carbon.}

\vocabentry{Clear-cutting}
{/ˈklɪrkʌtɪŋ/ (n)}
{The process of cutting down all trees in an area.}
{Clear-cutting leads to severe soil erosion and habitat loss.}
{Đây là một danh từ dùng để chỉ việc chặt phá trắng rừng.}

\vocabentry{Coexistence}
{/ˌkoʊɪɡˈzɪstəns/ (n)}
{The state of living together in the same place at the same time.}
{We must find a way for human development and wildlife coexistence.}
{Đây là một danh từ dùng để chỉ sự chung sống hòa bình.}

\vocabentry{Consumption}
{/kənˈsʌmpʃn/ (n)}
{The using up of a resource.}
{Reducing energy consumption is vital for a sustainable future.}
{Đây là một danh từ dùng để chỉ sự tiêu thụ.}

\vocabentry{Corridors}
{/ˈkɔːrɪdɔːrz/ (n)}
{Strips of natural habitat connecting populations of wildlife.}
{Wildlife corridors allow animals to travel safely between forests.}
{Đây là một danh từ dùng để chỉ hành lang sinh thái.}

\vocabentry{Critically endangered}
{/ˌkrɪtɪkli ɪnˈdeɪndʒərd/ (adj)}
{Facing an extremely high risk of extinction in the wild.}
{The Sumatran rhino is now critically endangered.}
{Đây là một tính từ dùng để chỉ mức độ cực kỳ nguy cấp.}

\vocabentry{Decompose}
{/ˌdiːkəmˈpoʊz/ (v)}
{To make or become rotten; decay.}
{Plastic bottles can take hundreds of years to decompose.}
{Đây là một động từ dùng để chỉ sự phân hủy.}

\vocabentry{Desalination}
{/ˌdiːˌsælɪˈneɪʃn/ (n)}
{The process of removing salt from seawater.}
{Desalination plants provide fresh water for many coastal cities.}
{Đây là một danh từ dùng để chỉ sự khử muối.}

\vocabentry{Ecotourism}
{/ˈiːkoʊtʊrɪzəm/ (n)}
{Tourism directed toward natural environments.}
{Ecotourism supports local conservation efforts and economies.}
{Đây là một danh từ dùng để chỉ du lịch sinh thái.}

\vocabentry{Endemic}
{/enˈdemɪk/ (adj)}
{(Of a plant or animal) native and restricted to a certain place.}
{The lemur is endemic to the island of Madagascar.}
{Đây là một tính từ dùng để chỉ loài đặc hữu.}

\vocabentry{Energy-efficient}
{/ˌenərdʒi ɪˈfɪʃnt/ (adj)}
{Using less energy to provide the same level of performance.}
{Energy-efficient light bulbs last longer and save money.}
{Đây là một tính từ dùng để chỉ tính tiết kiệm năng lượng.}

\vocabentry{Environmental impact}
{/ɪnˌvaɪrənˈmentl ˈɪmpækt/ (n)}
{The effect of an activity on the environment.}
{Companies must assess the environmental impact of their projects.}
{Đây là một cụm danh từ dùng để chỉ tác động môi trường.}

\vocabentry{Estuary}
{/ˈestʃueri/ (n)}
{The tidal mouth of a large river where it meets the stream.}
{Estuaries are important breeding grounds for many fish species.}
{Đây là một danh từ dùng để chỉ cửa sông.}

\vocabentry{Eutrophication}
{/juːˌtrɑːfɪˈkeɪʃn/ (n)}
{Excessive richness of nutrients in a body of water.}
{Eutrophication causes algae to grow rapidly, killing fish.}
{Đây là một danh từ dùng để chỉ hiện tượng phú dưỡng.}

\vocabentry{Fertile}
{/ˈfɜːrtl/ (adj)}
{(Of soil or land) producing or capable of producing abundant crops.}
{The river valley has extremely fertile soil for farming.}
{Đây là một tính từ dùng để chỉ đất đai màu mỡ.}

\vocabentry{Food chain}
{/ˈfuːd tʃeɪn/ (n)}
{A series of organisms each dependent on the next as a food source.}
{Mercury pollution can move up the food chain to humans.}
{Đây là một cụm danh từ dùng để chỉ chuỗi thức ăn.}

\vocabentry{Genetic diversity}
{/dʒəˈnetɪk daɪˈvɜːrsəti/ (n)}
{The total number of genetic characteristics in the genetic makeup of a species.}
{Genetic diversity helps a species survive changing environments.}
{Đây là một cụm danh từ dùng để chỉ sự đa dạng di truyền.}

\vocabentry{Geology}
{/dʒiˈɑːlədʒi/ (n)}
{The science that deals with the earth's physical structure.}
{Understanding local geology is important for groundwater management.}
{Đây là một danh từ dùng để chỉ địa chất học.}

\vocabentry{Harvest}
{/ˈhɑːrvɪst/ (v)}
{To gather a crop as a harvest.}
{Farmers harvest wheat in late summer.}
{Đây là một động từ dùng để chỉ việc thu hoạch.}

\vocabentry{Hydroelectric}
{/ˌhaɪdroʊɪˈlektrɪk/ (adj)}
{Relating to the generation of electricity using flowing water.}
{Hydroelectric dams provide a significant portion of the country's power.}
{Đây là một tính từ dùng để chỉ thuộc về thủy điện.}

\vocabentry{Imbalance}
{/ɪmˈbæləns/ (n)}
{Lack of proportion or relation between corresponding things.}
{Overhunting can cause a serious imbalance in the ecosystem.}
{Đây là một danh từ dùng để chỉ sự mất cân bằng.}

\vocabentry{Industrialization}
{/ɪnˌdʌstriələˈzeɪʃn/ (n)}
{The development of industries in a country or region.}
{Rapid industrialization has led to increased carbon emissions.}
{Đây là một danh từ dùng để chỉ sự công nghiệp hóa.}

\vocabentry{Intensive farming}
{/ɪnˈtensɪv ˈfɑːrmɪŋ/ (n)}
{Farming that uses a lot of machinery and chemicals.}
{Intensive farming can lead to soil exhaustion over time.}
{Đây là một cụm danh từ dùng để chỉ thâm canh/nông nghiệp cường độ cao.}

\vocabentry{Invasive species}
{/ɪnˈveɪsɪv ˈspiːʃiːz/ (n)}
{A species that is not native to a specific location and has a tendency to spread.}
{Invasive species can outcompete native plants for resources.}
{Đây là một cụm danh từ dùng để chỉ loài xâm lấn.}

\vocabentry{Irrigation}
{/ˌɪrɪˈɡeɪʃn/ (n)}
{The supply of water to land or crops to help growth.}
{Efficient irrigation systems can save a huge amount of water.}
{Đây là một danh từ dùng để chỉ sự tưới tiêu.}

\vocabentry{Land reclamation}
{/ˈlænd ˌrekləˈmeɪʃn/ (n)}
{The process of creating new land from oceans or riverbeds.}
{Land reclamation projects are common in crowded coastal cities.}
{Đây là một cụm danh từ dùng để chỉ sự lấn biển/cải tạo đất.}

\vocabentry{Lead-free}
{/ˌled ˈfriː/ (adj)}
{Not containing lead.}
{Lead-free gasoline has significantly reduced toxic air emissions.}
{Đây là một tính từ dùng để chỉ sản phẩm không chứa chì.}

\vocabentry{Leaching}
{/ˈliːtʃɪŋ/ (n)}
{The process of a substance being washed out from the soil by rain.}
{Leaching of chemicals from landfills can contaminate groundwater.}
{Đây là một danh từ dùng để chỉ sự rửa trôi/rò rỉ chất độc.}

\vocabentry{Magnification}
{/ˌmæɡnɪfɪˈkeɪʃn/ (n)}
{The increase in concentration of a substance as it moves up the food chain.}
{Biomagnification makes predators more vulnerable to toxins.}
{Đây là một danh từ dùng để chỉ sự tích tụ sinh học.}

\vocabentry{Microplastics}
{/ˈmaɪkroʊplæstɪks/ (n)}
{Extremely small pieces of plastic debris in the environment.}
{Microplastics are being ingested by marine life worldwide.}
{Đây là một danh từ dùng để chỉ hạt vi nhựa.}

\vocabentry{Natural habitat}
{/ˌnætʃrəl ˈhæbɪtæt/ (n)}
{The environment in which an animal or plant normally lives.}
{Many animals are losing their natural habitat to suburban sprawl.}
{Đây là một cụm danh từ dùng để chỉ môi trường sống tự nhiên.}

\vocabentry{Net-zero}
{/ˌnet ˈzɪroʊ/ (n)}
{Achieving a balance between the greenhouse gases put into and taken out of the atmosphere.}
{To stop global warming, the world must reach net-zero by 2050.}
{Đây là một danh từ dùng để chỉ phát thải ròng bằng không.}

\vocabentry{Nitrogen cycle}
{/ˈnaɪtrədʒən ˌsaɪkl/ (n)}
{The series of processes by which nitrogen circulates.}
{The nitrogen cycle is essential for providing nutrients to plants.}
{Đây là một cụm danh từ dùng để chỉ chu kỳ nitơ.}

\vocabentry{Ocean currents}
{/ˈoʊʃn ˌkɜːrənts/ (n)}
{Continuous, directed movements of seawater.}
{Ocean currents play a major role in regulating global climate.}
{Đây là một cụm danh từ dùng để chỉ các dòng hải lưu.}

\vocabentry{Off-grid}
{/ˌɔːf ˈɡrɪd/ (adj)}
{Not using or depending on public utilities (electricity, water).}
{Living off-grid often involves using solar panels and wind turbines.}
{Đây là một tính từ dùng để chỉ sống độc lập với mạng lưới điện quốc gia.}

\vocabentry{Organic farming}
{/ɔːrˈɡænɪk ˈfɑːrmɪŋ/ (n)}
{Farming without the use of synthetic chemicals.}
{Organic farming is better for the long-term health of the soil.}
{Đây là một cụm danh từ dùng để chỉ nông nghiệp hữu cơ.}

\vocabentry{Organism}
{/ˈɔːrɡənɪzəm/ (n)}
{An individual animal, plant, or single-celled life form.}
{Bacteria are simple organisms found almost everywhere on Earth.}
{Đây là một danh từ dùng để chỉ một sinh vật sống.}

\vocabentry{Particulate matter}
{/pɑːrˈtɪkjələt ˈmætər/ (n)}
{Small particles of solid or liquid matter suspended in air.}
{Particulate matter from diesel engines is harmful to human lungs.}
{Đây là một cụm danh từ dùng để chỉ bụi mịn/vật chất dạng hạt.}

\vocabentry{Permeable}
{/ˈpɜːrmiebl/ (adj)}
{(Of a material or membrane) allowing liquids or gases to pass through it.}
{Permeable pavement helps reduce urban flooding by absorbing rain.}
{Đây là một tính từ dùng để chỉ tính thấm được.}

\vocabentry{Photosynthesis}
{/ˌfoʊtoʊˈsɪnθəsɪs/ (n)}
{The process by which plants use sunlight to synthesize foods.}
{Photosynthesis produces the oxygen that most living things need to survive.}
{Đây là một danh từ dùng để chỉ quá trình quang hợp.}

\vocabentry{Plate tectonics}
{/ˌpleɪt tekˈtɑːnɪks/ (n)}
{The theory that the earth's outer shell is divided into several plates.}
{Plate tectonics explains how mountains are formed and earthquakes occur.}
{Đây là một cụm danh từ dùng để chỉ kiến tạo mảng.}

\vocabentry{Population density}
{/ˌpɑːpjuˈleɪʃn ˈdensəti/ (n)}
{The number of people living in each unit of area.}
{High population density in cities leads to more waste production.}
{Đây là một cụm danh từ dùng để chỉ mật độ dân số.}

\vocabentry{Potable}
{/ˈpoʊtəbl/ (adj)}
{Safe to drink; drinkable.}
{In many parts of the world, potable water is becoming scarce.}
{Đây là một tính từ dùng để chỉ nước có thể uống được.}

\vocabentry{Recharge}
{/ˌriːˈtʃɑːrdʒ/ (v)}
{To refill or replenish (water in an aquifer).}
{Rainwater is necessary to recharge underground aquifers.}
{Đây là một động từ dùng để chỉ sự bổ cập/làm đầy lại nguồn nước ngầm.}

\vocabentry{Red list}
{/ˈred lɪst/ (n)}
{A list of species at risk of extinction.}
{The IUCN Red List is a critical indicator of the health of biodiversity.}
{Đây là một cụm danh từ dùng để chỉ Sách Đỏ.}

\vocabentry{Residual}
{/rɪˈzɪdʒuəl/ (adj)}
{Remaining after the greater part or quantity has gone.}
{Residual pesticides can remain on vegetables even after washing.}
{Đây là một tính từ dùng để chỉ dư lượng, thường nói về hóa chất.}

\vocabentry{Restoration}
{/ˌrestəˈreɪʃn/ (n)}
{The action of returning something to its former condition.}
{The restoration of the local wetlands has attracted many birds.}
{Đây là một danh từ dùng để chỉ sự phục hồi môi trường.}

\vocabentry{Salination}
{/ˌsælɪˈneɪʃn/ (n)}
{The process by which water-soluble salts accumulate in the soil.}
{Over-irrigation in dry climates can lead to soil salination.}
{Đây là một danh từ dùng để chỉ sự nhiễm mặn.}

\vocabentry{Sanctuary}
{/ˈsæŋktʃueri/ (n)}
{A place where birds or animals are protected from being hunted.}
{The island is a sanctuary for rare and endangered seabirds.}
{Đây là một danh từ dùng để chỉ khu bảo tồn/nơi trú ẩn.}

\vocabentry{Sediment}
{/ˈsedɪmənt/ (n)}
{Matter that settles to the bottom of a liquid.}
{Sediment build-up can block river flow and affect fish.}
{Đây là một danh từ dùng để chỉ trầm tích.}

\vocabentry{Slash-and-burn}
{/ˌslæʃ ənd ˈbɜːrn/ (adj)}
{A method of agriculture in which trees are cut down and burned to clear land.}
{Slash-and-burn agriculture is a major cause of deforestation.}
{Đây là một tính từ dùng để chỉ phương pháp du canh đốt rừng.}

\vocabentry{Sludge}
{/slʌdʒ/ (n)}
{Thick, soft, wet mud or a similar viscous mixture.}
{Sewage sludge can be treated and used as a fertilizer.}
{Đây là một danh từ dùng để chỉ bùn thải.}

\vocabentry{Solar panel}
{/ˌsoʊlər ˈpænl/ (n)}
{A panel designed to absorb the sun's rays as a source of energy.}
{Installing solar panels can significantly reduce your electricity bill.}
{Đây là một cụm danh từ dùng để chỉ tấm pin mặt trời.}

\vocabentry{Stratosphere}
{/ˈstrætəsfɪr/ (n)}
{The layer of the earth's atmosphere above the troposphere.}
{The ozone layer is located within the Earth's stratosphere.}
{Đây là một danh từ dùng để chỉ tầng bình lưu.}

\vocabentry{Subsidy}
{/ˈsʌbsədi/ (n)}
{A sum of money granted by the government to assist an industry.}
{Government subsidies can help lower the cost of electric cars.}
{Đây là một danh từ dùng để chỉ sự trợ cấp, thường dùng cho năng lượng sạch.}

\vocabentry{Sulphur dioxide}
{/ˌsʌlfər daɪˈɑːksaɪd/ (n)}
{A colorless pungent toxic gas formed by burning sulfur in air.}
{Sulphur dioxide is one of the main gases that cause acid rain.}
{Đây là một cụm danh từ dùng để chỉ khí SO2.}

\vocabentry{Thermal}
{/ˈθɜːrml/ (adj)}
{Relating to heat.}
{Thermal pollution occurs when hot water is released into a river.}
{Đây là một tính từ dùng để chỉ thuộc về nhiệt.}

\vocabentry{Topsoil}
{/ˈtɑːpsɔɪl/ (n)}
{The top layer of soil.}
{It takes hundreds of years to form just one inch of topsoil.}
{Đây là một danh từ dùng để chỉ tầng đất mặt.}

\vocabentry{Trace element}
{/ˈtreɪs ˌelɪmənt/ (n)}
{A chemical element present in very small amounts.}
{Some trace elements are necessary for the growth of plants.}
{Đây là một cụm danh từ dùng để chỉ nguyên tố vi lượng.}

\vocabentry{Transpiration}
{/ˌtrænspɪˈreɪʃn/ (n)}
{The process by which plants give off water vapor through leaves.}
{Transpiration helps cool the surrounding air in a forest.}
{Đây là một danh từ dùng để chỉ sự thoát hơi nước ở cây.}

\vocabentry{Turbine}
{/ˈtɜːrbaɪn/ (n)}
{A machine for producing continuous power by rotating wheels.}
{Wind turbines convert the wind's kinetic energy into electricity.}
{Đây là một danh từ dùng để chỉ tua-bin.}

\vocabentry{Ultraviolet}
{/ˌʌltrəˈvaɪələt/ (adj)}
{Having a wavelength shorter than that of the violet end of the visible spectrum.}
{The ozone layer protects us from harmful ultraviolet rays.}
{Đây là một tính từ dùng để chỉ tia cực tím.}

\vocabentry{Vulnerable}
{/ˈvʌlnərəbl/ (adj)}
{Susceptible to physical or emotional attack or harm.}
{Developing nations are the most vulnerable to climate change.}
{Đây là một tính từ dùng để chỉ sự dễ bị tổn thương.}

\vocabentry{Waste disposal}
{/ˈweɪst dɪˌspoʊzl/ (n)}
{The act of getting rid of waste material.}
{Improper waste disposal can lead to serious environmental pollution.}
{Đây là một cụm danh từ dùng để chỉ sự xử lý rác thải.}

\vocabentry{Wastewater}
{/ˈweɪstwaːtər/ (n)}
{Water that has been used in the home, in a business, or as part of an industrial process.}
{Wastewater treatment plants clean the water before it returns to rivers.}
{Đây là một danh từ dùng để chỉ nước thải.}

\vocabentry{Water pollution}
{/ˈwɔːtər pəˌluːʃn/ (n)}
{The contamination of water bodies.}
{Water pollution can kill fish and make people sick.}
{Đây là một cụm danh từ dùng để chỉ ô nhiễm nguồn nước.}

\vocabentry{Water table}
{/ˈwɔːtər ˌteɪbl/ (n)}
{The level below which the ground is saturated with water.}
{During a drought, the water table can drop significantly.}
{Đây là một cụm danh từ dùng để chỉ mực nước ngầm.}

\vocabentry{Weathering}
{/ˈweðərɪŋ/ (n)}
{The wear away or change of appearance of something by long exposure to air.}
{Weathering of rocks is a natural part of soil formation.}
{Đây là một danh từ dùng để chỉ sự phong hóa.}

\vocabentry{Wetland}
{/ˈwetlænd/ (n)}
{Land consisting of marshes or swamps; saturated land.}
{Wetlands act as natural filters for the water system.}
{Đây là một danh từ dùng để chỉ đất ngập nước.}

\vocabentry{Wildlife conservation}
{/ˈwaɪldlaɪf ˌkɑːnsərˈveɪʃn/ (n)}
{The practice of protecting wild species and their habitats.}
{Wildlife conservation is essential for maintaining global biodiversity.}
{Đây là một cụm danh từ dùng để chỉ sự bảo tồn động vật hoang dã.}

\vocabentry{Wildlife trade}
{/ˈwaɪldlaɪf treɪd/ (n)}
{The sale and exchange of wild animals and plants.}
{Illegal wildlife trade is a multibillion-dollar global industry.}
{Đây là một cụm danh từ dùng để chỉ việc buôn bán động vật hoang dã.}

\vocabentry{Wind energy}
{/ˈwɪnd ˌenərdʒi/ (n)}
{Power obtained by harnessing the energy of the wind.}
{Wind energy is one of the fastest-growing renewable sources.}
{Đây là một cụm danh từ dùng để chỉ năng lượng gió.}

\vocabentry{Zero-waste}
{/ˌzɪroʊ ˈweɪst/ (adj)}
{Producing little or no waste.}
{Adopting a zero-waste lifestyle helps reduce environmental impact.}
{Đây là một tính từ dùng để chỉ lối sống không rác thải.}

\vocabentry{Zoology}
{/zuˈɑːlə dʒi/ (n)}
{The scientific study of the behavior, structure, and classification of animals.}
{Zoology is a field that is critical for conservation biology.}
{Đây là một danh từ dùng để chỉ động vật học.}

\vocabentry{Algal bloom}
{/ˈælɡl bluːm/ (n)}
{A rapid increase or accumulation in the population of algae in freshwater or marine water systems.}
{Fertilizer runoff can cause massive algal blooms that kill fish.}
{Đây là một cụm danh từ dùng để chỉ hiện tượng tảo nở hoa.}

\vocabentry{Ambient}
{/ˈæmbiənt/ (adj)}
{Relating to the immediate surroundings of something.}
{The ambient air quality in the city has improved recently.}
{Đây là một tính từ dùng để chỉ môi trường xung quanh.}

\vocabentry{Net-zero emissions}
{/ˌnet ˈzɪroʊ iˈmɪʃnz/ (n)}
{A state in which the greenhouse gases going into the atmosphere are balanced by removal from the atmosphere.}
{Many nations have pledged to reach net-zero emissions by mid-century.}
{Đây là một cụm danh từ dùng để chỉ phát thải ròng bằng không.}

