% Auto-generated from raw.txt
% Usage:
%   \vocabentry{word}{/ipa/ (pos)}{definition}{example}{note}

\vocabentry{Chronic}
{/ˈkrɑːnɪk/ (adj)}
{(Of an illness) persisting for a long time or constantly recurring.}
{Chronic back pain can significantly affect a person's quality of life and mental health.}
{Đây là một tính từ dùng để chỉ các bệnh mãn tính, kéo dài và khó chữa dứt điểm.}

\vocabentry{Diagnosis}
{/ˌdaɪəɡˈnoʊsɪs/ (n)}
{The identification of the nature of an illness or other problem by examination of the symptoms.}
{Early diagnosis of the disease is crucial for effective treatment and recovery.}
{Đây là một danh từ dùng để chỉ sự chẩn đoán bệnh.}

\vocabentry{Prescription}
{/prɪˈskrɪpʃn/ (n)}
{An instruction written by a medical practitioner that authorizes a patient to be provided with a medicine or treatment.}
{You cannot buy these antibiotics without a valid prescription from a licensed physician.}
{Đây là một danh từ dùng để chỉ đơn thuốc do bác sĩ kê.}

\vocabentry{Vaccination}
{/ˌvæksɪˈneɪʃn/ (n)}
{Treatment with a vaccine to produce immunity against a disease.}
{Mass vaccination programs have helped to eradicate several deadly diseases worldwide.}
{Đây là một danh từ dùng để chỉ việc tiêm chủng vắc-xin.}

\vocabentry{Epidemic}
{/ˌepɪˈdemɪk/ (n)}
{A widespread occurrence of an infectious disease in a community at a particular time.}
{The government is taking urgent measures to prevent a flu epidemic this winter.}
{Đây là một danh từ dùng để chỉ một đợt dịch bệnh bùng phát trong một cộng đồng.}

\vocabentry{Immune system}
{/ɪˈmjuːn ˌsɪstəm/ (n)}
{The organs and processes of the body that provide resistance to infection and toxins.}
{A healthy diet and regular exercise are essential for strengthening your immune system.}
{Đây là một cụm danh từ dùng để chỉ hệ miễn dịch của cơ thể.}

\vocabentry{Hygiene}
{/ˈhaɪdʒiːn/ (n)}
{Conditions or practices conducive to maintaining health and preventing disease, especially through cleanliness.}
{Good personal hygiene is the most effective way to prevent the spread of germs.}
{Đây là một danh từ dùng để chỉ vệ sinh cá nhân hoặc cộng đồng.}

\vocabentry{Nutrition}
{/nuˈtrɪʃn/ (n)}
{The process of providing or obtaining the food necessary for health and growth.}
{Poor nutrition in childhood can lead to long-term health problems in adulthood.}
{Đây là một danh từ dùng để chỉ chế độ dinh dưỡng hoặc sự cung cấp dưỡng chất.}

\vocabentry{Obesity}
{/oʊˈbiːsəti/ (n)}
{The state of being grossly fat or overweight.}
{Obesity is a major risk factor for heart disease, diabetes, and certain types of cancer.}
{Đây là một danh từ dùng để chỉ bệnh béo phì.}

\vocabentry{Sedentary}
{/ˈsednteri/ (adj)}
{(Of a person) tending to spend much time seated; somewhat inactive.}
{A sedentary lifestyle can lead to weight gain and various metabolic disorders.}
{Đây là một tính từ dùng để chỉ lối sống ít vận động, ngồi nhiều.}

\vocabentry{Therapy}
{/ˈθerəpi/ (n)}
{Treatment intended to relieve or heal a disorder.}
{He is undergoing physical therapy to regain strength in his leg after the accident.}
{Đây là một danh từ dùng để chỉ liệu pháp điều trị, bao gồm cả vật lý và tâm lý.}

\vocabentry{Antibiotic}
{/ˌæntibaɪˈɑːtɪk/ (n)}
{A medicine that inhibits the growth of or destroys microorganisms.}
{Overuse of antibiotics can lead to the development of drug-resistant bacteria.}
{Đây là một danh từ dùng để chỉ thuốc kháng sinh.}

\vocabentry{Symptoms}
{/ˈsɪmptəm/ (n)}
{A physical or mental feature which is regarded as indicating a condition of disease.}
{Common symptoms of the cold include a runny nose, cough, and sore throat.}
{Đây là một danh từ dùng để chỉ các triệu chứng của bệnh.}

\vocabentry{Infectious}
{/ɪnˈfekʃəs/ (adj)}
{(Of a disease or disease-causing organism) likely to be transmitted to people through the environment.}
{Flu is highly infectious and can spread quickly in crowded places like schools.}
{Đây là một tính từ dùng để chỉ tính lây nhiễm.}

\vocabentry{Metabolism}
{/məˈtæbəlɪzəm/ (n)}
{The chemical processes that occur within a living organism in order to maintain life.}
{Regular physical activity can help to boost your metabolism and burn more calories.}
{Đây là một danh từ dùng để chỉ sự trao đổi chất.}

\vocabentry{Check-up}
{/ˈtʃek ʌp/ (n)}
{A thorough medical examination to ensure that a person is healthy.}
{It is recommended to have a medical check-up at least once a year.}
{Đây là một danh từ dùng để chỉ việc đi khám sức khỏe tổng quát.}

\vocabentry{Fatigue}
{/fəˈtiːɡ/ (n)}
{Extreme tiredness resulting from mental or physical exertion or illness.}
{Chronic fatigue can be a symptom of an underlying medical condition.}
{Đây là một danh từ dùng để chỉ sự mệt mỏi, kiệt sức.}

\vocabentry{Mental health}
{/ˌmentl ˈhelθ/ (n)}
{A person’s condition with regard to their psychological and emotional well-being.}
{Taking care of your mental health is just as important as looking after your physical body.}
{Đây là một cụm danh từ dùng để chỉ sức khỏe tâm thần.}

\vocabentry{Wellness}
{/ˈwelnəs/ (n)}
{The state of being in good health, especially as an actively pursued goal.}
{The company offers various wellness programs to support the health of its employees.}
{Đây là một danh từ dùng để chỉ trạng thái khỏe mạnh toàn diện.}

\vocabentry{Rehabilitation}
{/ˌriːəˌbɪlɪˈteɪʃn/ (n)}
{The action of restoring someone to health or normal life through training and therapy.}
{The center provides rehabilitation services for patients recovering from a stroke.}
{Đây là một danh từ dùng để chỉ sự phục hồi chức năng.}

\vocabentry{Acute}
{/əˈkjuːt/ (adj)}
{(Of a physical disease or its symptoms) severe or intense degree.}
{He was rushed to the hospital with acute abdominal pain.}
{Tính từ chỉ bệnh cấp tính, khởi phát nhanh và dữ dội.}

\vocabentry{Ailment}
{/ˈeɪlmənt/ (n)}
{An illness, typically a minor one.}
{I have a minor ailment that requires a few days of rest and fluids.}
{Danh từ chỉ một sự ốm đau, thường là bệnh nhẹ.}

\vocabentry{Allergy}
{/ˈælərdʒi/ (n)}
{A damaging immune response by the body to a substance.}
{My brother has a severe allergy to peanuts and must be very careful.}
{Danh từ chỉ sự dị ứng với một chất nào đó.}

\vocabentry{Ambulance}
{/ˈæmbjələns/ (n)}
{A vehicle equipped for taking sick or injured people to hospital.}
{The ambulance arrived within minutes to take the injured cyclist to the ER.}
{Danh từ chỉ xe cứu thương.}

\vocabentry{Anemia}
{/əˈniːmiə/ (n)}
{A condition marked by a deficiency of red blood cells.}
{She was diagnosed with iron-deficiency anemia and prescribed supplements.}
{Danh từ chỉ bệnh thiếu máu.}

\vocabentry{Anesthesia}
{/ˌænəsˈθiːʒə/ (n)}
{Insensitivity to pain, especially as artificially induced by gases or drugs.}
{The surgeon performed the operation under general anesthesia.}
{Danh từ chỉ sự gây mê.}

\vocabentry{Antiseptic}
{/ˌæntiˈseptɪk/ (n)}
{A substance that prevents the growth of disease-causing microorganisms.}
{Clean the wound thoroughly with an antiseptic before applying a bandage.}
{Danh từ chỉ chất sát trùng.}

\vocabentry{Anxiety}
{/æŋˈzaɪəti/ (n)}
{A feeling of worry, nervousness, or unease.}
{Regular meditation can help people manage their daily anxiety levels.}
{Danh từ chỉ sự lo âu, rối loạn lo âu.}

\vocabentry{Appetite}
{/ˈæpɪtaɪt/ (n)}
{A natural desire to satisfy a bodily need, especially for food.}
{Loss of appetite is often one of the first signs of an infection.}
{Danh từ chỉ sự thèm ăn.}

\vocabentry{Artery}
{/ˈɑːrtəri/ (n)}
{Any of the muscular-walled tubes forming part of the circulation system.}
{A diet high in saturated fat can lead to the clogging of arteries.}
{Danh từ chỉ động mạch.}

\vocabentry{Asthma}
{/ˈæzmə/ (n)}
{A respiratory condition marked by attacks of spasm in the bronchi of the lungs.}
{Air pollution can trigger severe asthma attacks in vulnerable children.}
{Danh từ chỉ bệnh hen suyễn.}

\vocabentry{Bacteria}
{/bækˈtɪriə/ (n)}
{Unicellular microorganisms which have cell walls but lack organelles.}
{Not all bacteria are harmful; some are essential for our digestion.}
{Danh từ chỉ vi khuẩn.}

\vocabentry{Bandage}
{/ˈbændɪdʒ/ (n)}
{A strip of material used to bind a wound or a protected injured part of the body.}
{He wrapped a clean bandage around his sprained ankle.}
{Danh từ chỉ băng gạc.}

\vocabentry{Benign}
{/bɪˈnaɪn/ (adj)}
{(Of a disease) not harmful in effect; in particular, not malignant.}
{The biopsy results showed that the tumor was benign, much to our relief.}
{Tính từ chỉ khối u lành tính, không gây hại.}

\vocabentry{Blood pressure}
{/ˈblʌd ˌpreʃər/ (n)}
{The pressure of the blood in the circulatory system.}
{High blood pressure can increase the risk of having a stroke or heart attack.}
{Cụm danh từ chỉ huyết áp.}

\vocabentry{Blood vessel}
{/ˈblʌd ˌvesl/ (n)}
{A tubular structure carrying blood through the tissues and organs.}
{Small blood vessels are responsible for delivering oxygen to every cell.}
{Cụm danh từ chỉ mạch máu.}

\vocabentry{Body mass index}
{/ˌbɑːdi ˈmæs ˌɪndeks/ (n)}
{A weight-to-height ratio used as an indicator of obesity.}
{Your body mass index can help you determine if you are at a healthy weight.}
{Cụm danh từ chỉ chỉ số khối cơ thể - BMI.}

\vocabentry{Boost}
{/buːst/ (v)}
{Help or encourage (something) to increase or improve.}
{Vitamin C is known to help boost the body's natural defenses.}
{Động từ chỉ sự thúc đẩy, làm tăng cường sức khỏe.}

\vocabentry{Brainstorm}
{/ˈbreɪnstɔːrm/ (n)}
{A sudden clever idea; also a period of mental confusion.}
{Extreme fatigue can sometimes lead to a temporary brainstorm or confusion.}
{Trong y tế có thể chỉ cơn rối loạn tâm trí đột ngột.}

\vocabentry{Bronchitis}
{/brɑːŋˈkaɪtɪs/ (n)}
{Inflammation of the mucous membrane in the bronchial tubes.}
{Smoking is the leading cause of chronic bronchitis in adults.}
{Danh từ chỉ bệnh viêm phế quản.}

\vocabentry{Bruise}
{/bruːz/ (n)}
{An injury appearing as an area of discolored skin on the body.}
{She had a large bruise on her arm after falling off her bicycle.}
{Danh từ chỉ vết bầm tím.}

\vocabentry{Calorie}
{/ˈkæləri/ (n)}
{A unit of energy often used to measure the energy value of foods.}
{To lose weight, you need to consume fewer calories than you burn each day.}
{Danh từ chỉ đơn vị năng lượng calo.}

\vocabentry{Cancer}
{/ˈkænsər/ (n)}
{A disease caused by an uncontrolled division of abnormal cells.}
{Early screening is vital for the successful treatment of many types of cancer.}
{Danh từ chỉ bệnh ung thư.}

\vocabentry{Capsule}
{/ˈkæpsuːl/ (n)}
{A small case or container, especially a soluble one enclosing a dose of medicine.}
{Please take one capsule of this medicine twice a day with water.}
{Danh từ chỉ thuốc viên nang.}

\vocabentry{Carbohydrate}
{/ˌkɑːrboʊˈhaɪdreɪt/ (n)}
{Large group of organic compounds occurring in foods and living tissues.}
{Athletes often eat high-carbohydrate meals before a big race for energy.}
{Danh từ chỉ chất bột đường/tinh bột.}

\vocabentry{Cardiac}
{/ˈkɑːrdiæk/ (adj)}
{Relating to the heart.}
{Regular cardio exercises are beneficial for maintaining cardiac health.}
{Tính từ chỉ những gì liên quan đến tim mạch.}

\vocabentry{Cardiovascular}
{/ˌkɑːrdioʊˈvæskjələr/ (adj)}
{Relating to the heart and blood vessels.}
{Cardiovascular disease remains one of the leading causes of death globally.}
{Tính từ chỉ hệ tim mạch.}

\vocabentry{Casualty}
{/ˈkæʒuəlti/ (n)}
{A person killed or injured in a war or accident.}
{The casualties were immediately taken to the nearest hospital for treatment.}
{Danh từ chỉ thương vong hoặc phòng cấp cứu.}

\vocabentry{Checkup}
{/ˈtʃekʌp/ (n)}
{A thorough medical examination.}
{Routine checkups help in the early detection of potential health issues.}
{Lặp lại để nhấn mạnh nghĩa danh từ viết liền.}

\vocabentry{Chemotherapy}
{/ˌkiːmoʊˈθerəpi/ (n)}
{The treatment of disease by the use of chemical substances.}
{Chemotherapy is often used in combination with surgery to treat cancer.}
{Danh từ chỉ hóa trị.}

\vocabentry{Cholesterol}
{/kəˈlestərɔːl/ (n)}
{A compound found in most body tissues.}
{High levels of LDL cholesterol can lead to heart disease.}
{Danh từ chỉ cholesterol trong máu.}

\vocabentry{Clinic}
{/ˈklɪnɪk/ (n)}
{An establishment or hospital department where outpatients are given medical treatment.}
{She works as a pediatrician at a local health clinic.}
{Danh từ chỉ phòng khám.}

\vocabentry{Cognitive}
{/ˈkɑːɡnətɪv/ (adj)}
{Relating to cognition; mental action or process of acquiring knowledge.}
{Cognitive decline can be a natural part of aging for some individuals.}
{Tính từ chỉ nhận thức.}

\vocabentry{Condition}
{/kənˈdɪʃn/ (n)}
{A state of physical or mental fitness or a specific illness.}
{The patient is in a stable condition after the successful surgery.}
{Danh từ chỉ tình trạng sức khỏe hoặc bệnh lý.}

\vocabentry{Contagious}
{/kənˈteɪdʒəs/ (adj)}
{(Of a disease) spread from one person or organism to another by direct or indirect contact.}
{Chickenpox is extremely contagious among children who haven't been vaccinated.}
{Tính từ chỉ bệnh truyền nhiễm/lây lan nhanh.}

\vocabentry{Convalescence}
{/ˌkɑːnvəˈlesns/ (n)}
{Time spent recovering from an illness or medical treatment.}
{After the operation, he needed a long period of convalescence at home.}
{Danh từ chỉ thời kỳ dưỡng bệnh/hồi phục.}

\vocabentry{Cough}
{/kɔːf/ (n/v)}
{Expel air from the lungs with a sudden sharp sound.}
{She has a persistent cough that hasn't gone away for two weeks.}
{Danh từ/Động từ chỉ triệu chứng ho.}

\vocabentry{Cure}
{/kjʊr/ (n/v)}
{Relieve (a person or animal) of the symptoms of a disease or condition.}
{Scientists are still searching for a definitive cure for the common cold.}
{Danh từ/Động từ chỉ sự chữa khỏi/liều thuốc.}

\vocabentry{Dehydration}
{/ˌdiːhaɪˈdreɪʃn/ (n)}
{A harmful reduction in the amount of water in the body.}
{Drinking plenty of water during exercise prevents dehydration.}
{Danh từ chỉ sự mất nước.}

\vocabentry{Dementia}
{/dɪˈmenʃə/ (n)}
{A chronic or persistent disorder of the mental processes caused by brain disease.}
{Alzheimer's is the most common form of dementia in elderly people.}
{Danh từ chỉ chứng sa sút trí tuệ.}

\vocabentry{Dentistry}
{/ˈdentɪstri/ (n)}
{The profession or practice of a dentist.}
{Advances in dentistry have made dental procedures much less painful.}
{Danh từ chỉ nha khoa.}

\vocabentry{Depression}
{/dɪˈpreʃn/ (n)}
{A mental health disorder characterized by persistently depressed mood.}
{Depression is a serious medical illness that requires professional help.}
{Danh từ chỉ bệnh trầm cảm.}

\vocabentry{Dermatologist}
{/ˌdɜːrməˈtɑːlədʒɪst/ (n)}
{A medical practitioner qualified to diagnose and treat skin disorders.}
{You should see a dermatologist if you notice any unusual moles on your skin.}
{Danh từ chỉ bác sĩ da liễu.}

\vocabentry{Diabetes}
{/ˌdaɪəˈbiːtiːz/ (n)}
{A disease in which the body’s ability to produce or respond to the hormone insulin is impaired.}
{People with diabetes need to monitor their blood sugar levels carefully.}
{Danh từ chỉ bệnh tiểu đường.}

\vocabentry{Diet}
{/ˈdaɪət/ (n)}
{The kinds of food that a person, animal, or community habitually eats.}
{A balanced diet includes a variety of fruits, vegetables, and proteins.}
{Danh từ chỉ chế độ ăn uống.}

\vocabentry{Dietitian}
{/ˌdaɪəˈtɪʃn/ (n)}
{An expert on diet and nutrition.}
{The dietitian helped him create a meal plan to manage his weight.}
{Danh từ chỉ chuyên gia dinh dưỡng.}

\vocabentry{Digest}
{/daɪˈdʒest/ (v)}
{Break down food in the alimentary canal into substances that can be absorbed by the body.}
{It takes several hours for the body to fully digest a heavy meal.}
{Động từ chỉ sự tiêu hóa.}

\vocabentry{Disorder}
{/dɪsˈɔːrdər/ (n)}
{A state of confusion; also a physical or mental condition that is not normal.}
{Eating disorders can have severe consequences for a person's physical health.}
{Danh từ chỉ sự rối loạn chức năng hoặc bệnh tật.}

\vocabentry{Dizziness}
{/ˈdɪzinəs/ (n)}
{A sensation of spinning and losing one's balance.}
{Low blood sugar can often cause sudden bouts of dizziness.}
{Danh từ chỉ sự chóng mặt.}

\vocabentry{Dose}
{/doʊs/ (n)}
{A quantity of a medicine or drug taken or recommended to be taken at a particular time.}
{Do not exceed the recommended dose of this painkiller.}
{Danh từ chỉ liều lượng thuốc.}

\vocabentry{Drugstore}
{/ˈdrʌɡstɔːr/ (n)}
{A shop where medicinal drugs are dispensed and sold.}
{I need to stop by the drugstore to pick up some cough syrup.}
{Danh từ chỉ hiệu thuốc.}

\vocabentry{Dyslexia}
{/dɪsˈleksiə/ (n)}
{A general term for disorders that involve difficulty in learning to read or interpret words.}
{Many famous and successful people have struggled with dyslexia.}
{Danh từ chỉ chứng khó đọc.}

\vocabentry{Earache}
{/ˈɪreɪk/ (n)}
{Pain in the inside of the ear.}
{Children often get an earache after suffering from a cold.}
{Danh từ chỉ sự đau tai.}

\vocabentry{Eating disorder}
{/ˈiːtɪŋ dɪsˌɔːrdər/ (n)}
{Any of a range of psychological disorders characterized by abnormal or disturbed eating habits.}
{Anorexia and bulimia are two common types of eating disorders.}
{Cụm danh từ chỉ rối loạn ăn uống.}

\vocabentry{Elderly}
{/ˈeldərli/ (adj)}
{(Of a person) old or aging.}
{Special health care services are provided for the elderly in our community.}
{Tính từ chỉ người cao tuổi.}

\vocabentry{Emergency}
{/ɪˈmɜːrdʒənsi/ (n)}
{A serious, unexpected, and often dangerous situation requiring immediate action.}
{In case of an emergency, please dial 911 immediately.}
{Danh từ chỉ trường hợp khẩn cấp.}

\vocabentry{Emotional}
{/ɪˈmoʊʃənl/ (adj)}
{Relating to a person’s emotions.}
{Support from friends is crucial for maintaining emotional well-being.}
{Tính từ chỉ cảm xúc.}

\vocabentry{Endurance}
{/ɪnˈdʊrəns/ (n)}
{The fact or power of enduring an unpleasant or difficult process or situation without giving way.}
{Running marathons requires a high level of physical endurance.}
{Danh từ chỉ sức chịu đựng/sức bền.}

\vocabentry{Energy}
{/ˈenərdʒi/ (n)}
{The strength and vitality required for sustained physical or mental activity.}
{Eating a healthy breakfast gives you the energy you need for the day.}
{Danh từ chỉ năng lượng.}

\vocabentry{Enzyme}
{/ˈenzaɪm/ (n)}
{A substance produced by a living organism which acts as a catalyst to bring about a specific biochemical reaction.}
{Enzymes play a vital role in breaking down food during digestion.}
{Danh từ chỉ men tiêu hóa/enzym.}

\vocabentry{Examination}
{/ɪɡˌzæmɪˈneɪʃn/ (n)}
{A detailed inspection or investigation.}
{The doctor conducted a physical examination to find the cause of the pain.}
{Trong y tế là sự thăm khám bệnh.}

\vocabentry{Exercise}
{/ˈeksərsaɪz/ (n/v)}
{Activity requiring physical effort, carried out to sustain or improve health and fitness.}
{Regular exercise can help reduce the risk of many chronic diseases.}
{Danh từ/Động từ chỉ sự tập thể dục.}

\vocabentry{Exhale}
{/eksˈheɪl/ (v)}
{Breathe out.}
{Inhale deeply through your nose and exhale slowly through your mouth.}
{Động từ chỉ sự thở ra.}

\vocabentry{Expire}
{/ɪkˈspaɪər/ (v)}
{(Of a period of time) come to an end; (of a medicine) reach its used-by date.}
{Check the label to make sure the medicine hasn't expired.}
{Động từ chỉ sự hết hạn dùng thuốc.}

\vocabentry{Eye-strain}
{/ˈaɪ streɪn/ (n)}
{Fatigue of the eyes caused by excessive use or uncorrected defects of vision.}
{Working at a computer for long hours can cause severe eye-strain.}
{Danh từ chỉ sự mỏi mắt.}

\vocabentry{Fever}
{/ˈfiːvər/ (n)}
{An abnormally high body temperature.}
{A high fever is often a sign that your body is fighting an infection.}
{Danh từ chỉ cơn sốt.}

\vocabentry{First aid}
{/ˌfɜːrst ˈeɪd/ (n)}
{Help given to a sick or injured person until full medical treatment is available.}
{Everyone should learn basic first aid to help in case of an accident.}
{Cụm danh từ chỉ sơ cứu.}

\vocabentry{Fitness}
{/ˈfɪtnəs/ (n)}
{The condition of being physically fit and healthy.}
{She goes to the gym every morning to maintain her level of fitness.}
{Danh từ chỉ sự sung sức, thể trạng khỏe mạnh.}

\vocabentry{Flu}
{/fluː/ (n)}
{Influenza; a highly contagious viral infection of the respiratory passages.}
{Getting a flu shot every year is a good way to stay healthy.}
{Danh từ chỉ bệnh cúm.}

\vocabentry{Fracture}
{/ˈfræktʃər/ (n/v)}
{The cracking or breaking of a hard object or material, typically a bone.}
{The X-ray confirmed that he had a small fracture in his wrist.}
{Danh từ/Động từ chỉ sự gãy xương.}

\vocabentry{Fungus}
{/ˈfʌŋɡəs/ (n)}
{Any of a group of spore-producing organisms feeding on organic matter.}
{Athlete's foot is a common skin infection caused by a fungus.}
{Danh từ chỉ nấm/bệnh nấm.}

\vocabentry{General practitioner}
{/ˌdʒenrəl prækˈtɪʃənər/ (n)}
{A doctor who does not specialize in a particular area of medicine.}
{If you have a health problem, you should see your general practitioner first.}
{Cụm danh từ chỉ bác sĩ đa khoa - GP.}

\vocabentry{Genetic}
{/dʒəˈnetɪk/ (adj)}
{Relating to genes or heredity.}
{Some diseases are genetic and can be passed from parents to children.}
{Tính từ chỉ di truyền.}

\vocabentry{Germ}
{/dʒɜːrm/ (n)}
{A microorganism, especially one which causes disease.}
{Washing your hands frequently helps to kill germs and prevent illness.}
{Danh từ chỉ vi trùng, mầm bệnh.}

\vocabentry{Gland}
{/ɡlænd/ (n)}
{An organ which secretes particular chemical substances for use in the body. (Danh từ chỉ tuyến (tuyến giáp, tuyến mồ hôi...).).}
{The thyroid gland plays a crucial role in regulating metabolism.}
{}

\vocabentry{Glucose}
{/ˈɡluːkoʊs/ (n)}
{A simple sugar which is an important energy source in living organisms.}
{The body converts carbohydrates into glucose to provide energy for cells.}
{Danh từ chỉ đường glucose.}

\vocabentry{Handicap}
{/ˈhændikæp/ (n)}
{A condition that markedly restricts a person’s ability to function physically, mentally, or socially.}
{Technology has provided many new ways for people with a handicap to lead independent lives.}
{Danh từ chỉ sự khuyết tật.}

\vocabentry{Health}
{/helθ/ (n)}
{The state of being free from illness or injury.}
{Wealth is useless if you don't have your health.}
{Danh từ gốc chỉ sức khỏe.}

\vocabentry{Healthcare}
{/ˈhelθker/ (n)}
{The organized provision of medical care to individuals or a community.}
{Access to affordable healthcare is a major issue in many countries.}
{Danh từ chỉ dịch vụ y tế/chăm sóc sức khỏe.}

\vocabentry{Healthy}
{/ˈhelθi/ (adj)}
{In good health.}
{Eating a healthy diet is essential for maintaining a long life.}
{Tính từ chỉ sự khỏe mạnh.}

\vocabentry{Heart attack}
{/ˈhɑːrt əˌtæk/ (n)}
{A sudden and sometimes fatal occurrence of coronary thrombosis, typically resulting in the death of part of a heart muscle.}
{He was hospitalized after suffering a minor heart attack.}
{Cụm danh từ chỉ cơn đau tim.}

\vocabentry{Heartbeat}
{/ˈhɑːrtbiːt/ (n)}
{The pulsation of the heart.}
{Your heartbeat increases when you are exercising or stressed.}
{Danh từ chỉ nhịp tim.}

\vocabentry{Hereditary}
{/həˈredɪteri/ (adj)}
{(Of a characteristic or disease) determined by genetic factors and therefore able to be passed on from parents to offspring.}
{Some forms of high blood pressure are hereditary.}
{Tính từ chỉ tính di truyền.}

\vocabentry{Homeopathy}
{/ˌhoʊmiˈɑːpəθi/ (n)}
{The treatment of disease by minute doses of natural substances.}
{Many people use homeopathy as a complementary therapy alongside conventional medicine.}
{Danh từ chỉ phép chữa bệnh bằng phương pháp vi lượng đồng căn.}

\vocabentry{Hormone}
{/ˈhɔːrmoʊn/ (n)}
{A regulatory substance produced in an organism and transported in tissue fluids.}
{Hormones control many important functions in the body, including growth and reproduction.}
{Danh từ chỉ hoóc-môn/nội tiết tố.}

\vocabentry{Hospice}
{/ˈhɑːspɪs/ (n)}
{A home providing care for the sick, especially the terminally ill.}
{The hospice provides a peaceful and caring environment for patients in their final days.}
{Danh từ chỉ nhà tế bần/nơi chăm sóc bệnh nhân giai đoạn cuối.}

\vocabentry{Hospital}
{/ˈhɑːspɪtl/ (n)}
{An institution providing medical and surgical treatment and nursing care for sick or injured people.}
{She was taken to the hospital for further tests and observation.}
{Danh từ chỉ bệnh viện.}

\vocabentry{Hydrotherapy}
{/ˌhaɪdroʊˈθerəpi/ (n)}
{The use of exercises in a pool as part of treatment for conditions such as arthritis or partial paralysis.}
{Hydrotherapy can be very effective for relieving joint pain and stiffness.}
{Danh từ chỉ thủy liệu pháp.}

\vocabentry{Hypothermia}
{/ˌhaɪpoʊˈθɜːrmiə/ (n)}
{The condition of having an abnormally low body temperature.}
{Hikers need to be careful of hypothermia when traveling in cold and wet conditions.}
{Danh từ chỉ chứng hạ thân nhiệt.}

\vocabentry{Immune}
{/ɪˈmjuːn/ (adj)}
{Resistant to a particular infection or toxin owing to the presence of specific antibodies.}
{Vaccination makes you immune to certain infectious diseases.}
{Tính từ chỉ sự miễn dịch.}

\vocabentry{Immunity}
{/ɪˈmjuːnəti/ (n)}
{The ability of an organism to resist a particular infection.}
{Babies get some natural immunity from their mother's milk.}
{Danh từ chỉ sự miễn dịch.}

\vocabentry{Immunize}
{/ˈɪmjunaɪz/ (v)}
{Make (a person or animal) immune to infection, typically by inoculation.}
{It is important to immunize children against common childhood diseases.}
{Động từ chỉ việc gây miễn dịch.}

\vocabentry{Incurable}
{/ɪnˈkjʊrəbl/ (adj)}
{(Of a sick person or a disease) not able to be cured.}
{Some forms of cancer are still incurable despite advances in medical science.}
{Tính từ chỉ bệnh nan y, không thể chữa khỏi.}

\vocabentry{Indigestion}
{/ˌɪndɪˈdʒestʃən/ (n)}
{Pain or discomfort in the stomach associated with difficulty in digesting food.}
{Eating too quickly can often cause a painful bout of indigestion.}
{Danh từ chỉ chứng khó tiêu.}

\vocabentry{Infection}
{/ɪnˈfekʃn/ (n)}
{The process of infecting or the state of being infected.}
{Use an antiseptic to clean the cut and prevent an infection.}
{Danh từ chỉ sự nhiễm trùng.}

\vocabentry{Inflammation}
{/ˌɪnfləˈmeɪʃn/ (n)}
{A physical condition in which part of the body becomes reddened, swollen, hot, and often painful.}
{Arthritis causes painful inflammation of the joints.}
{Danh từ chỉ sự viêm nhiễm.}

\vocabentry{Influenza}
{/ˌɪnfluˈenzə/ (n)}
{Flu; a highly contagious viral infection of the respiratory passages.}
{The influenza virus can mutate rapidly, making it difficult to control.}
{Danh từ đầy đủ của bệnh cúm.}

\vocabentry{Inhale}
{/ɪnˈheɪl/ (v)}
{Breathe in.}
{Deep breathing exercises help you inhale more oxygen and relax.}
{Động từ chỉ sự hít vào.}

\vocabentry{Injection}
{/ɪnˈdʒekʃn/ (n)}
{An instance of injecting or being injected.}
{The nurse gave him a quick injection to relieve the pain.}
{Danh từ chỉ sự tiêm thuốc.}

\vocabentry{Injure}
{/ˈɪndʒər/ (v)}
{Do physical harm or damage to (someone).}
{Be careful not to injure your back when lifting heavy objects.}
{Động từ chỉ việc làm bị thương.}

\vocabentry{Injury}
{/ˈɪndʒəri/ (n)}
{An instance of being injured.}
{He suffered a serious head injury in the car accident.}
{Danh từ chỉ vết thương.}

\vocabentry{Inpatient}
{/ˈɪnpeɪʃnt/ (n)}
{A patient who stays in a hospital while receiving treatment.}
{The hospital has over five hundred beds for inpatients.}
{Danh từ chỉ bệnh nhân nội trú.}

\vocabentry{Insomnia}
{/ɪnˈsɑːmniə/ (n)}
{Habitual sleeplessness; inability to sleep.}
{Stress and anxiety are common causes of chronic insomnia.}
{Danh từ chỉ chứng mất ngủ.}

\vocabentry{Insulin}
{/ˈɪnsəlɪn/ (n)}
{A hormone produced in the pancreas by the islets of Langerhans.}
{People with Type 1 diabetes need daily injections of insulin.}
{Danh từ chỉ hoóc-môn insulin.}

\vocabentry{Intensive care}
{/ɪnˌtensɪv ˈker/ (n)}
{Specialized medical facilities and services for patients who are seriously ill.}
{After the surgery, the patient was moved to the intensive care unit.}
{Danh từ chỉ sự chăm sóc đặc biệt/hồi sức cấp cứu.}

\vocabentry{Internal}
{/ɪnˈtɜːrnl/ (adj)}
{Of or situated on the inside.}
{Internal bleeding can be a life-threatening emergency.}
{Tính từ chỉ nội tạng/bên trong cơ thể.}

\vocabentry{Invalid}
{/ˈɪnvəlɪd/ (n)}
{A person made weak or disabled by illness or injury.}
{He spent his final years as an invalid, cared for by his family.}
{Danh từ chỉ người tàn tật hoặc đau yếu lâu ngày.}

\vocabentry{Itch}
{/ɪtʃ/ (n/v)}
{An uncomfortable sensation on the skin that causes a desire to scratch.}
{Mosquito bites often cause an annoying itch that lasts for several days.}
{Danh từ/Động từ chỉ sự ngứa ngáy.}

\vocabentry{Jaundice}
{/ˈdʒɔːndɪs/ (n)}
{A medical condition with yellowing of the skin or whites of the eyes.}
{Jaundice can be a sign of a liver or gallbladder problem.}
{Danh từ chỉ bệnh vàng da.}

\vocabentry{Joint}
{/dʒɔɪnt/ (n)}
{A structure in the human or animal body at which two parts of the skeleton are fitted together.}
{Regular stretching can help keep your joints flexible and healthy.}
{Danh từ chỉ khớp xương.}

\vocabentry{Kidney}
{/ˈkɪdni/ (n)}
{Each of a pair of organs in the abdominal cavity that excrete urine.}
{Drinking plenty of water is essential for maintaining kidney health.}
{Danh từ chỉ quả thận.}

\vocabentry{Laceration}
{/ˌlæsəˈreɪʃn/ (n)}
{A deep cut or tear in skin or flesh.}
{The doctor needed to put several stitches in the laceration on his arm.}
{Danh từ chỉ vết rách da sâu.}

\vocabentry{Latent}
{/ˈleɪtnt/ (adj)}
{(Of a quality or state) existing but not yet developed or manifest; hidden or concealed.}
{The virus can remain latent in the body for many years before symptoms appear.}
{Trong y tế chỉ mầm bệnh đang ủ bệnh.}

\vocabentry{Laxative}
{/ˈlæksətɪv/ (n)}
{A medicine which has a laxative effect.}
{Natural laxatives like fiber and water are usually better than chemical ones.}
{Danh từ chỉ thuốc nhuận tràng.}

\vocabentry{Lethargic}
{/ləˈθɑːrdʒɪk/ (adj)}
{Affected by lethargy; sluggish and apathetic.}
{Iron deficiency can make you feel very tired and lethargic.}
{Tính từ chỉ sự lờ đờ, uể oải.}

\vocabentry{Leukemia}
{/luːˈkiːmiə/ (n)}
{A malignant progressive disease in which the bone marrow and other blood-forming organs produce increased numbers of immature or abnormal leukocytes.}
{Leukemia is a type of cancer that affects the blood and bone marrow.}
{Danh từ chỉ bệnh bạch cầu.}

\vocabentry{Lifespan}
{/ˈlaɪfspæn/ (n)}
{The length of time for which a person or animal lives or a thing functions.}
{Modern medicine has significantly increased the average human lifespan.}
{Danh từ chỉ tuổi thọ.}

\vocabentry{Ligament}
{/ˈlɪɡəmənt/ (n)}
{A short band of tough, flexible fibrous connective tissue which connects two bones.}
{He tore a ligament in his knee while playing soccer.}
{Danh từ chỉ dây chằng.}

\vocabentry{Liver}
{/ˈlɪvər/ (n)}
{A large lobed glandular organ in the abdomen of vertebrates.}
{Excessive alcohol consumption can cause serious damage to the liver.}
{Danh từ chỉ lá gan.}

\vocabentry{Longevity}
{/lɔːnˈdʒevəti/ (n)}
{Long life.}
{A healthy lifestyle and positive attitude are keys to longevity.}
{Danh từ chỉ sự trường thọ.}

\vocabentry{Lump}
{/lʌmp/ (n)}
{A swelling under the skin, especially one caused by injury or disease.}
{You should see a doctor if you find a lump anywhere on your body.}
{Danh từ chỉ cục bướu hoặc vết sưng.}

\vocabentry{Lung}
{/lʌŋ/ (n)}
{Each of the pair of organs situated within the rib cage.}
{Smoking increases the risk of developing lung cancer and other respiratory diseases.}
{Danh từ chỉ lá phổi.}

\vocabentry{Malaria}
{/məˈleriə/ (n)}
{An intermittent and remittent fever caused by a protozoan parasite that invades the red blood cells.}
{Malaria is transmitted to humans through the bites of infected mosquitoes.}
{Danh từ chỉ bệnh sốt rét.}

\vocabentry{Malignant}
{/məˈlɪɡnənt/ (adj)}
{(Of a disease) very virulent or infectious.}
{A malignant tumor can spread to other parts of the body if not treated early.}
{Tính từ chỉ khối u ác tính.}

\vocabentry{Malnutrition}
{/ˌmælnuˈtrɪʃn/ (n)}
{Lack of proper nutrition, caused by not having enough to eat, not eating enough of the right things, or being unable to use the food that one does eat.}
{Malnutrition is a serious problem in many developing countries.}
{Danh từ chỉ sự suy dinh dưỡng.}

\vocabentry{Medical}
{/ˈmedɪkl/ (adj)}
{Relating to the science or practice of medicine.}
{He decided to pursue a medical career and become a surgeon.}
{Tính từ chỉ những gì thuộc về y tế.}

\vocabentry{Medication}
{/ˌmedɪˈkeɪʃn/ (n)}
{A substance used for medical treatment, especially a medicine or drug.}
{Always follow the instructions provided with your medication carefully.}
{Danh từ chỉ dược phẩm/thuốc men.}

\vocabentry{Medicine}
{/ˈmedɪsn/ (n)}
{The science or practice of the diagnosis, treatment, and prevention of disease.}
{He has been practicing medicine for over thirty years.}
{Danh từ chỉ ngành y hoặc thuốc.}

\vocabentry{Melanoma}
{/ˌmeləˈnoʊmə/ (n)}
{A tumor of melanin-forming cells, typically a malignant tumor associated with skin cancer.}
{Overexposure to the sun can increase the risk of developing melanoma.}
{Danh từ chỉ ung thư hắc tố da.}

\vocabentry{Migraine}
{/ˈmaɪɡreɪn/ (n)}
{A recurrent throbbing headache that typically affects one side of the head.}
{For some people, certain foods or smells can trigger a severe migraine.}
{Danh từ chỉ chứng đau nửa đầu.}

\vocabentry{Mineral}
{/ˈmɪnərəl/ (n)}
{An inorganic substance that is essential for the healthy functioning of the body.}
{Calcium is an essential mineral for building strong bones and teeth.}
{Danh từ chỉ chất khoáng.}

\vocabentry{Mobility}
{/moʊˈbɪləti/ (n)}
{The ability to move or be moved freely and easily.}
{Physical therapy can help to improve mobility in elderly patients.}
{Danh từ chỉ khả năng vận động.}

\vocabentry{Muscle}
{/ˈmʌsl/ (n)}
{A band or bundle of fibrous tissue in a human or animal body that has the ability to contract.}
{Regular strength training helps to build and maintain healthy muscles.}
{Danh từ chỉ cơ bắp.}

\vocabentry{Nausea}
{/ˈnɔːziə/ (n)}
{A feeling of sickness with an inclination to vomit.}
{Sea travel often causes nausea and dizziness for many people.}
{Danh từ chỉ sự buồn nôn.}

\vocabentry{Neurological}
{/ˌnʊrəˈlɑːdʒɪkl/ (adj)}
{Relating to the anatomy, functions, and organic disorders of nerves and the nervous system.}
{Parkinson's disease is a progressive neurological disorder.}
{Tính từ chỉ thuộc về hệ thần kinh.}

\vocabentry{Neurosis}
{/nʊˈroʊsɪs/ (n)}
{A relatively mild mental illness that is not caused by organic disease.}
{Anxiety neurosis can manifest as physical symptoms like palpitations or sweating.}
{Danh từ chỉ chứng loạn thần kinh chức năng.}

\vocabentry{Nightmare}
{/ˈnaɪtmer/ (n)}
{A frightening or unpleasant dream or experience. (Danh từ chỉ cơn ác mộng (có thể do tâm lý).).}
{Stress and trauma can often lead to recurring nightmares.}
{}

\vocabentry{Nurse}
{/nɜːrs/ (n/v)}
{A person trained to care for the sick or infirm, especially in a hospital.}
{The nurses at the hospital were very kind and helpful during my stay.}
{Danh từ/Động từ chỉ y tá hoặc sự chăm sóc người bệnh.}

\vocabentry{Nursery}
{/ˈnɜːrsəri/ (n)}
{A room in a house or other building set apart for the use of babies and children.}
{The hospital's newborn nursery is located on the second floor.}
{Thường liên quan đến chăm sóc trẻ em trong y tế.}

\vocabentry{Nurture}
{/ˈnɜːrtʃər/ (v)}
{Care for and encourage the growth or development of.}
{Parents nurture their children to help them grow up healthy and strong.}
{Động từ chỉ sự nuôi dưỡng/chăm sóc.}

\vocabentry{Nutrient}
{/ˈnuːtriənt/ (n)}
{A substance that provides nourishment essential for growth and the maintenance of life.}
{Fruits and vegetables are rich in essential nutrients like vitamins and minerals.}
{Danh từ chỉ chất dinh dưỡng.}

\vocabentry{Nutritious}
{/nuˈtrɪʃəs/ (adj)}
{Efficient as food; nourishing.}
{Eggs are a very nutritious food that is high in protein and vitamins.}
{Tính từ chỉ sự bổ dưỡng.}

\vocabentry{Ointment}
{/ˈɔɪntmənt/ (n)}
{A smooth oily preparation that is rubbed on the skin for medicinal purposes.}
{Apply the ointment to the affected area twice a day.}
{Danh từ chỉ thuốc mỡ.}

\vocabentry{Operation}
{/ˌɑːpəˈreɪʃn/ (n)}
{An act of surgery performed on a patient.}
{He had a minor operation to remove a lump on his arm.}
{Danh từ chỉ ca phẫu thuật.}

\vocabentry{Optician}
{/ɑːpˈtɪʃn/ (n)}
{A person qualified to prescribe and dispense glasses and contact lenses.}
{You should visit an optician regularly to have your eyesight checked.}
{Danh từ chỉ chuyên gia nhãn khoa/người làm kính mắt.}

\vocabentry{Orthopedic}
{/ˌɔːrθəˈpiːdɪk/ (adj)}
{Relating to the branch of medicine dealing with the correction of deformities of bones or muscles.}
{He needed orthopedic surgery to repair his broken leg.}
{Tính từ chỉ thuộc về chỉnh hình.}

\vocabentry{Outpatient}
{/ˈaʊtpeɪʃnt/ (n)}
{A patient who attends a hospital for treatment without staying overnight.}
{The outpatient clinic is open from 8 am to 5 pm every weekday.}
{Danh từ chỉ bệnh nhân ngoại trú.}

\vocabentry{Over-the-counter}
{/ˌoʊvər ðə ˈkaʊntər/ (adj)}
{(Of a drug) sold legally without a prescription.}
{You can buy common pain relievers over-the-counter at any drugstore.}
{Tính từ chỉ thuốc bán không cần đơn bác sĩ.}

\vocabentry{Overweight}
{/ˌoʊvərˈweɪt/ (adj)}
{Above a weight considered normal or desirable.}
{Being overweight can put extra strain on your heart and joints.}
{Tính từ chỉ sự thừa cân.}

\vocabentry{Oxygen}
{/ˈɑːksɪdʒən/ (n)}
{A colorless, odorless reactive gas, the chemical element of atomic number 8 and the life-supporting component of the air.}
{Deep breathing exercises help increase the level of oxygen in your blood.}
{Danh từ chỉ khí oxy.}

\vocabentry{Pacemaker}
{/ˈpeɪsmeɪkər/ (n)}
{An artificial device for stimulating the heart muscle and regulating its contractions.}
{He had a pacemaker fitted to help regulate his irregular heartbeat.}
{Danh từ chỉ máy trợ tim.}

\vocabentry{Painkiller}
{/ˈpeɪnkɪlər/ (n)}
{A medicine used to relieve pain.}
{Aspirin is a commonly used painkiller for headaches and minor aches.}
{Danh từ chỉ thuốc giảm đau.}

\vocabentry{Palliative care}
{/ˈpæliətɪv ker/ (n)}
{Specialized medical care for people living with a serious illness.}
{Palliative care focuses on providing relief from the symptoms and stress of the illness.}
{Cụm danh từ chỉ chăm sóc giảm nhẹ.}

\vocabentry{Pancreas}
{/ˈpæŋkriəs/ (n)}
{A large gland behind the stomach which secretes digestive enzymes.}
{The pancreas produces insulin, which is essential for regulating blood sugar.}
{Danh từ chỉ tuyến tụy.}

\vocabentry{Pandemic}
{/pænˈdemɪk/ (n)}
{(Of a disease) prevalent over a whole country or the world.}
{The 1918 Spanish flu pandemic killed millions of people worldwide.}
{Danh từ chỉ đại dịch toàn cầu.}

\vocabentry{Paralysis}
{/pəˈræləsɪs/ (n)}
{The loss of the ability to move (and sometimes to feel anything) in part or most of the body.}
{A spinal cord injury can result in permanent paralysis of the legs.}
{Danh từ chỉ sự liệt/tê liệt.}

\vocabentry{Pathogen}
{/ˈpæθədʒən/ (n)}
{A bacterium, virus, or other microorganism that can cause disease.}
{Pathogens can enter the body through contaminated food or water.}
{Danh từ chỉ mầm bệnh/tác nhân gây bệnh.}

\vocabentry{Patient}
{/ˈpeɪʃnt/ (n)}
{A person receiving or registered to receive medical treatment.}
{The doctor spent a lot of time talking to each patient and answering their questions.}
{Danh từ chỉ bệnh nhân.}

\vocabentry{Pediatrician}
{/ˌpiːdiəˈtrɪʃn/ (n)}
{A medical practitioner specializing in children and their diseases.}
{Our pediatrician is very good with children and always puts them at ease.}
{Danh từ chỉ bác sĩ nhi khoa.}

\vocabentry{Pharmaceutical}
{/ˌfɑːrməˈsuːtɪkl/ (adj)}
{Relating to medicinal drugs, or their preparation, use, or sale.}
{The pharmaceutical industry is constantly developing new and better medicines.}
{Tính từ chỉ thuộc về dược phẩm.}

\vocabentry{Pharmacist}
{/ˈfɑːrməsɪst/ (n)}
{A person who is professionally qualified to prepare and dispense medicinal drugs.}
{The pharmacist can give you advice on how to take your medicine correctly.}
{Danh từ chỉ dược sĩ.}

\vocabentry{Pharmacy}
{/ˈfɑːrməsi/ (n)}
{A shop or hospital department where medicinal drugs are prepared or sold.}
{I need to go to the pharmacy to pick up my prescription.}
{Danh từ chỉ nhà thuốc/khoa dược.}

\vocabentry{Physical}
{/ˈfɪzɪkl/ (adj)}
{Relating to the body as opposed to the mind.}
{Regular physical activity is essential for maintaining a healthy weight.}
{Tính từ chỉ thuộc về thể chất.}

\vocabentry{Physician}
{/fɪˈzɪʃn/ (n)}
{A person qualified to practice medicine; a doctor.}
{He is a highly respected physician who specializes in internal medicine.}
{Danh từ chỉ bác sĩ y khoa.}

\vocabentry{Physiotherapy}
{/ˌfɪzioʊˈθerəpi/ (n)}
{The treatment of disease, injury, or deformity by physical methods.}
{Physiotherapy can help to restore movement and function after an injury.}
{Danh từ chỉ vật lý trị liệu.}

\vocabentry{Placebo}
{/pləˈsiːboʊ/ (n)}
{A harmless pill, medicine, or procedure prescribed more for the psychological benefit to the patient than for any physiological effect.}
{In some medical trials, one group of patients is given a placebo.}
{Danh từ chỉ giả dược.}

\vocabentry{Pneumonia}
{/nuːˈmoʊniə/ (n)}
{Lung inflammation caused by bacterial or viral infection.}
{Pneumonia can be a very serious illness, especially for elderly people.}
{Danh từ chỉ bệnh viêm phổi.}

\vocabentry{Poison}
{/ˈpɔɪzn/ (n/v)}
{A substance that is capable of causing the illness or death of a living organism when introduced or absorbed.}
{Many common household products contain substances that are poisonous if swallowed.}
{Danh từ/Động từ chỉ chất độc hoặc sự đầu độc.}

\vocabentry{Post-traumatic}
{/ˌpoʊst trəˈmætɪk/ (adj)}
{Occurring after a traumatic event.}
{Post-traumatic stress disorder (PTSD) is a serious condition that can affect soldiers or accident victims.}
{Tính từ chỉ hậu sang chấn tâm lý.}

\vocabentry{Practitioner}
{/prækˈtɪʃənər/ (n)}
{A person actively engaged in an art, discipline, or profession, especially medicine.}
{She is a skilled practitioner of traditional Chinese medicine.}
{Danh từ chỉ người hành nghề, thường là bác sĩ.}

\vocabentry{Prevent}
{/prɪˈvent/ (v)}
{Keep (something) from happening or arising.}
{Vaccination is the most effective way to prevent the spread of infectious diseases.}
{Động từ chỉ sự ngăn ngừa bệnh tật.}

\vocabentry{Prevention}
{/prɪˈvenʃn/ (n)}
{The action of stopping something from happening or arising.}
{Prevention is always better than cure when it comes to your health.}
{Danh từ chỉ sự ngăn ngừa/phòng bệnh.}

\vocabentry{Primary care}
{/ˌpraɪmeri ˈker/ (n)}
{Health care at a basic rather than specialized level.}
{Primary care services are provided by general practitioners and nurses.}
{Danh từ chỉ chăm sóc sức khỏe ban đầu.}

\vocabentry{Prognosis}
{/prɑːɡˈnoʊsɪs/ (n)}
{The likely course of a disease or ailment.}
{With early treatment, the prognosis for most patients is very good.}
{Danh từ chỉ sự tiên lượng bệnh.}

\vocabentry{Protein}
{/ˈproʊtiːn/ (n)}
{Large group of organic compounds which are essential for the body's structure and function.}
{Lean meat, fish, and beans are all excellent sources of protein.}
{Danh từ chỉ chất đạm/protein.}

\vocabentry{Psychological}
{/ˌsaɪkəˈlɑːdʒɪkl/ (adj)}
{Relating to the mental and emotional state of a person.}
{Stress can have both physical and psychological effects on your health.}
{Tính từ chỉ thuộc về tâm lý.}

\vocabentry{Pulse}
{/pʌls/ (n)}
{A rhythmical throbbing of the arteries as blood is propelled through them.}
{The doctor checked his pulse to see how fast his heart was beating.}
{Danh từ chỉ mạch đập.}

\vocabentry{Quarantine}
{/ˈkwɔːrəntiːn/ (n/v)}
{A state, period, or place of isolation in which people or animals that have arrived from elsewhere or been exposed to infectious or contagious disease are placed.}
{The passengers were placed in quarantine for two weeks to prevent the spread of the virus.}
{Danh từ/Động từ chỉ sự cách ly kiểm dịch.}

\vocabentry{Radiation}
{/ˌreɪdiˈeɪʃn/ (n)}
{The emission of energy as electromagnetic waves or as moving subatomic particles.}
{Radiation therapy is a common treatment for many types of cancer.}
{Danh từ chỉ bức xạ/phóng xạ/xạ trị.}

\vocabentry{Rash}
{/ræʃ/ (n)}
{An area of reddening of a person's skin, sometimes with raised spots, appearing especially as a result of a medical condition.}
{He developed a red rash on his arms after eating some strawberries.}
{Danh từ chỉ vết phát ban.}

